### 2020-03-18 korea.tex
pushing way too many stuff..

\documentclass[12pt]{article}
\usepackage[top=1in,left=1in, right = 1in, footskip=1in]{geometry}

\usepackage{graphicx}
%\usepackage{adjustbox}

%% \newcommand{\comment}{\showcomment}
\newcommand{\comment}{\nocomment}

\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
\newcommand{\nocomment}[3]{}

\newcommand{\jd}[1]{\comment{cyan}{JD}{#1}}
\newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}

\newcommand{\eref}[1]{(Eq.~\ref{eq:#1})}
\newcommand{\fref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\Fref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\sref}[1]{Sec.~\ref{#1}}
\newcommand{\frange}[2]{Fig.~\ref{fig:#1}--\ref{fig:#2}}
\newcommand{\tref}[1]{Table~\ref{tab:#1}}
\newcommand{\tlab}[1]{\label{tab:#1}}
\newcommand{\seminar}{SE\mbox{$^m$}I\mbox{$^n$}R}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

% \usepackage{lineno}
% \linenumbers

\usepackage[pdfencoding=auto, psdextra]{hyperref}

\usepackage{natbib}
\bibliographystyle{chicago}
\date{\today}

\usepackage{xspace}
\newcommand*{\ie}{i.e.\@\xspace}

\usepackage{color}

\newcommand{\Rx}[1]{\ensuremath{{\mathcal R}_{#1}}} 
\newcommand{\Ro}{\Rx{0}}
\newcommand{\RR}{\ensuremath{{\mathcal R}}}
\newcommand{\Rhat}{\ensuremath{{\hat\RR}}}
\newcommand{\tsub}[2]{#1_{{\textrm{\tiny #2}}}}

\begin{document}

\begin{flushleft}{
	\Large
	\textbf\newline{
		Estimating the effects of social distancing on preventing the spread of coronavirus disease 2019 (COVID-19) in South Korea
	}
}
\end{flushleft}

\pagebreak

Here, we use traffic data to compare and contrast the potential role of social distancing in preventing the spread of COVID-19 in two geographically separated major cities in South Korea.

\section{Data description}

Here, we analyze epidemiological data, collected bewteen January 20--March 16, 2020, describing the COVID-19 outbreak in Korea.
Daily number of reported cases in each geographic region was transcribed from press releases by Korea Centers for Disease Control and Prevention (KCDC).
Partial line lists were transcribed from press releases and reports from KCDC and various local and provincial government websites.
All data and original reports are stored in a publicly available GitHub repoository: https://github.com/parksw3/COVID19-Korea.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{figure_compare_report.pdf}
\caption{
\textbf{Comparison of epidemiological and traffic data from Daegu and Seoul.}
Solid lines represent the daily metro traffic volume in 2020 (red) and in 2017--2019 (gray).
Black lines represent the average daily metro traffic volume in previous years.
}
\label{fig:data}
\end{figure}

We focus our analysis on two cities in which the biggest number of COVID-19 cases have been reported: Daegu and Seoul.
Bewteen January 20--March 16, 2020, 6,083 cases were reported from Daegu and 248 from Seoul.
Need one or two sentences about epidemic.

Daily metro traffic from Daegu and Seoul were obtained from their corresponding government websites.
Need one or two sentences about metro. 
About 80\% decrease in Daegu and 50\% decrease in Seoul after the identification of the church-related case.

%% https://www.data.go.kr/dataset/15002503/fileData.do
%% https://data.seoul.go.kr/dataList/OA-12914/S/1/datasetView.do

\section{Effective reproduction number}

We first estimate the time-varying effective reproduction number $R_t$ for both epidemics.
We account for changes in case definition by scaling the number of reported cases by the relative detection rate (i.e., a proportion of cases tested positive based a case definition divided by the mean detection rate among case definitions); 
using a narrow case definition is more likely to have a higher detection rate and a lower reporting rate than a wide one.
Likewise, we account for changes in the number of tests by scaling the number of reported cases again by the relative number of tests performed on each day; this step is performed separately for each case definition because the widening of a case definition necessarily increases the number of tests.

We then reconstruct the incidence time series from the scaled time series of reported cases by using the backward confirmation-to-onset delay distribution, inferred from the partial line list, and previously estimated incubation period distribution.
To account for right-censoring (i.e., cases that are infected but have not been reported yet), we estimate the forward onset-to-confirmation delay distribution and estimate the probability that a case infected on a given day will be reported before March 16, 2020. 
Then, we divide the reconstructed incidence time series by this probability and estimate the effective reproduction number using the renewal equation using a 14-day sliding window.
We only calculate the effective reproduction number until March 10, 2020 because the degree of right-censoring too large afterwards.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_compare_R_t.pdf}
\caption{
\textbf{Comparison of effective reproduction number and reconstructed incidence time series in Daegu and Seoul.}
}
\label{fig:eff}
\end{figure}

\fref{eff} compares the estimates of $R_t$ in Daegu and Seoul.
While $R_t$ drops below 1 in Daegu shortly after its first case was confirmed (February 18, 2020), 
the decrease is unclear in Seoul.
Similar patterns in $R_t$ can be found in adjacent provinces -- a sharp decrease in Gyeongsangbuk-do, which surrounds Daegu, and a persistent $R_t$ around 1 in Gyeonggi-do, which surrounds Seoul (Supplementary Materials).

\section{Effect of social distancing on the effective reproduction number}

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_traffic_regression.pdf}
\caption{
\textbf{Relationship between daily traffic volume and effective reproduction number.}
}
\label{fig:traffic}
\end{figure}

\begin{itemize}
  \item Traffic negatively correlated with $R_t$
  \item Geographically differential effect of traffic on $R_t$; other routes of transportation, other factors we don't consider
  \item Even if we can get traffic to 0 in Seoul, $R_t$ is not less than 1...?
\end{itemize}

\section{Discussion}

\begin{itemize}
  \item Provides indirect evidence of effect of social distancing on epidemic dynamics
  \item But social distancing alone might not be sufficient... others have said something
  \item SK might need stronger intervention in some places
\end{itemize}

Caveats:
\begin{itemize}
  \item Only two cities...although two coupled provinces show strikingly similar patterns
  \item Correlation not causation
  \item Did not account for geographic heterogeneity in testing or etc. Estimation of delay distribution etc all based on national line list.
\end{itemize}

\pagebreak

\section{Supplementary Materials}

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_R_t_all.pdf}
\caption{
\textbf{Comparison of effective reproduction number in Daegu, Seoul, Gyeongsangbuk-do, and Gyeonggi-do.}
}
\label{fig:eff}
\end{figure}

\end{document}

### 2020-03-18 R_t_daegu_censor_raw.R

@@ -1,0 +1,50 @@
+library(dplyr)
+load("reconstruct_time_series_daegu_raw.rda")
+load("degree_censor.rda")
+
+reconstruct_censor_raw <- R_t_daegu_censor_raw <- vector('list', length(reconstruct_time_series_daegu_raw))
+
+set.seed(101)
+for (i in 1:length(reconstruct_time_series_daegu_raw)) {
+  rr <- reconstruct_time_series_daegu_raw[[i]]
+  
+  # pgamma(7, 62, 62/5) - pgamma(4, 62, 62/5) approx .95
+  
+  meangen <- rgamma(1, 62, 62/5)
+  size <- rgamma(1, 20, 20/5)
+  
+  gen <- dnbinom(0:13, mu=meangen, size=size)
+  gen <- gen/sum(gen)
+  
+  recondata <- rr %>%
+    group_by(infection) %>%
+    summarize(
+      case=length(infection)
+    ) %>%
+    rename(
+      date=infection
+    ) %>%
+    filter(date >= as.Date("2020-01-19"), date <= as.Date("2020-03-10"))
+  
+  dd <- data.frame(
+    date=seq.Date(as.Date("2020-01-19"), as.Date("2020-03-10"), 1),
+    case=0
+  )
+  
+  dd$case[match(recondata$date, dd$date)] <- recondata$case
+  dd$case <- dd$case/(1-degree_censor$median)
+  
+  reconstruct_censor_raw[[i]] <- dd
+  
+  IRt <- tail(dd$case, -14)/sapply(1:(length(dd$case)-14), function(x) sum(dd$case[x:(x+13)]*rev(gen)))
+  
+  Rt <- sapply(1:(length(IRt)-13), function(x) sum(IRt[x:(x+13)]*gen))
+  
+  R_t_daegu_censor_raw[[i]] <- data.frame(
+    date=seq.Date(as.Date("2020-01-19")+14, as.Date("2020-03-10"), 1),
+    IRt=IRt,
+    Rt=c(Rt, rep(NA, 13))
+  )
+}
+
+save("R_t_daegu_censor_raw", "reconstruct_censor_raw", file="R_t_daegu_censor_raw.rda")

### 2020-03-18 R_t_gyeonggi_censor.R

@@ -1,0 +1,50 @@
+library(dplyr)
+load("reconstruct_time_series_gyeonggi.rda")
+load("degree_censor.rda")
+
+reconstruct_censor <- R_t_gyeonggi_censor <- vector('list', length(reconstruct_time_series_gyeonggi))
+
+set.seed(101)
+for (i in 1:length(reconstruct_time_series_gyeonggi)) {
+  rr <- reconstruct_time_series_gyeonggi[[i]]
+  
+  # pgamma(7, 62, 62/5) - pgamma(4, 62, 62/5) approx .95
+  
+  meangen <- rgamma(1, 62, 62/5)
+  size <- rgamma(1, 20, 20/5)
+  
+  gen <- dnbinom(0:13, mu=meangen, size=size)
+  gen <- gen/sum(gen)
+  
+  recondata <- rr %>%
+    group_by(infection) %>%
+    summarize(
+      case=length(infection)
+    ) %>%
+    rename(
+      date=infection
+    ) %>%
+    filter(date >= as.Date("2020-01-19"), date <= as.Date("2020-03-10"))
+  
+  dd <- data.frame(
+    date=seq.Date(as.Date("2020-01-19"), as.Date("2020-03-10"), 1),
+    case=0
+  )
+  
+  dd$case[match(recondata$date, dd$date)] <- recondata$case
+  dd$case <- dd$case/(1-degree_censor$median)
+  
+  reconstruct_censor[[i]] <- dd
+  
+  IRt <- tail(dd$case, -14)/sapply(1:(length(dd$case)-14), function(x) sum(dd$case[x:(x+13)]*rev(gen)))
+  
+  Rt <- sapply(1:(length(IRt)-13), function(x) sum(IRt[x:(x+13)]*gen))
+  
+  R_t_gyeonggi_censor[[i]] <- data.frame(
+    date=seq.Date(as.Date("2020-01-19")+14, as.Date("2020-03-10"), 1),
+    IRt=IRt,
+    Rt=c(Rt, rep(NA, 13))
+  )
+}
+
+save("R_t_gyeonggi_censor", "reconstruct_censor", file="R_t_gyeonggi_censor.rda")

### 2020-03-18 R_t_seoul_censor.R

@@ -1,0 +1,50 @@
+library(dplyr)
+load("reconstruct_time_series_seoul.rda")
+load("degree_censor.rda")
+
+reconstruct_censor <- R_t_seoul_censor <- vector('list', length(reconstruct_time_series_seoul))
+
+set.seed(101)
+for (i in 1:length(reconstruct_time_series_seoul)) {
+  rr <- reconstruct_time_series_seoul[[i]]
+  
+  # pgamma(7, 62, 62/5) - pgamma(4, 62, 62/5) approx .95
+  
+  meangen <- rgamma(1, 62, 62/5)
+  size <- rgamma(1, 20, 20/5)
+  
+  gen <- dnbinom(0:13, mu=meangen, size=size)
+  gen <- gen/sum(gen)
+  
+  recondata <- rr %>%
+    group_by(infection) %>%
+    summarize(
+      case=length(infection)
+    ) %>%
+    rename(
+      date=infection
+    ) %>%
+    filter(date >= as.Date("2020-01-19"), date <= as.Date("2020-03-10"))
+  
+  dd <- data.frame(
+    date=seq.Date(as.Date("2020-01-19"), as.Date("2020-03-10"), 1),
+    case=0
+  )
+  
+  dd$case[match(recondata$date, dd$date)] <- recondata$case
+  dd$case <- dd$case/(1-degree_censor$median)
+  
+  reconstruct_censor[[i]] <- dd
+  
+  IRt <- tail(dd$case, -14)/sapply(1:(length(dd$case)-14), function(x) sum(dd$case[x:(x+13)]*rev(gen)))
+  
+  Rt <- sapply(1:(length(IRt)-13), function(x) sum(IRt[x:(x+13)]*gen))
+  
+  R_t_seoul_censor[[i]] <- data.frame(
+    date=seq.Date(as.Date("2020-01-19")+14, as.Date("2020-03-10"), 1),
+    IRt=IRt,
+    Rt=c(Rt, rep(NA, 13))
+  )
+}
+
+save("R_t_seoul_censor", "reconstruct_censor", file="R_t_seoul_censor.rda")

### 2020-03-18 R_t_seoul_censor_raw.R

@@ -1,0 +1,50 @@
+library(dplyr)
+load("reconstruct_time_series_seoul_raw.rda")
+load("degree_censor.rda")
+
+reconstruct_censor_raw <- R_t_seoul_censor_raw <- vector('list', length(reconstruct_time_series_seoul_raw))
+
+set.seed(101)
+for (i in 1:length(reconstruct_time_series_seoul_raw)) {
+  rr <- reconstruct_time_series_seoul_raw[[i]]
+  
+  # pgamma(7, 62, 62/5) - pgamma(4, 62, 62/5) approx .95
+  
+  meangen <- rgamma(1, 62, 62/5)
+  size <- rgamma(1, 20, 20/5)
+  
+  gen <- dnbinom(0:13, mu=meangen, size=size)
+  gen <- gen/sum(gen)
+  
+  recondata <- rr %>%
+    group_by(infection) %>%
+    summarize(
+      case=length(infection)
+    ) %>%
+    rename(
+      date=infection
+    ) %>%
+    filter(date >= as.Date("2020-01-19"), date <= as.Date("2020-03-10"))
+  
+  dd <- data.frame(
+    date=seq.Date(as.Date("2020-01-19"), as.Date("2020-03-10"), 1),
+    case=0
+  )
+  
+  dd$case[match(recondata$date, dd$date)] <- recondata$case
+  dd$case <- dd$case/(1-degree_censor$median)
+  
+  reconstruct_censor_raw[[i]] <- dd
+  
+  IRt <- tail(dd$case, -14)/sapply(1:(length(dd$case)-14), function(x) sum(dd$case[x:(x+13)]*rev(gen)))
+  
+  Rt <- sapply(1:(length(IRt)-13), function(x) sum(IRt[x:(x+13)]*gen))
+  
+  R_t_seoul_censor_raw[[i]] <- data.frame(
+    date=seq.Date(as.Date("2020-01-19")+14, as.Date("2020-03-10"), 1),
+    IRt=IRt,
+    Rt=c(Rt, rep(NA, 13))
+  )
+}
+
+save("R_t_seoul_censor_raw", "reconstruct_censor_raw", file="R_t_seoul_censor_raw.rda")

### 2020-03-18 degree_censor.R

@@ -1,0 +1,59 @@
+library(rstan)
+
+load("report_delay_censor.rda")
+load("pred_report_delay.rda")
+
+ee <- rstan::extract(report_delay_censor)
+
+shape <- ee$shape
+
+nsim <- 1000
+
+degree_censor <- vector('list', length(shape))
+
+set.seed(101)
+for (i in 1:length(shape)) {
+  if (i %% 100 == 0) print(i)
+  
+  meaninc <- rgamma(1, 145, 145/6.5)
+  ## pgamma(4.2, 25, 25/2.6) - pgamma(1.8, 25, 25/2.6)
+  sdinc <- rgamma(1, 25, 25/2.6)
+  
+  meandelay <- estmean[,i]
+  shapedelay <- ee$shape[i]
+  
+  dd <- data.frame(
+    date=seq.Date(from=as.Date("2020-01-10"), to=as.Date("2020-03-10"), by=1),
+    delay=meandelay
+  ) %>%
+    filter(date >= as.Date("2020-01-19"))
+  
+  censor <- sapply(1:nrow(dd), function(x) {
+    inc <- floor(rgamma(nsim, shape=shape, rate=shape/meaninc))
+    
+    symp <- as.Date("2020-01-19")-1+inc+x
+    
+    mu <- dd$delay[match(symp, dd$date)]
+    mu[is.na(mu)] <- tail(meandelay, 1)
+    
+    report <- symp + rnbinom(length(symp), mu=mu, size=shapedelay)
+    
+    mean(report >= as.Date("2020-03-16"))
+  })
+  
+  degree_censor[[i]] <- data.frame(
+    date=seq.Date(from=as.Date("2020-01-19"), to=as.Date("2020-03-10"), by=1),
+    censor=censor
+  )
+}
+
+degree_censor <- degree_censor %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(censor),
+    lwr=quantile(censor, 0.025),
+    upr=quantile(censor, 0.975)
+  )
+
+save("degree_censor", file="degree_censor.rda")

### 2020-03-18 figure_R_t_all.R

@@ -1,0 +1,59 @@
+library(dplyr)
+library(ggplot2); theme_set(theme_bw())
+
+load("R_t_chungnam_censor.rda")
+load("R_t_gyeongbuk_censor.rda")
+load("R_t_daegu_censor.rda")
+load("R_t_seoul_censor.rda")
+load("R_t_gyeonggi_censor.rda")
+
+R_t_all <- bind_rows(
+  R_t_daegu_censor %>%
+    bind_rows(.id="sim") %>%
+    mutate(
+      region="Daegu"
+    ),
+  R_t_seoul_censor %>%
+    bind_rows(.id="sim") %>%
+    mutate(
+      region="Seoul"
+    ),
+  R_t_gyeongbuk_censor %>%
+    bind_rows(.id="sim") %>%
+    mutate(
+      region="Gyeongsangbuk-do"
+    ),
+  R_t_gyeonggi_censor %>%
+    bind_rows(.id="sim") %>%
+    mutate(
+      region="Gyeonggi-do"
+    )
+) %>%
+  group_by(region, date) %>%
+  summarize(
+    median=median(IRt, na.rm=TRUE),
+    upr=quantile(IRt, 0.975, na.rm=TRUE),
+    lwr=quantile(IRt, 0.025, na.rm=TRUE)
+  ) %>%
+  ungroup %>%
+  mutate(
+    region=factor(region, levels=c("Daegu", "Seoul", "Gyeongsangbuk-do", "Gyeonggi-do"))
+  )
+
+g1 <- ggplot(R_t_all) +
+  geom_hline(yintercept=1, lty=2) +
+  geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.3) +
+  geom_line(aes(date, median)) +
+  scale_x_date("Date") +
+  scale_y_continuous("Effective reproduction number", limits=c(0, NA), expand=c(0, 0)) +
+  facet_wrap(~region, scale="free")+
+  theme(
+    panel.grid = element_blank(),
+    panel.border = element_blank(),
+    axis.line = element_line(),
+    legend.position = "none",
+    legend.title = element_blank(),
+    strip.background = element_blank()
+  )
+
+ggsave("figure_R_t_all.pdf", g1, width=8, height=6)

### 2020-03-18 figure_R_t_seoul.R

@@ -1,0 +1,137 @@
+library(ggplot2); theme_set(theme_bw())
+
+load("R_t_seoul_censor.rda")
+load("R_t_seoul_censor_detect.rda")
+load("R_t_seoul_censor_raw.rda")
+
+rt_censor <- R_t_seoul_censor %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(IRt, na.rm=TRUE),
+    lwr=quantile(IRt, 0.025, na.rm=TRUE),
+    upr=quantile(IRt, 0.975, na.rm=TRUE)
+  ) %>%
+  mutate(
+    group="adjusted for detection rate + number of tests"
+  )
+
+rt_censor_detect <- R_t_seoul_censor_detect %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(IRt, na.rm=TRUE),
+    lwr=quantile(IRt, 0.025, na.rm=TRUE),
+    upr=quantile(IRt, 0.975, na.rm=TRUE)
+  ) %>%
+  mutate(
+    group="adjusted for detection rate"
+  )
+
+rt_censor_raw <- R_t_seoul_censor_raw %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(IRt, na.rm=TRUE),
+    lwr=quantile(IRt, 0.025, na.rm=TRUE),
+    upr=quantile(IRt, 0.975, na.rm=TRUE)
+  ) %>%
+  mutate(
+    group="raw"
+  )
+
+rt_all <- bind_rows(rt_censor, rt_censor_detect, rt_censor_raw) %>%
+  mutate(
+    group=factor(group, level=c("raw",
+                                "adjusted for detection rate",
+                                "adjusted for detection rate + number of tests"))
+  )
+
+geo <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA", sheet=3)
+
+geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
+
+reconstruct_case <- reconstruct_censor %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(case),
+    lwr=quantile(case, 0.025),
+    upr=quantile(case, 0.975)
+  ) %>%
+  mutate(
[…]

### 2020-03-18 figure_adjusted_timeseries.R

@@ -1,0 +1,99 @@
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(ggplot2); theme_set(theme_bw())
+library(gridExtra)
+
+load("test_weight.rda")
+
+geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
+  mutate(
+    date_report=as.Date(date_report)
+  ) 
+
+geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
+
+daegu <- geo %>% 
+  group_by(date_report) %>%
+  summarize(
+    cases=sum(Daegu, na.rm=TRUE)
+  ) %>%
+  merge(test_weight) %>%
+  mutate(
+    cases2=cases*weight,
+    cases3=round(cases2*weight2),
+    cases2=round(cases2)
+  ) %>%
+  gather(
+    key, value,-date_report, -meanp, -test, -weight, -weight2, -group
+  ) %>%
+  mutate(
+    date_report=as.Date(date_report),
+    key=factor(key, levels=c("cases", "cases2", "cases3"),
+               labels=c("raw", "adjusted for detection rate", "adjusted for detection rate + number of tests"))
+  )
+
+seoul <- geo %>% 
+  group_by(date_report) %>%
+  summarize(
+    cases=sum(Seoul, na.rm=TRUE)
+  ) %>%
+  merge(test_weight) %>%
+  mutate(
+    cases2=cases*weight,
+    cases3=round(cases2*weight2),
+    cases2=round(cases2)
+  ) %>%
+  gather(
+    key, value,-date_report, -meanp, -test, -weight, -weight2, -group
+  ) %>%
+  mutate(
+    date_report=as.Date(date_report),
+    key=factor(key, levels=c("cases", "cases2", "cases3"),
+               labels=c("raw", "adjusted for detection rate", "adjusted for detection rate + number of tests"))
+  )
+
+g1 <- ggplot(daegu) +
+  geom_point(aes(date_report, value, col=key, shape=key)) +
+  geom_line(aes(date_report, value, col=key, lty=key)) +
+  scale_x_date("Date", expand=c(0, 0)) +
+  scale_y_continuous("Scaled number of reported cases", expand=c(0, 0)) +
+  scale_color_manual(values=c(1, 2, 4)) +
+  ggtitle("A. Daegu") +
+  theme(
+    panel.grid = element_blank(),
[…]

### 2020-03-18 figure_compare_R_t.R

@@ -1,0 +1,90 @@
+library(ggplot2); theme_set(theme_bw())
+library(gridExtra)
+
+load("R_t_daegu_censor.rda")
+
+rt_daegu <- R_t_daegu_censor %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(IRt, na.rm=TRUE),
+    lwr=quantile(IRt, 0.025, na.rm=TRUE),
+    upr=quantile(IRt, 0.975, na.rm=TRUE)
+  )
+
+case_daegu <- reconstruct_censor %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(case, na.rm=TRUE),
+    lwr=quantile(case, 0.025, na.rm=TRUE),
+    upr=quantile(case, 0.975, na.rm=TRUE)
+  )
+
+load("R_t_seoul_censor.rda")
+
+rt_seoul <- R_t_seoul_censor %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(IRt, na.rm=TRUE),
+    lwr=quantile(IRt, 0.025, na.rm=TRUE),
+    upr=quantile(IRt, 0.975, na.rm=TRUE)
+  )
+
+case_seoul <- reconstruct_censor %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(case, na.rm=TRUE),
+    lwr=quantile(case, 0.025, na.rm=TRUE),
+    upr=quantile(case, 0.975, na.rm=TRUE)
+  )
+
+g1 <- ggplot(rt_daegu) +
+  geom_point(data=case_daegu, aes(date, median/100), col=2) +
+  geom_errorbar(data=case_daegu, aes(date, ymin=lwr/100, ymax=upr/100), width=0, col=2) +
+  geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.3) +
+  geom_line(aes(date, median)) +
+  geom_hline(yintercept=1, lty=2) + 
+  scale_color_manual(values=c(1, 2, 4)) +
+  scale_fill_manual(values=c(1, 2, 4)) +
+  scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-19", "2020-03-10"))) +
+  scale_y_continuous("Effective reproduction number", limits=c(0, 8.5), expand=c(0, 0),
[…]

### 2020-03-18 figure_compare_report.R

@@ -1,0 +1,150 @@
+library(ggplot2); theme_set(theme_bw())
+library(gridExtra)
+
+geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
+  mutate(
+    date_report=as.Date(date_report)
+  ) 
+
+geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
+
+daegu <- geo %>% 
+  group_by(date_report) %>%
+  summarize(
+    cases=sum(Daegu, na.rm=TRUE)
+  )
+
+seoul <- geo %>% 
+  group_by(date_report) %>%
+  summarize(
+    cases=sum(Seoul, na.rm=TRUE)
+  )
+
+load("traffic_daegu.rda")
+load("traffic_seoul.rda")
+
+traffic_daegu1 <- select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2017-01-20")+3, date <= as.Date("2017-03-01")+3
+  )
+
+traffic_daegu2 <- select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2018-01-20")+2, date <= as.Date("2018-03-01")+2
+  )
+
+traffic_daegu3 <- select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2019-01-20")+1, date <= as.Date("2019-03-01")+1
+  )
+
+traffic_daegu4 <- select(ungroup(filter(traffic_daegu,year==2020)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2020-01-20"), date <= as.Date("2020-02-29")
+  )
+
+traffic_daegu_all <- data.frame(
+  date=seq.Date(as.Date("2020-01-20"), as.Date("2020-02-29"), 1),
+  traffic=c(traffic_daegu1$total,traffic_daegu2$total,traffic_daegu3$total,traffic_daegu4$total),
+  mean=((traffic_daegu1$total + traffic_daegu2$total + traffic_daegu3$total)/3),
[…]

### 2020-03-18 figure_report_delay.R

@@ -1,0 +1,183 @@
+library(brms)
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(lubridate)
+library(mgcv)
+library(rstan)
+library(gridExtra)
+library(ggplot2); theme_set(theme_bw())
+load("report_delay_censor.rda")
+
+covid_test <- read_xlsx("data/COVID19-Korea-2020-03-11.xlsx", na="NA", sheet=2) 
+
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+  mutate(
+    age=as.numeric(age),
+    date_onset=as.Date(date_onset)
+  )
+
+casenum <- covid_line$case[!is.na(covid_line$case)]
+
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=4, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_all <- rbind(covid_line, covid_seoul, covid_chungnam, covid_busan, covid_gyeongnam)
+
+covid_delay <- covid_all %>%
+  filter(!is.na(date_onset)) %>%
+  mutate(
+    delay=yday(date_confirm)-yday(date_onset),
+    day=yday(date_onset),
+    day=day-min(day),
+    absmax=yday(as.Date("2020-03-11"))-yday(date_onset)
+  ) %>%
[…]

### 2020-03-18 figure_traffic.R

@@ -1,0 +1,103 @@
+library(ggplot2); theme_set(theme_bw())
+library(gridExtra)
+
+load("traffic_daegu.rda")
+load("traffic_seoul.rda")
+
+traffic_daegu1 <- select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2017-02-02")+3, date <= as.Date("2017-03-01")+3
+  )
+
+traffic_daegu2 <- select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2018-02-02")+2, date <= as.Date("2018-03-01")+2
+  )
+
+traffic_daegu3 <- select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2019-02-02")+1, date <= as.Date("2019-03-01")+1
+  )
+
+traffic_daegu4 <- select(ungroup(filter(traffic_daegu,year==2020)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2020-02-02"), date <= as.Date("2020-02-29")
+  )
+
+traffic_daegu_all <- data.frame(
+  date=seq.Date(as.Date("2020-02-02"), as.Date("2020-02-29"), 1),
+  traffic=c(traffic_daegu1$total,traffic_daegu2$total,traffic_daegu3$total,traffic_daegu4$total),
+  mean=((traffic_daegu1$total + traffic_daegu2$total + traffic_daegu3$total)/3),
+  year=factor(rep(c(2017, 2018, 2019, 2020), each=28))
+)
+
+g1 <- ggplot(traffic_daegu_all) +
+  geom_line(aes(date, traffic, group=year, col=year)) +
+  geom_line(aes(date, mean)) +
+  scale_x_date("Date") +
+  scale_y_continuous("Daily metro traffic") +
+  scale_color_manual(values=c("gray", "gray", "gray", "red")) +
+  theme(
+    panel.grid = element_blank(),
+    panel.border = element_blank(),
+    axis.line = element_line(),
[…]

### 2020-03-18 infection_delay.R

@@ -1,0 +1,23 @@
+library(rstan)
+load("report_delay_censor.rda")
+load("pred_report_delay.rda")
+
+# pgamma(7.7, 175, 175/6.4) - pgamma(5.6, 175, 175/6.4)
+# pgamma(3.7, 35, 35/2.3) - pgamma(1.7, 35, 35/2.3)
+
+ee <- extract(report_delay_censor)
+
+for (i in 1:8000) {
+  mu_inc <- rgamma(1, 175, 175/6.4)
+  sd_inc <- rgamma(1, 35, 35/2.3)
+  
+  
+}
+
+plot(estmean[,1])
+
+ee$shape
+
+
+
+dnbinom()

### 2020-03-18 reconstruct_time_series_daegu_raw.R

@@ -1,0 +1,118 @@
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(lubridate)
+library(brms)
+library(rstan)
+
+load("report_delay_backward.rda")
+
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+  mutate(
+    age=as.numeric(age),
+    date_onset=as.Date(date_onset)
+  )
+
+casenum <- covid_line$case[!is.na(covid_line$case)]
+
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=4, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_all <- rbind(covid_line, covid_seoul, covid_chungnam, covid_busan, covid_gyeongnam)
+
+covid_delay <- covid_all %>%
+  filter(!is.na(date_onset), !is.na(date_confirm)) %>%
+  mutate(
+    delay=yday(date_confirm)-yday(date_onset),
+    day=yday(date_confirm),
+    day=day-min(day)
+  ) %>%
+  filter(yday(date_confirm) <= yday(as.Date("2020-03-11")))
+
+covid_delay_fake <- data.frame(
+  day=covid_delay$day[covid_delay$date_confirm==as.Date("2020-02-18")]:56,
+  delay=1
+)
+
[…]

### 2020-03-18 reconstruct_time_series_gyeongbuk.R

@@ -1,0 +1,124 @@
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(lubridate)
+library(brms)
+library(rstan)
+
+load("test_weight.rda")
+load("report_delay_backward.rda")
+
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+  mutate(
+    age=as.numeric(age),
+    date_onset=as.Date(date_onset)
+  )
+
+casenum <- covid_line$case[!is.na(covid_line$case)]
+
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=4, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_all <- rbind(covid_line, covid_seoul, covid_chungnam, covid_busan, covid_gyeongnam)
+
+covid_delay <- covid_all %>%
+  filter(!is.na(date_onset), !is.na(date_confirm)) %>%
+  mutate(
+    delay=yday(date_confirm)-yday(date_onset),
+    day=yday(date_confirm),
+    day=day-min(day)
+  ) %>%
+  filter(yday(date_confirm) <= yday(as.Date("2020-03-11")))
+
+covid_delay_fake <- data.frame(
+  day=covid_delay$day[covid_delay$date_confirm==as.Date("2020-01-20")]:56,
+  delay=1
+)
+
[…]

### 2020-03-18 reconstruct_time_series_gyeonggi.R

@@ -1,0 +1,124 @@
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(lubridate)
+library(brms)
+library(rstan)
+
+load("test_weight.rda")
+load("report_delay_backward.rda")
+
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+  mutate(
+    age=as.numeric(age),
+    date_onset=as.Date(date_onset)
+  )
+
+casenum <- covid_line$case[!is.na(covid_line$case)]
+
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=4, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_all <- rbind(covid_line, covid_seoul, covid_chungnam, covid_busan, covid_gyeongnam)
+
+covid_delay <- covid_all %>%
+  filter(!is.na(date_onset), !is.na(date_confirm)) %>%
+  mutate(
+    delay=yday(date_confirm)-yday(date_onset),
+    day=yday(date_confirm),
+    day=day-min(day)
+  ) %>%
+  filter(yday(date_confirm) <= yday(as.Date("2020-03-11")))
+
+covid_delay_fake <- data.frame(
+  day=covid_delay$day[covid_delay$date_confirm==as.Date("2020-01-20")]:56,
+  delay=1
+)
+
[…]

### 2020-03-18 reconstruct_time_series_seoul.R

@@ -1,0 +1,124 @@
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(lubridate)
+library(brms)
+library(rstan)
+
+load("test_weight.rda")
+load("report_delay_backward.rda")
+
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+  mutate(
+    age=as.numeric(age),
+    date_onset=as.Date(date_onset)
+  )
+
+casenum <- covid_line$case[!is.na(covid_line$case)]
+
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=4, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_all <- rbind(covid_line, covid_seoul, covid_chungnam, covid_busan, covid_gyeongnam)
+
+covid_delay <- covid_all %>%
+  filter(!is.na(date_onset), !is.na(date_confirm)) %>%
+  mutate(
+    delay=yday(date_confirm)-yday(date_onset),
+    day=yday(date_confirm),
+    day=day-min(day)
+  ) %>%
+  filter(yday(date_confirm) <= yday(as.Date("2020-03-11")))
+
+covid_delay_fake <- data.frame(
+  day=covid_delay$day[covid_delay$date_confirm==as.Date("2020-01-20")]:56,
+  delay=1
+)
+
[…]

### 2020-03-18 reconstruct_time_series_seoul_raw.R

@@ -1,0 +1,118 @@
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(lubridate)
+library(brms)
+library(rstan)
+
+load("report_delay_backward.rda")
+
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+  mutate(
+    age=as.numeric(age),
+    date_onset=as.Date(date_onset)
+  )
+
+casenum <- covid_line$case[!is.na(covid_line$case)]
+
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=4, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_all <- rbind(covid_line, covid_seoul, covid_chungnam, covid_busan, covid_gyeongnam)
+
+covid_delay <- covid_all %>%
+  filter(!is.na(date_onset), !is.na(date_confirm)) %>%
+  mutate(
+    delay=yday(date_confirm)-yday(date_onset),
+    day=yday(date_confirm),
+    day=day-min(day)
+  ) %>%
+  filter(yday(date_confirm) <= yday(as.Date("2020-03-11")))
+
+covid_delay_fake <- data.frame(
+  day=covid_delay$day[covid_delay$date_confirm==as.Date("2020-01-20")]:56,
+  delay=1
+)
+
[…]

### 2020-03-18 report_delay_backward.R

@@ -1,0 +1,66 @@
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(lubridate)
+library(brms)
+library(rstan)
+
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+  mutate(
+    age=as.numeric(age),
+    date_onset=as.Date(date_onset)
+  )
+
+casenum <- covid_line$case[!is.na(covid_line$case)]
+
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=4, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_all <- rbind(covid_line, covid_seoul, covid_chungnam, covid_busan, covid_gyeongnam)
+
+covid_delay <- covid_all %>%
+  filter(!is.na(date_onset), !is.na(date_confirm)) %>%
+  mutate(
+    delay=yday(date_confirm)-yday(date_onset),
+    day=yday(date_confirm),
+    day=day-min(day)
+  ) %>%
+  filter(yday(date_confirm) <= yday(as.Date("2020-03-11")))
+
+report_delay_backward <- brm(delay~s(day),
+                             data=covid_delay,
+                             family=negbinomial,
+                             prior=prior(normal(0, 2), class=b) +
+                               prior(normal(0, 2), class=Intercept),
+                             seed=102, thin=5)
+
+save("report_delay_backward", file="report_delay_backward.rda")

### 2020-03-18 report_delay_censor.R

@@ -1,0 +1,73 @@
+library(dplyr)
+library(tidyr)
+library(readxl)
+library(lubridate)
+library(brms)
+library(rstan)
+
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+  mutate(
+    age=as.numeric(age),
+    date_onset=as.Date(date_onset)
+  )
+
+casenum <- covid_line$case[!is.na(covid_line$case)]
+
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=4, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  filter(
+    !(case %in% casenum)
+  )
+
+covid_all <- rbind(covid_line, covid_seoul, covid_chungnam, covid_busan, covid_gyeongnam)
+
+covid_delay <- covid_all %>%
+  filter(!is.na(date_onset), !is.na(date_confirm)) %>%
+  mutate(
+    delay=yday(date_confirm)-yday(date_onset),
+    day=yday(date_onset),
+    day=day-min(day),
+    absmax=yday(as.Date("2020-03-11"))-yday(date_onset)
+  ) %>%
+  filter(yday(date_confirm) <= yday(as.Date("2020-03-11")))
+
+standata <- make_standata(delay~s(day),
+              data=covid_delay,
+              family=negbinomial)
+
+standata$absmax <- covid_delay$absmax
+
+smodel <- stan_model(file="report_delay_censor.stan")
+
+report_delay_censor <- sampling(smodel, data=standata,
[…]

### 2020-03-18 test_weight.R

@@ -1,0 +1,47 @@
+library(readxl)
+
+tt <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=2)
+
+tt$negative[7] <- (tt$negative[6]+tt$negative[8])/2
+tt$negative[13] <- (tt$negative[12]+tt$negative[14])/2
+
+test <- tt %>%
+  mutate(
+    date_report=as.Date(date_report),
+    time_report=hour(time_report)
+  ) %>%
+  filter(
+    is.na(time_report) | time_report != 16
+  ) %>%
+  mutate(
+    test=positive+negative,
+    test=diff(c(0, test)),
+    positive=diff(c(0, positive))
+  ) %>%
+  filter(
+    date_report >= as.Date("2020-01-20")
+  ) %>%
+  mutate(
+    group=1,
+    group=ifelse(date_report >= as.Date("2020-01-28"), 2, group),
+    group=ifelse(date_report >= as.Date("2020-02-07"), 3, group),
+    group=ifelse(date_report >= as.Date("2020-02-20"), 4, group),
+    group=ifelse(date_report >= as.Date("2020-03-02"), 5, group),
+    group=factor(group)
+  )
+
+test_weight <- test %>%
+  group_by(group) %>%
+  summarize(
+    meanp=sum(positive)/sum(test)
+  ) %>%
+  mutate(
+    weight=meanp/mean(meanp)
+  ) %>%
+  merge(select(test, date_report, group, test)) %>%
+  group_by(group) %>%
+  mutate(
+    weight2=mean(test)/test
+  )
+
+save("test_weight", file="test_weight.rda")

### 2020-03-18 traffic_daegu.R

@@ -1,0 +1,52 @@
+library(dplyr)
+library(tidyr)
+library(lubridate)
+library(ggplot2); theme_set(theme_bw())
+
+summfun <- function(x) {
+  x %>%
+    filter(승하=="승차") %>%
+    group_by(월, 일) %>%
+    summarize(
+      total=sum(일계)
+    )
+}
+
+summfun2 <- function(x) {
+  x %>%
+    filter(승하차=="승차") %>%
+    group_by(월, 일) %>%
+    summarize(
+      total=sum(일계)
+    )
+}
+
+dd2017 <- read.csv("daegu_traffic/daegu_metro_daily_onoff_20171231.csv", stringsAsFactors=FALSE,
+                   fileEncoding="EUC-KR") %>% 
+  summfun %>%
+  mutate(year=2017)
+
+dd2018 <- read.csv("daegu_traffic/daegu_metro_daily_onoff_20181231.csv", stringsAsFactors=FALSE,
+                   fileEncoding="EUC-KR") %>% 
+  summfun2 %>%
+  mutate(year=2018)
+
+dd2019 <- read.csv("daegu_traffic/daegu_metro_daily_onoff_20191231.csv", stringsAsFactors=FALSE,
+                   fileEncoding="EUC-KR") %>% 
+  summfun %>%
+  mutate(year=2019)
+
+dd2020 <- read.csv("daegu_traffic/daegu_metro_daily_onoff_20200229.csv", stringsAsFactors=FALSE,
+               fileEncoding="EUC-KR") %>% 
+  summfun %>%
+  mutate(year=2020)
+
+traffic_daegu <- ddall <- bind_rows(dd2017, dd2018, dd2019, dd2020) %>%
+  mutate(
+    date=as.Date(paste0(year, "-", 월, "-", 일)),
+    day=yday(date),
+    year=factor(year),
+    location="Daegu"
+  )
+
+save("traffic_daegu", file="traffic_daegu.rda")

### 2020-03-18 traffic_seoul.R

@@ -1,0 +1,175 @@
+library(dplyr)
+library(tidyr)
+library(lubridate)
+library(ggplot2); theme_set(theme_bw())
+
+dataname <- c("사용일자", "노선명", "역ID", "역명", "승차총승객수", "하차총승객수", "등록일자")
+
+ss2020 <- read.csv("seoul_subway/서울시 지하철호선별 역별 승하차 인원 정보.csv", fileEncoding="EUC-KR") 
+
+ss2020a <- ss2020 %>%
+  mutate(
+    사용일자=gsub("2020", "2020-", 사용일자),
+    사용일자=gsub("2020-03", "2020-03-", 사용일자),
+    date=as.Date(gsub("2020-02", "2020-02-", 사용일자))
+  ) %>%
+  group_by(date) %>%
+  summarize(
+    total=sum(승차총승객수),
+    year=2020
+  )
+
+ss2020b <- read.csv("seoul_subway/CARD_SUBWAY_MONTH_202002.csv", stringsAsFactors=FALSE, fileEncoding="latin1") %>%
+  setNames(dataname) %>%
+  mutate(
+    사용일자=gsub("2020", "2020-", 사용일자),
+    date=as.Date(gsub("2020-02", "2020-02-", 사용일자))
+  ) %>%
+  group_by(date) %>%
+  summarize(
+    total=sum(승차총승객수),
+    year=2020
+  ) %>%
+  filter(date=="2020-02-01")
+
+ss2020c <- read.csv("seoul_subway/CARD_SUBWAY_MONTH_202001.csv", stringsAsFactors=FALSE, fileEncoding="latin1",
+                    row.names = NULL) %>%
+  setNames(c(dataname, "empty")) %>%
+  mutate(
+    사용일자=gsub("2020", "2020-", 사용일자),
+    date=as.Date(gsub("2020-01", "2020-01-", 사용일자))
+  ) %>%
+  group_by(date) %>%
+  summarize(
+    total=sum(승차총승객수),
+    year=2020
+  )
+
+ss2019a <- read.csv("seoul_subway/CARD_SUBWAY_MONTH_201901.csv", stringsAsFactors=FALSE, fileEncoding="latin1",
+                    row.names = NULL) %>%
+  setNames(c(dataname)) %>%
+  mutate(
+    사용일자=gsub("2019", "2019-", 사용일자),
+    date=as.Date(gsub("2019-01", "2019-01-", 사용일자))
+  ) %>%
[…]

### 2020-03-21 korea.tex
update

@@
 	\textbf\newline{
-		Estimating the effects of social distancing on preventing the spread of coronavirus disease 2019 (COVID-19) in South Korea
+		Potential roles of social distancing in preventing the spread of coronavirus disease 2019 (COVID-19) in South Korea
 	}
@@
 
-Here, we use traffic data to compare and contrast the potential role of social distancing in preventing the spread of COVID-19 in two geographically separated major cities in South Korea.
+Since its appearance in December 2019, the novel coronavirus disease (COVID-19) has spread internationally, including to South Korea.
+The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China.
+The epidemic took off in mid-February when COVID-19 began to spread within a church from a city of Daegu --- as of March 18, 2020, 8,413 cases have been confirmed, of which 60\% are linked to the church.
+Despite the lack of draconian measures such as in China, the epidemic has been subsiding in South Korea.
+Its success in controlling the spread has been widely attributed to its intensive testing.
+Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by comparing epidemics in two geographically separated major cities in South Korea.
 
@@
 
-Here, we analyze epidemiological data, collected bewteen January 20--March 16, 2020, describing the COVID-19 outbreak in Korea.
-Daily number of reported cases in each geographic region was transcribed from press releases by Korea Centers for Disease Control and Prevention (KCDC).
+We analyze epidemiological data, collected bewteen January 20--March 16, 2020, describing the COVID-19 outbreak in Korea.
+Daily number of reported cases in each geographic region was transcribed from press releases by \cite{kcdc}.
 Partial line lists were transcribed from press releases and reports from KCDC and various local and provincial government websites.
-All data and original reports are stored in a publicly available GitHub repoository: https://github.com/parksw3/COVID19-Korea.
+All data, sources, and original reports are stored in a publicly available GitHub repoository: \url{https://github.com/parksw3/COVID19-Korea}.
 
@@
 \textbf{Comparison of epidemiological and traffic data from Daegu and Seoul.}
-Solid lines represent the daily metro traffic volume in 2020 (red) and in 2017--2019 (gray).
-Black lines represent the average daily metro traffic volume in previous years.
+Red lines represent the daily metro traffic volume in 2020.
+Black lines represent the average daily metro traffic volume in previous years, 2017--2019.
+Daily traffic from previous years have been shifted by 1--3 days to align day of the week.
 }
@@
 
-We focus our analysis on two cities in which the biggest number of COVID-19 cases have been reported: Daegu and Seoul.
+We compare epidemiological dynamics of COVID-19 from two cities in which the biggest number of COVID-19 cases have been reported: Daegu and Seoul.
+% population size: 2.4 and 9.7 million \citep{kosis}.
 Bewteen January 20--March 16, 2020, 6,083 cases were reported from Daegu and 248 from Seoul.
-Need one or two sentences about epidemic.
+Unlike the epidemic in Daegu, which is characterized by a single, large peak followed by a persistent decrease, the epidemic in Seoul consists of several small outbreaks (\fref{data}).
 
-Daily metro traffic from Daegu and Seoul were obtained from their corresponding government websites.
-Need one or two sentences about metro. 
+Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from their corresponding government websites.
+
+\fref{data}
 About 80\% decrease in Daegu and 50\% decrease in Seoul after the identification of the church-related case.
@@
 
-We first estimate the time-varying effective reproduction number $R_t$ for both epidemics.
+We estimate the time-varying effective reproduction number $R_t$ for both epidemics.
 We account for changes in case definition by scaling the number of reported cases by the relative detection rate (i.e., a proportion of cases tested positive based a case definition divided by the mean detection rate among case definitions);
@@
 \caption{
-\textbf{Relationship between daily traffic volume and effective reproduction number.}
+\textbf{Other effects.}
 }
@@
 
+\bibliography{korea}
+
 \section{Supplementary Materials}
@@
 }
-\label{fig:eff}
 \end{figure}

### 2020-03-21 R_t_daegu_censor_detect.R

@@ -13,7 +13,7 @@ for (i in 1:length(reconstruct_time_series_daegu_detect)) {
   meangen <- rgamma(1, 62, 62/5)
   size <- rgamma(1, 20, 20/5)
   
-  gen <- dnbinom(0:13, mu=meangen, size=size)
+  gen <- dnbinom(1:14, mu=meangen, size=size)
   gen <- gen/sum(gen)
   
   recondata <- rr %>%

### 2020-03-21 R_t_gyeongbuk_censor_detect.R

@@ -1,12 +1,12 @@
 library(dplyr)
-load("reconstruct_time_series_gyeongbuk.rda")
+load("reconstruct_time_series_gyeongbuk_detect.rda")
 load("degree_censor.rda")
 
-reconstruct_censor <- R_t_gyeongbuk_censor <- vector('list', length(reconstruct_time_series_gyeongbuk))
+reconstruct_censor_detect <- R_t_gyeongbuk_censor_detect <- vector('list', length(reconstruct_time_series_gyeongbuk_detect))
 
 set.seed(101)
-for (i in 1:length(reconstruct_time_series_gyeongbuk)) {
-  rr <- reconstruct_time_series_gyeongbuk[[i]]
+for (i in 1:length(reconstruct_time_series_gyeongbuk_detect)) {
+  rr <- reconstruct_time_series_gyeongbuk_detect[[i]]
   
   # pgamma(7, 62, 62/5) - pgamma(4, 62, 62/5) approx .95
   
@@ -34,17 +34,17 @@ for (i in 1:length(reconstruct_time_series_gyeongbuk)) {
   dd$case[match(recondata$date, dd$date)] <- recondata$case
   dd$case <- dd$case/(1-degree_censor$median)
   
-  reconstruct_censor[[i]] <- dd
+  reconstruct_censor_detect[[i]] <- dd
   
   IRt <- tail(dd$case, -14)/sapply(1:(length(dd$case)-14), function(x) sum(dd$case[x:(x+13)]*rev(gen)))
   
   Rt <- sapply(1:(length(IRt)-13), function(x) sum(IRt[x:(x+13)]*gen))
   
-  R_t_gyeongbuk_censor[[i]] <- data.frame(
+  R_t_gyeongbuk_censor_detect[[i]] <- data.frame(
     date=seq.Date(as.Date("2020-01-19")+14, as.Date("2020-03-10"), 1),
     IRt=IRt,
     Rt=c(Rt, rep(NA, 13))
   )
 }
 
-save("R_t_gyeongbuk_censor", "reconstruct_censor", file="R_t_gyeongbuk_censor.rda")
+save("R_t_gyeongbuk_censor_detect", "reconstruct_censor_detect", file="R_t_gyeongbuk_censor_detect.rda")

### 2020-03-21 R_t_gyeongbuk_censor_raw.R

@@ -1,12 +1,12 @@
 library(dplyr)
-load("reconstruct_time_series_gyeongbuk.rda")
+load("reconstruct_time_series_gyeongbuk_raw.rda")
 load("degree_censor.rda")
 
-reconstruct_censor <- R_t_gyeongbuk_censor <- vector('list', length(reconstruct_time_series_gyeongbuk))
+reconstruct_censor_raw <- R_t_gyeongbuk_censor_raw <- vector('list', length(reconstruct_time_series_gyeongbuk_raw))
 
 set.seed(101)
-for (i in 1:length(reconstruct_time_series_gyeongbuk)) {
-  rr <- reconstruct_time_series_gyeongbuk[[i]]
+for (i in 1:length(reconstruct_time_series_gyeongbuk_raw)) {
+  rr <- reconstruct_time_series_gyeongbuk_raw[[i]]
   
   # pgamma(7, 62, 62/5) - pgamma(4, 62, 62/5) approx .95
   
@@ -34,17 +34,17 @@ for (i in 1:length(reconstruct_time_series_gyeongbuk)) {
   dd$case[match(recondata$date, dd$date)] <- recondata$case
   dd$case <- dd$case/(1-degree_censor$median)
   
-  reconstruct_censor[[i]] <- dd
+  reconstruct_censor_raw[[i]] <- dd
   
   IRt <- tail(dd$case, -14)/sapply(1:(length(dd$case)-14), function(x) sum(dd$case[x:(x+13)]*rev(gen)))
   
   Rt <- sapply(1:(length(IRt)-13), function(x) sum(IRt[x:(x+13)]*gen))
   
-  R_t_gyeongbuk_censor[[i]] <- data.frame(
+  R_t_gyeongbuk_censor_raw[[i]] <- data.frame(
     date=seq.Date(as.Date("2020-01-19")+14, as.Date("2020-03-10"), 1),
     IRt=IRt,
     Rt=c(Rt, rep(NA, 13))
   )
 }
 
-save("R_t_gyeongbuk_censor", "reconstruct_censor", file="R_t_gyeongbuk_censor.rda")
+save("R_t_gyeongbuk_censor_raw", "reconstruct_censor_raw", file="R_t_gyeongbuk_censor_raw.rda")

### 2020-03-21 R_t_gyeonggi_censor_detect.R

@@ -1,12 +1,12 @@
 library(dplyr)
-load("reconstruct_time_series_gyeongbuk.rda")
+load("reconstruct_time_series_gyeonggi_detect.rda")
 load("degree_censor.rda")
 
-reconstruct_censor <- R_t_gyeongbuk_censor <- vector('list', length(reconstruct_time_series_gyeongbuk))
+reconstruct_censor_detect <- R_t_gyeonggi_censor_detect <- vector('list', length(reconstruct_time_series_gyeonggi_detect))
 
 set.seed(101)
-for (i in 1:length(reconstruct_time_series_gyeongbuk)) {
-  rr <- reconstruct_time_series_gyeongbuk[[i]]
+for (i in 1:length(reconstruct_time_series_gyeonggi_detect)) {
+  rr <- reconstruct_time_series_gyeonggi_detect[[i]]
   
   # pgamma(7, 62, 62/5) - pgamma(4, 62, 62/5) approx .95
   
@@ -34,17 +34,17 @@ for (i in 1:length(reconstruct_time_series_gyeongbuk)) {
   dd$case[match(recondata$date, dd$date)] <- recondata$case
   dd$case <- dd$case/(1-degree_censor$median)
   
-  reconstruct_censor[[i]] <- dd
+  reconstruct_censor_detect[[i]] <- dd
   
   IRt <- tail(dd$case, -14)/sapply(1:(length(dd$case)-14), function(x) sum(dd$case[x:(x+13)]*rev(gen)))
   
   Rt <- sapply(1:(length(IRt)-13), function(x) sum(IRt[x:(x+13)]*gen))
   
-  R_t_gyeongbuk_censor[[i]] <- data.frame(
+  R_t_gyeonggi_censor_detect[[i]] <- data.frame(
     date=seq.Date(as.Date("2020-01-19")+14, as.Date("2020-03-10"), 1),
     IRt=IRt,
     Rt=c(Rt, rep(NA, 13))
   )
 }
 
-save("R_t_gyeongbuk_censor", "reconstruct_censor", file="R_t_gyeongbuk_censor.rda")
+save("R_t_gyeonggi_censor_detect", "reconstruct_censor_detect", file="R_t_gyeonggi_censor_detect.rda")

### 2020-03-21 R_t_seoul_censor.R

@@ -13,7 +13,7 @@ for (i in 1:length(reconstruct_time_series_seoul)) {
   meangen <- rgamma(1, 62, 62/5)
   size <- rgamma(1, 20, 20/5)
   
-  gen <- dnbinom(0:13, mu=meangen, size=size)
+  gen <- dnbinom(1:14, mu=meangen, size=size)
   gen <- gen/sum(gen)
   
   recondata <- rr %>%

### 2020-03-21 R_t_seoul_censor_detect.R

@@ -13,7 +13,7 @@ for (i in 1:length(reconstruct_time_series_seoul_detect)) {
   meangen <- rgamma(1, 62, 62/5)
   size <- rgamma(1, 20, 20/5)
   
-  gen <- dnbinom(0:13, mu=meangen, size=size)
+  gen <- dnbinom(1:14, mu=meangen, size=size)
   gen <- gen/sum(gen)
   
   recondata <- rr %>%

### 2020-03-21 R_t_seoul_censor_raw.R

@@ -13,7 +13,7 @@ for (i in 1:length(reconstruct_time_series_seoul_raw)) {
   meangen <- rgamma(1, 62, 62/5)
   size <- rgamma(1, 20, 20/5)
   
-  gen <- dnbinom(0:13, mu=meangen, size=size)
+  gen <- dnbinom(1:14, mu=meangen, size=size)
   gen <- gen/sum(gen)
   
   recondata <- rr %>%

### 2020-03-21 figure_R_t_daegu.R

@@ -1,3 +1,4 @@
+library(readxl)
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
 
@@ -122,7 +123,7 @@ g2 <- ggplot(rt_all) +
   scale_color_manual(values=c(1, 2, 4)) +
   scale_fill_manual(values=c(1, 2, 4)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-19", "2020-03-14"))) +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 8.5), expand=c(0, 0)) +
+  scale_y_continuous("Effective reproduction number", limits=c(0, 6.5), expand=c(0, 0)) +
   theme(
     panel.grid = element_blank(),
     panel.border = element_blank(),

### 2020-03-21 figure_R_t_gyeongbuk.R

@@ -1,10 +1,11 @@
 library(ggplot2); theme_set(theme_bw())
+library(gridExtra)
 
-load("R_t_seoul_censor.rda")
-load("R_t_seoul_censor_detect.rda")
-load("R_t_seoul_censor_raw.rda")
+load("R_t_gyeongbuk_censor.rda")
+load("R_t_gyeongbuk_censor_detect.rda")
+load("R_t_gyeongbuk_censor_raw.rda")
 
-rt_censor <- R_t_seoul_censor %>%
+rt_censor <- R_t_gyeongbuk_censor %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   summarize(
@@ -16,7 +17,7 @@ rt_censor <- R_t_seoul_censor %>%
     group="adjusted for detection rate + number of tests"
   )
 
-rt_censor_detect <- R_t_seoul_censor_detect %>%
+rt_censor_detect <- R_t_gyeongbuk_censor_detect %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   summarize(
@@ -28,7 +29,7 @@ rt_censor_detect <- R_t_seoul_censor_detect %>%
     group="adjusted for detection rate"
   )
 
-rt_censor_raw <- R_t_seoul_censor_raw %>%
+rt_censor_raw <- R_t_gyeongbuk_censor_raw %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   summarize(
@@ -95,15 +96,15 @@ reconstruct_case_all <- bind_rows(reconstruct_case, reconstruct_case_detect, rec
   )
 
 g1 <- ggplot(geo) +
-  geom_bar(aes(as.Date(date_report)-0.5, Seoul), stat="identity", alpha=0.5) +
+  geom_bar(aes(as.Date(date_report)-0.5, `Gyeongsangbuk-do`), stat="identity", alpha=0.5) +
+  geom_vline(xintercept=as.Date("2020-03-10"), lty=2) +
   geom_ribbon(data=reconstruct_case_all, aes(date, ymin=lwr, ymax=upr, col=group, fill=group, lty=group), alpha=0.3) +
   geom_line(data=reconstruct_case_all, aes(date, median, col=group, lty=group)) +
-  geom_vline(xintercept=as.Date("2020-03-10"), lty=2) +
   scale_color_manual(values=c(1, 2, 4)) +
   scale_fill_manual(values=c(1, 2, 4)) +
[…]

### 2020-03-21 figure_R_t_gyeonggi.R

@@ -1,10 +1,11 @@
 library(ggplot2); theme_set(theme_bw())
+library(gridExtra)
 
-load("R_t_seoul_censor.rda")
-load("R_t_seoul_censor_detect.rda")
-load("R_t_seoul_censor_raw.rda")
+load("R_t_gyeonggi_censor.rda")
+load("R_t_gyeonggi_censor_detect.rda")
+load("R_t_gyeonggi_censor_raw.rda")
 
-rt_censor <- R_t_seoul_censor %>%
+rt_censor <- R_t_gyeonggi_censor %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   summarize(
@@ -16,7 +17,7 @@ rt_censor <- R_t_seoul_censor %>%
     group="adjusted for detection rate + number of tests"
   )
 
-rt_censor_detect <- R_t_seoul_censor_detect %>%
+rt_censor_detect <- R_t_gyeonggi_censor_detect %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   summarize(
@@ -28,7 +29,7 @@ rt_censor_detect <- R_t_seoul_censor_detect %>%
     group="adjusted for detection rate"
   )
 
-rt_censor_raw <- R_t_seoul_censor_raw %>%
+rt_censor_raw <- R_t_gyeonggi_censor_raw %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   summarize(
@@ -95,15 +96,15 @@ reconstruct_case_all <- bind_rows(reconstruct_case, reconstruct_case_detect, rec
   )
 
 g1 <- ggplot(geo) +
-  geom_bar(aes(as.Date(date_report)-0.5, Seoul), stat="identity", alpha=0.5) +
+  geom_bar(aes(as.Date(date_report)-0.5, `Gyeonggi-do`), stat="identity", alpha=0.5) +
+  geom_vline(xintercept=as.Date("2020-03-10"), lty=2) +
   geom_ribbon(data=reconstruct_case_all, aes(date, ymin=lwr, ymax=upr, col=group, fill=group, lty=group), alpha=0.3) +
   geom_line(data=reconstruct_case_all, aes(date, median, col=group, lty=group)) +
-  geom_vline(xintercept=as.Date("2020-03-10"), lty=2) +
   scale_color_manual(values=c(1, 2, 4)) +
   scale_fill_manual(values=c(1, 2, 4)) +
[…]

### 2020-03-21 figure_compare_R_t.R

@@ -2,6 +2,41 @@ library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
 
 load("R_t_daegu_censor.rda")
+load("R_t_seoul_censor.rda")
+load("traffic_daegu.rda")
+load("traffic_seoul.rda")
+
+traffic_daegu1 <- select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2017-01-20")+3, date <= as.Date("2017-03-01")+3
+  )
+
+traffic_daegu2 <- select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2018-01-20")+2, date <= as.Date("2018-03-01")+2
+  )
+
+traffic_daegu3 <- select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2019-01-20")+1, date <= as.Date("2019-03-01")+1
+  )
+
+traffic_daegu4 <- select(ungroup(filter(traffic_daegu,year==2020)), -월, -일) %>%
+  mutate(
+    wday=wday(date)
+  ) %>%
+  filter(
+    date >= as.Date("2020-01-20"), date <= as.Date("2020-02-29")
+  )
 
 rt_daegu <- R_t_daegu_censor %>%
   bind_rows(.id="sim") %>%
@@ -9,19 +44,46 @@ rt_daegu <- R_t_daegu_censor %>%
   summarize(
     median=median(IRt, na.rm=TRUE),
     lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE)
+    upr=quantile(IRt, 0.975, na.rm=TRUE),
+    location="Daegu"
   )
 
-case_daegu <- reconstruct_censor %>%
-  bind_rows(.id="sim") %>%
-  group_by(date) %>%
-  summarize(
-    median=median(case, na.rm=TRUE),
-    lwr=quantile(case, 0.025, na.rm=TRUE),
-    upr=quantile(case, 0.975, na.rm=TRUE)
+daegu_traffic <- data.frame(
+  date=seq.Date(as.Date("2020-01-20"), as.Date("2020-02-29"), 1),
[…]

### 2020-03-21 figure_compare_report.R

@@ -1,5 +1,10 @@
+library(dplyr)
+library(tidyr)
+library(readxl)
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
+library(lubridate)
+library(raster)
 
 geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
   mutate(
@@ -101,15 +106,14 @@ traffic_seoul_all <- data.frame(
   year=factor(rep(c(2017, 2018, 2019, 2020), each=51))
 )
 
-g1 <- ggplot(traffic_daegu_all) +
+g1 <- ggplot(filter(traffic_daegu_all, year==2020)) +
   geom_bar(data=daegu, aes(date_report, cases), stat="identity", alpha=0.5) + 
-  geom_line(aes(date, traffic/3e2, group=year, col=year, lwd=factor(year))) +
   geom_line(aes(date, mean/3e2), lwd=1) +
+  geom_line(aes(date, traffic/3e2, group=year, col=year), lwd=1) +
   scale_x_date("Date", expand=c(0, 0)) +
   scale_y_continuous("Daily number of reported cases",
                      sec.axis = sec_axis(~ .*3e2, name = "Daily traffic volume, 2020"), expand=c(0, 0)) +
-  scale_color_manual(values=c("gray", "gray", "gray", "red")) +
-  scale_size_manual(values=c(0.7, 0.7, 0.7, 1)) +
+  scale_color_manual(values=c("red")) +
   ggtitle("A. Daegu") +
   theme(
     panel.grid = element_blank(),
@@ -123,15 +127,14 @@ g1 <- ggplot(traffic_daegu_all) +
     axis.text.y.right = element_text(color="red")
   )
 
-g2 <- ggplot(traffic_seoul_all) +
+g2 <- ggplot(filter(traffic_seoul_all, year==2020)) +
   geom_bar(data=seoul, aes(date_report, cases), stat="identity", alpha=0.5) + 
-  geom_line(aes(date, traffic/1e5, group=year, col=year, lwd=factor(year))) +
   geom_line(aes(date, mean/1e5), lwd=1) +
+  geom_line(aes(date, traffic/1e5, group=year, col=year), lwd=1) +
   scale_x_date("Date", expand=c(0, 0)) +
   scale_y_continuous("Daily number of reported cases",
[…]

### 2020-03-21 reconstruct_time_series_daegu_detect.R

@@ -79,7 +79,7 @@ estmean <- sapply(1:length(ee$b_Intercept), function(i) {
 geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
   mutate(
     date_report=as.Date(date_report)
-  ) 
+  )
 
 geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
 

### 2020-03-21 reconstruct_time_series_gyeongbuk_detect.R

@@ -5,6 +5,7 @@ library(lubridate)
 library(brms)
 library(rstan)
 
+load("test_weight.rda")
 load("report_delay_backward.rda")
 
 covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
@@ -78,18 +79,22 @@ estmean <- sapply(1:length(ee$b_Intercept), function(i) {
 geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
   mutate(
     date_report=as.Date(date_report)
-  ) 
+  )
 
 geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
 
-seoul <- geo %>% group_by(date_report) %>%
+gyeongbuk <- geo %>% group_by(date_report) %>%
   summarize(
-    cases=sum(Seoul, na.rm=TRUE)
+    cases=sum(`Gyeongsangbuk-do`, na.rm=TRUE)
+  ) %>%
+  merge(test_weight) %>%
+  mutate(
+    cases2=round(cases*weight)
   )
 
-cases <- filter(seoul, date_report <= as.Date("2020-03-16"))$cases
+cases <- filter(gyeongbuk, date_report <= as.Date("2020-03-16"))$cases2
 
-reconstruct_time_series_seoul_raw <- vector('list', length(ee$b_Intercept))
+reconstruct_time_series_gyeongbuk_detect <- vector('list', length(ee$b_Intercept))
 
 set.seed(101)
 for (i in 1:length(ee$b_Intercept)) {
@@ -105,14 +110,14 @@ for (i in 1:length(ee$b_Intercept)) {
   meandelay <- estmean[,i]
   shape <- ee$shape[i]
   
-  reconstruct_time_series_seoul_raw[[i]] <- sapply(1:length(cases), function(x) {
+  reconstruct_time_series_gyeongbuk_detect[[i]] <- sapply(1:length(cases), function(x) {
     data_frame(
       confirm=as.Date("2020-01-20")+x-1,
-      onset=confirm-rnbinom(cases[x], mu=meandelay[x], size=shape),
-      infection=onset-floor(rgamma(cases[x], shape=shape, rate=shape/meaninc))
+      onset=confirm-rnbinom(round(cases[x]), mu=meandelay[x], size=shape),
[…]

### 2020-03-21 reconstruct_time_series_gyeongbuk_raw.R

@@ -5,6 +5,7 @@ library(lubridate)
 library(brms)
 library(rstan)
 
+load("test_weight.rda")
 load("report_delay_backward.rda")
 
 covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
@@ -78,18 +79,18 @@ estmean <- sapply(1:length(ee$b_Intercept), function(i) {
 geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
   mutate(
     date_report=as.Date(date_report)
-  ) 
+  )
 
 geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
 
-seoul <- geo %>% group_by(date_report) %>%
+gyeongbuk <- geo %>% group_by(date_report) %>%
   summarize(
-    cases=sum(Seoul, na.rm=TRUE)
+    cases=sum(`Gyeongsangbuk-do`, na.rm=TRUE)
   )
 
-cases <- filter(seoul, date_report <= as.Date("2020-03-16"))$cases
+cases <- filter(gyeongbuk, date_report <= as.Date("2020-03-16"))$cases
 
-reconstruct_time_series_seoul_raw <- vector('list', length(ee$b_Intercept))
+reconstruct_time_series_gyeongbuk_raw <- vector('list', length(ee$b_Intercept))
 
 set.seed(101)
 for (i in 1:length(ee$b_Intercept)) {
@@ -105,14 +106,14 @@ for (i in 1:length(ee$b_Intercept)) {
   meandelay <- estmean[,i]
   shape <- ee$shape[i]
   
-  reconstruct_time_series_seoul_raw[[i]] <- sapply(1:length(cases), function(x) {
+  reconstruct_time_series_gyeongbuk_raw[[i]] <- sapply(1:length(cases), function(x) {
     data_frame(
       confirm=as.Date("2020-01-20")+x-1,
-      onset=confirm-rnbinom(cases[x], mu=meandelay[x], size=shape),
-      infection=onset-floor(rgamma(cases[x], shape=shape, rate=shape/meaninc))
+      onset=confirm-rnbinom(round(cases[x]), mu=meandelay[x], size=shape),
+      infection=onset-floor(rgamma(round(cases[x]), shape=shape, rate=shape/meaninc))
     )
[…]

### 2020-03-21 reconstruct_time_series_gyeonggi_raw.R

@@ -5,6 +5,7 @@ library(lubridate)
 library(brms)
 library(rstan)
 
+load("test_weight.rda")
 load("report_delay_backward.rda")
 
 covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
@@ -82,14 +83,14 @@ geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
 
 geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
 
-seoul <- geo %>% group_by(date_report) %>%
+gyeonggi <- geo %>% group_by(date_report) %>%
   summarize(
-    cases=sum(Seoul, na.rm=TRUE)
+    cases=sum(`Gyeonggi-do`, na.rm=TRUE)
   )
 
-cases <- filter(seoul, date_report <= as.Date("2020-03-16"))$cases
+cases <- filter(gyeonggi, date_report <= as.Date("2020-03-16"))$cases
 
-reconstruct_time_series_seoul_raw <- vector('list', length(ee$b_Intercept))
+reconstruct_time_series_gyeonggi_raw <- vector('list', length(ee$b_Intercept))
 
 set.seed(101)
 for (i in 1:length(ee$b_Intercept)) {
@@ -105,14 +106,14 @@ for (i in 1:length(ee$b_Intercept)) {
   meandelay <- estmean[,i]
   shape <- ee$shape[i]
   
-  reconstruct_time_series_seoul_raw[[i]] <- sapply(1:length(cases), function(x) {
+  reconstruct_time_series_gyeonggi_raw[[i]] <- sapply(1:length(cases), function(x) {
     data_frame(
       confirm=as.Date("2020-01-20")+x-1,
-      onset=confirm-rnbinom(cases[x], mu=meandelay[x], size=shape),
-      infection=onset-floor(rgamma(cases[x], shape=shape, rate=shape/meaninc))
+      onset=confirm-rnbinom(round(cases[x]), mu=meandelay[x], size=shape),
+      infection=onset-floor(rgamma(round(cases[x]), shape=shape, rate=shape/meaninc))
     )
   }, simplify = FALSE) %>%
     bind_rows
 }
 
-save("reconstruct_time_series_seoul_raw", file="reconstruct_time_series_seoul_raw.rda")
[…]

### 2020-03-21 traffic_seoul.R

@@ -2,12 +2,15 @@ library(dplyr)
 library(tidyr)
 library(lubridate)
 library(ggplot2); theme_set(theme_bw())
+library(readxl)
 
-dataname <- c("사용일자", "노선명", "역ID", "역명", "승차총승객수", "하차총승객수", "등록일자")
+whichfilter <- c("1호선", "2호선", "3호선", "4호선", "5호선", "6호선", "7호선", "8호선", "9호선",
+                 "9호선2~3단계", "우이신설선", "9호선2단계")
 
 ss2020 <- read.csv("seoul_subway/서울시 지하철호선별 역별 승하차 인원 정보.csv", fileEncoding="EUC-KR") 
 
 ss2020a <- ss2020 %>%
+  filter(호선명 %in% whichfilter) %>%
   mutate(
     사용일자=gsub("2020", "2020-", 사용일자),
     사용일자=gsub("2020-03", "2020-03-", 사용일자),
@@ -19,8 +22,8 @@ ss2020a <- ss2020 %>%
     year=2020
   )
 
-ss2020b <- read.csv("seoul_subway/CARD_SUBWAY_MONTH_202002.csv", stringsAsFactors=FALSE, fileEncoding="latin1") %>%
-  setNames(dataname) %>%
+ss2020b <- read.csv("seoul_subway/CARD_SUBWAY_MONTH_202002.csv", stringsAsFactors=FALSE, fileEncoding="utf8") %>%
+  filter(노선명 %in% whichfilter) %>%
   mutate(
     사용일자=gsub("2020", "2020-", 사용일자),
     date=as.Date(gsub("2020-02", "2020-02-", 사용일자))
@@ -32,9 +35,8 @@ ss2020b <- read.csv("seoul_subway/CARD_SUBWAY_MONTH_202002.csv", stringsAsFactor
   ) %>%
   filter(date=="2020-02-01")
 
-ss2020c <- read.csv("seoul_subway/CARD_SUBWAY_MONTH_202001.csv", stringsAsFactors=FALSE, fileEncoding="latin1",
-                    row.names = NULL) %>%
-  setNames(c(dataname, "empty")) %>%
+ss2020c <- read_xlsx("seoul_subway/CARD_SUBWAY_MONTH_202001.xlsx", skip=1) %>%
+  filter(노선명 %in% whichfilter) %>%
   mutate(
     사용일자=gsub("2020", "2020-", 사용일자),
     date=as.Date(gsub("2020-01", "2020-01-", 사용일자))
[…]

### 2020-03-21 korea.tex
update

@@
 
-Since its appearance in December 2019, the novel coronavirus disease (COVID-19) has spread internationally, including to South Korea.
+Since its first appearance in Wuhan, China, in December 2019, the novel coronavirus disease (COVID-19) has spread internationally, including to South Korea.
 The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China.
-The epidemic took off in mid-February when COVID-19 began to spread within a church from a city of Daegu --- as of March 18, 2020, 8,413 cases have been confirmed, of which 60\% are linked to the church.
-Despite the lack of draconian measures such as in China, the epidemic has been subsiding in South Korea.
-Its success in controlling the spread has been widely attributed to its intensive testing.
+The epidemic then took off in mid-February when COVID-19 began to spread within a church from a city of Daegu --- as of March 18, 2020, 8,413 cases have been confirmed, of which 60\% are linked to the church.
+While its success in controlling the epidemic has been widely attributed to its intensive testing, other factors also played a role. 
 Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by comparing epidemics in two geographically separated major cities in South Korea.
@@
 Partial line lists were transcribed from press releases and reports from KCDC and various local and provincial government websites.
-All data, sources, and original reports are stored in a publicly available GitHub repoository: \url{https://github.com/parksw3/COVID19-Korea}.
+All data and original reports are stored in a publicly available GitHub repoository: \url{https://github.com/parksw3/COVID19-Korea}.
 
@@
 We compare epidemiological dynamics of COVID-19 from two cities in which the biggest number of COVID-19 cases have been reported: Daegu and Seoul.
-% population size: 2.4 and 9.7 million \citep{kosis}.
 Bewteen January 20--March 16, 2020, 6,083 cases were reported from Daegu and 248 from Seoul.

### 2020-03-21 figure_compare_report.R

@@ -28,7 +28,7 @@ seoul <- geo %>%
 load("traffic_daegu.rda")
 load("traffic_seoul.rda")
 
-traffic_daegu1 <- select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) %>%
+traffic_daegu1 <- dplyr::select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) %>%
   mutate(
     wday=wday(date)
   ) %>%
@@ -36,7 +36,7 @@ traffic_daegu1 <- select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) 
     date >= as.Date("2017-01-20")+3, date <= as.Date("2017-03-01")+3
   )
 
-traffic_daegu2 <- select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) %>%
+traffic_daegu2 <- dplyr::select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) %>%
   mutate(
     wday=wday(date)
   ) %>%
@@ -44,7 +44,7 @@ traffic_daegu2 <- select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) 
     date >= as.Date("2018-01-20")+2, date <= as.Date("2018-03-01")+2
   )
 
-traffic_daegu3 <- select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) %>%
+traffic_daegu3 <- dplyr::select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) %>%
   mutate(
     wday=wday(date)
   ) %>%
@@ -52,7 +52,7 @@ traffic_daegu3 <- select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) 
     date >= as.Date("2019-01-20")+1, date <= as.Date("2019-03-01")+1
   )
 
-traffic_daegu4 <- select(ungroup(filter(traffic_daegu,year==2020)), -월, -일) %>%
+traffic_daegu4 <- dplyr::select(ungroup(filter(traffic_daegu,year==2020)), -월, -일) %>%
   mutate(
     wday=wday(date)
   ) %>%
@@ -110,6 +110,7 @@ g1 <- ggplot(filter(traffic_daegu_all, year==2020)) +
   geom_bar(data=daegu, aes(date_report, cases), stat="identity", alpha=0.5) + 
   geom_line(aes(date, mean/3e2), lwd=1) +
   geom_line(aes(date, traffic/3e2, group=year, col=year), lwd=1) +
[…]

### 2020-03-22 korea.tex
edits

@@
 
-\usepackage{natbib}
-\bibliographystyle{chicago}
+\usepackage[numbers]{natbib}
+\usepackage{hyperref,url}
+\bibliographystyle{unsrt}
 \date{\today}
@@
 }
+\newline
+\\
+Sang Woo Park\textsuperscript{1,*}
+Kaiyuan Sun\textsuperscript{2}
+C\'ecile Viboud\textsuperscript{2}
+Jonathan Dushoff\textsuperscript{3,4,5}
+Bryan T.\ Grenfell\textsuperscript{1,2,6}
+\\
+\bigskip
+\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
+\\
+\textbf{2} Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
+\\
+\textbf{3} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
+\\
+\textbf{4} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
+\\
+\textbf{5} Department of Biology, McMaster University, Hamilton, ON, Canada
+\\
+\textbf{6} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
+\\
+\bigskip
+
+*Corresponding author: swp2@princeton.edu
 \end{flushleft}
@@
 Since its first appearance in Wuhan, China, in December 2019, the novel coronavirus disease (COVID-19) has spread internationally, including to South Korea.
-The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China.
-The epidemic then took off in mid-February when COVID-19 began to spread within a church from a city of Daegu --- as of March 18, 2020, 8,413 cases have been confirmed, of which 60\% are linked to the church.
-While its success in controlling the epidemic has been widely attributed to its intensive testing, other factors also played a role. 
-Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by comparing epidemics in two geographically separated major cities in South Korea.
+The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
+The epidemic then took off in mid-February when COVID-19 began to spread within a church from a city of Daegu --- as of March 18, 2020, 8,413 cases have been confirmed, of which 60\% are related to the church \citep{kcdc}.
+While its success in controlling the epidemic has been widely attributed to its extensive testing \citep{science}, other factors, such as social distancing, are also likely to have played important roles.
+Here, we compare epidemics in two geographically separated major cities in South Korea and use traffic data to study potential roles of social distancing in mitigating the spread of COVID-19.
 
@@
 We analyze epidemiological data, collected bewteen January 20--March 16, 2020, describing the COVID-19 outbreak in Korea.
-Daily number of reported cases in each geographic region was transcribed from press releases by \cite{kcdc}.
+Daily number of reported cases in each geographic region was transcribed from press releases by Korea Centers for Disease Control and Prevention (KCDC) \cite{kcdc}.
 Partial line lists were transcribed from press releases and reports from KCDC and various local and provincial government websites.
@@
 Daily traffic from previous years have been shifted by 1--3 days to align day of the week.
+Vertical lines indicate Feb 18, 2020, when the first case was confirmed in Daegu.
 }
@@
 Bewteen January 20--March 16, 2020, 6,083 cases were reported from Daegu and 248 from Seoul.
-Unlike the epidemic in Daegu, which is characterized by a single, large peak followed by a persistent decrease, the epidemic in Seoul consists of several small outbreaks (\fref{data}).
+Unlike the epidemic in Daegu, which is characterized by a single, large peak followed by a gradual decrease in the number of confirmed cases, the epidemic in Seoul consists of several small outbreaks (\fref{data}).
 
-Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from their corresponding government websites.
-
-\fref{data}
-About 80\% decrease in Daegu and 50\% decrease in Seoul after the identification of the church-related case.
+Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from \url{data.go.kr} and \url{data.seoul.go.kr}.
+\fref{data} compares the daily number of individuals who got on the subway --- including a monorail in Daegu --- across all stations.
+Soon after the first church-related case was confirmed in Daegu on Feb 18, 2020, the daily traffic volume decreased by about 80\% and 50\% compared to previous years in Daegu and Seoul, respectively.
 
@@
 
-\section{Effective reproduction number}
+\section{Trends in time-dependent reproduction number and traffic volume}
 
-We estimate the time-varying effective reproduction number $R_t$ for both epidemics.
-We account for changes in case definition by scaling the number of reported cases by the relative detection rate (i.e., a proportion of cases tested positive based a case definition divided by the mean detection rate among case definitions);
-using a narrow case definition is more likely to have a higher detection rate and a lower reporting rate than a wide one.
-Likewise, we account for changes in the number of tests by scaling the number of reported cases again by the relative number of tests performed on each day; this step is performed separately for each case definition because the widening of a case definition necessarily increases the number of tests.
+We estimate the time-dependent reproduction number $R_t$ (i.e., the expected number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}) for both epidemics.
+We account for changes in testing criteria, which occurred 5 times, by multiplying the number of positive cases by their relative detection rate (i.e., a proportion of cases tested positive based on a criterion divided by the mean detection rate).
+Likewise, we account for changes in the number of tests completed on each day by dividing the number of positive cases again by the relative number of tests completed on each day; this step is performed separately for each testing criterion because the widening of a criterion necessarily increases the number of tests.
+A sensitivity analysis showed that qualitative patterns of inferred $R_t$ are robust to these adjustments (Supplementary Materials).
 
-We then reconstruct the incidence time series from the scaled time series of reported cases by using the backward confirmation-to-onset delay distribution, inferred from the partial line list, and previously estimated incubation period distribution.
-To account for right-censoring (i.e., cases that are infected but have not been reported yet), we estimate the forward onset-to-confirmation delay distribution and estimate the probability that a case infected on a given day will be reported before March 16, 2020.
-Then, we divide the reconstructed incidence time series by this probability and estimate the effective reproduction number using the renewal equation using a 14-day sliding window.
-We only calculate the effective reproduction number until March 10, 2020 because the degree of right-censoring too large afterwards.
+We estimate the time-varying \emph{backward} confirmation-to-onset delay distributions from the partial line list and combine them with previously estimated incubation period distribution \citep{backer2020incubation} to infer the probability distributions for date of infection of each confirmed case.
+To account for right-censoring (i.e., individuals that are infected but have not been confirmed yet), we estimate the time-varying \emph{forward} onset-to-confirmation delay distribution, and infer the probability that a case infected on a given day will be reported before March 16, 2020.
+Then, we divide the reconstructed incidence time series (i.e., the number of infected individuals on day $t$) by this probability and estimate the time-dependent reproduction number using the renewal equation using a 14-day sliding window \citep{fraser2007estimating}:
+\begin{equation}
+\mathcal R_t = \frac{I_t}{\sum_{k=1}^{14} I_{t-k} w_k},
+\end{equation}
+where $I_t$ is the censoring-adjusted incidence time series and $w_k$ is the generation-interval distribution.
+We only calculate the time-dependent reproduction number until March 10, 2020 because the degree of right-censoring too large afterwards.
+Details are provided in the Supplementary Materials.
 
@@
 \caption{
-\textbf{Comparison of effective reproduction number and reconstructed incidence time series in Daegu and Seoul.}
+\textbf{Comparison of time-dependent reproduction number and normalized traffic in Daegu and Seoul.}
+Black lines represent the median estimates of $R_t$. 
+Gray ribbons represent the corresponding 95\% credible intervals.
+Red lines represent the normalized traffic volume.
+Vertical lines indicate Feb 18, 2020, when the first case was confirmed in Daegu.
 }
@@
 
-\fref{eff} compares the estimates of $R_t$ in Daegu and Seoul.
-While $R_t$ drops below 1 in Daegu shortly after its first case was confirmed (February 18, 2020), 
-the decrease is unclear in Seoul.
-Similar patterns in $R_t$ can be found in adjacent provinces -- a sharp decrease in Gyeongsangbuk-do, which surrounds Daegu, and a persistent $R_t$ around 1 in Gyeonggi-do, which surrounds Seoul (Supplementary Materials).
-
-\section{Effect of social distancing on the effective reproduction number}
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figure_traffic_regression.pdf}
-\caption{
-\textbf{Other effects.}
-}
-\label{fig:traffic}
-\end{figure}
-
-\begin{itemize}
-  \item Traffic negatively correlated with $R_t$
-  \item Geographically differential effect of traffic on $R_t$; other routes of transportation, other factors we don't consider
-  \item Even if we can get traffic to 0 in Seoul, $R_t$ is not less than 1...?
-\end{itemize}
+\fref{eff} compares the estimates of $\mathcal R_t$ in Daegu and Seoul.
+Estimates of $\mathcal R_t$ gradually decrease in Daegu and eventually drop below 1 about a week after the introduction of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}A);
+our estimates of $\mathcal R_t$ are consistent with \cite{tempvar}, although their $\mathcal R_t$ estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
+On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}B), suggesting that the degree of social distancing may have been insufficient to prevent the spread in Seoul;
+stronger distancing or other control measures will be required to decrease $\mathcal R_t$ below 1.
+Similar patterns in the estimates of $\mathcal R_t$ are found in adjacent provinces, providing support for the robustness of our analysis (Supplementary Materials).
 
@@
 
-\begin{itemize}
-  \item Provides indirect evidence of effect of social distancing on epidemic dynamics
-  \item But social distancing alone might not be sufficient... others have said something
-  \item SK might need stronger intervention in some places
-\end{itemize}
+The ongoing COVID-19 outbreak in South Korea provides a unique perspective to understanding and controlling the pandemic.
+Its experience provides evidence that the epidemic can be contained without draconian measures as in China.
+Our analysis provides indirect, but clear, evidence that social distancing is likely to have assisted in mitigating the epidemic in South Korea.
+Even though social distancing alone may not be able to fully prevent the spread of the disease, its ability to flatten the epidemic curve (cf.
+\fref{eff}B) reduces burden for healthcare system and provides time to plan for the future \citep{anderson2020will}.
+
+There are several limitations to our study.
+
+Finally, while there is a clear
 
@@
 \end{itemize}
+
+\section*{Contribution}
+
+Data collection: SWP; conceptualisation: SWP and CV; analysis: SWP; first draft: SWP, JD. All authors contributed to the writing and approval of the final report.
 

### 2020-03-22 figure_compare_R_t.R

@@ -1,10 +1,33 @@
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
+library(lubridate)
+library(dplyr)
+library(tidyr)
+library(readxl)
 
 load("R_t_daegu_censor.rda")
 load("R_t_seoul_censor.rda")
 load("traffic_daegu.rda")
 load("traffic_seoul.rda")
+
+geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
+  mutate(
+    date_report=as.Date(date_report)
+  ) 
+
+geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
+
+daegu <- geo %>% 
+  group_by(date_report) %>%
+  summarize(
+    cases=sum(Daegu, na.rm=TRUE)
+  )
+
+seoul <- geo %>% 
+  group_by(date_report) %>%
+  summarize(
+    cases=sum(Seoul, na.rm=TRUE)
+  )
 
 traffic_daegu1 <- select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) %>%
   mutate(
@@ -101,12 +124,13 @@ seoul_traffic <- data.frame(
 )
 
 g1 <- ggplot(rt_daegu) +
-  geom_bar(data=daegu, aes(date_report, cases/max(daegu$cases)), stat="identity", alpha=0.5) + 
+  geom_bar(data=daegu, aes(date_report, cases/max(daegu$cases)), stat="identity", alpha=0.3) + 
   geom_hline(yintercept=6, lty=2, col=2) + 
-  geom_line(data=daegu_traffic, aes(date, traffic*6), col=2, lwd=1) +
+  geom_line(data=daegu_traffic, aes(date, traffic*6), col=2) +
   geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.3) +
-  geom_line(aes(date, median)) +
+  geom_line(aes(date, median), lwd=1) +
   geom_hline(yintercept=1, lty=2) + 
+  geom_vline(xintercept=as.Date("2020-02-18"), lty=2) +
   scale_color_manual(values=c(1, 2, 4)) +
   scale_fill_manual(values=c(1, 2, 4)) +
[…]

### 2020-03-22 korea.tex
edits

@@
 While its success in controlling the epidemic has been widely attributed to its extensive testing \citep{science}, other factors, such as social distancing, are also likely to have played important roles.
-Here, we compare epidemics in two geographically separated major cities in South Korea and use traffic data to study potential roles of social distancing in mitigating the spread of COVID-19.
+Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by using metro traffic data to compare epidemics in two geographically separated major cities in South Korea.
 
@@
 
-We analyze epidemiological data, collected bewteen January 20--March 16, 2020, describing the COVID-19 outbreak in Korea.
+We analyzed epidemiological data, collected bewteen January 20--March 16, 2020, describing the COVID-19 outbreak in South Korea.
 Daily number of reported cases in each geographic region was transcribed from press releases by Korea Centers for Disease Control and Prevention (KCDC) \cite{kcdc}.
-Partial line lists were transcribed from press releases and reports from KCDC and various local and provincial government websites.
-All data and original reports are stored in a publicly available GitHub repoository: \url{https://github.com/parksw3/COVID19-Korea}.
+Partial line lists were transcribed from press releases from KCDC and various local and provincial governments.
+All data and original reports are stored in a publicly available GitHub repository: \url{https://github.com/parksw3/COVID19-Korea}.
 
@@
 \textbf{Comparison of epidemiological and traffic data from Daegu and Seoul.}
-Red lines represent the daily metro traffic volume in 2020.
-Black lines represent the average daily metro traffic volume in previous years, 2017--2019.
+Solid lines represent the daily metro traffic volume in 2020 (red) and mean daily traffic volume between 2017--2019 (black).
 Daily traffic from previous years have been shifted by 1--3 days to align day of the week.
@@
 
-We compare epidemiological dynamics of COVID-19 from two cities in which the biggest number of COVID-19 cases have been reported: Daegu and Seoul.
-Bewteen January 20--March 16, 2020, 6,083 cases were reported from Daegu and 248 from Seoul.
-Unlike the epidemic in Daegu, which is characterized by a single, large peak followed by a gradual decrease in the number of confirmed cases, the epidemic in Seoul consists of several small outbreaks (\fref{data}).
+We compared epidemiological dynamics of COVID-19 from two cities in which the biggest number of COVID-19 cases have been reported: Daegu and Seoul.
+Between January 20--March 16, 2020, 6,083 cases were reported from Daegu and 248 from Seoul.
+Unlike the epidemic in Daegu, which is characterized by a single, large peak followed by a gradual decrease, the epidemic in Seoul consists of several small outbreaks (\fref{data}).
 
-Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from \url{data.go.kr} and \url{data.seoul.go.kr}.
-\fref{data} compares the daily number of individuals who got on the subway --- including a monorail in Daegu --- across all stations.
+Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from \url{data.go.kr} and \url{data.seoul.go.kr}, respectively.
+We compared the the daily number of individuals who got on the subway --- including a monorail in Daegu --- across all stations (\fref{data}).
 Soon after the first church-related case was confirmed in Daegu on Feb 18, 2020, the daily traffic volume decreased by about 80\% and 50\% compared to previous years in Daegu and Seoul, respectively.
@@
 
-We estimate the time-dependent reproduction number $R_t$ (i.e., the expected number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}) for both epidemics.
-We account for changes in testing criteria, which occurred 5 times, by multiplying the number of positive cases by their relative detection rate (i.e., a proportion of cases tested positive based on a criterion divided by the mean detection rate).
-Likewise, we account for changes in the number of tests completed on each day by dividing the number of positive cases again by the relative number of tests completed on each day; this step is performed separately for each testing criterion because the widening of a criterion necessarily increases the number of tests.
+\begin{figure}
+\includegraphics[width=\textwidth]{figure_report_delay.pdf}
+\caption{
+\textbf{Changes in the number of tests and delay distributions over time.}
+}
+Vertical lines indicate the date on which testing criteria expanded.
+\label{fig:dist}
+\end{figure}
+
+In order to estimate the time-dependent reproduction number $R_t$ (i.e., the expected number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), 
+we first accounted for changes in testing criteria (\fref{dist}B) by multiplying the daily number of reported cases by the relative detection rate for each criterion (i.e., a proportion of cases tested positive based on a criterion divided by the mean detection rate).
+Likewise, we accounted for changes in the number of tests completed on each day (\fref{dist}A) by dividing the scaled number of reported cases by the relative number of tests completed on each day; this step was performed separately for each testing criterion because widening of criteria necessarily increases the number of tests.
 A sensitivity analysis showed that qualitative patterns of inferred $R_t$ are robust to these adjustments (Supplementary Materials).
 
-We estimate the time-varying \emph{backward} confirmation-to-onset delay distributions from the partial line list and combine them with previously estimated incubation period distribution \citep{backer2020incubation} to infer the probability distributions for date of infection of each confirmed case.
-To account for right-censoring (i.e., individuals that are infected but have not been confirmed yet), we estimate the time-varying \emph{forward} onset-to-confirmation delay distribution, and infer the probability that a case infected on a given day will be reported before March 16, 2020.
-Then, we divide the reconstructed incidence time series (i.e., the number of infected individuals on day $t$) by this probability and estimate the time-dependent reproduction number using the renewal equation using a 14-day sliding window \citep{fraser2007estimating}:
+We then estimated time-dependent \emph{backward} confirmation-to-onset delay distributions from the partial line list using a negative-binomial regression with \texttt{brms} package \citep{burkner2017brms} (\fref{dist}C).
+We combined the estimated posterior distribution of delay distributions with previously estimated incubation period distribution \citep{backer2020incubation} to obtain posterior samples for date of infection for each confirmed case.
+Likewise, we estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using a negative-binomial regression while accounting for right-censoring (\fref{dist}D).
+This allowed us to calculate the degree of censoring (i.e., median probability that a case infected on a given day will be reported after March 16, 2020).
+Details are provided in the Supplementary Materials.
+
+For each posterior sample of the reconstructed incidence time series (i.e., the number of infected individuals on day $t$), we divided incidence by $1-\textrm{degree of censoring}$ and estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
 \begin{equation}
@@
 \end{equation}
-where $I_t$ is the censoring-adjusted incidence time series and $w_k$ is the generation-interval distribution.
-We only calculate the time-dependent reproduction number until March 10, 2020 because the degree of right-censoring too large afterwards.
-Details are provided in the Supplementary Materials.
+where $I_t$ is the censoring-adjusted incidence time series and $w_k$ is the assumed generation-interval distribution.
+We restricted the calculation of the time-dependent reproduction number between February 2, 2020 (14 days after the first confirmed case was imported) and March 10, 2020 (because the degree of censoring would be too large after this date).
 
@@
 \fref{eff} compares the estimates of $\mathcal R_t$ in Daegu and Seoul.
-Estimates of $\mathcal R_t$ gradually decrease in Daegu and eventually drop below 1 about a week after the introduction of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}A);
-our estimates of $\mathcal R_t$ are consistent with \cite{tempvar}, although their $\mathcal R_t$ estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
+Estimates of $\mathcal R_t$ gradually decrease in Daegu and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}A).
+Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
 On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}B), suggesting that the degree of social distancing may have been insufficient to prevent the spread in Seoul;
-stronger distancing or other control measures will be required to decrease $\mathcal R_t$ below 1.
+stronger distancing or further intervention will be necessary to decrease $\mathcal R_t$ below 1.
 Similar patterns in the estimates of $\mathcal R_t$ are found in adjacent provinces, providing support for the robustness of our analysis (Supplementary Materials).
@@
 
-There are several limitations to our study.
+There are many other factors that allowed South Korea to control the COVID-19 outbreak.
+In particular, the large number of tests as well as the intensity of the testing are likely to have played major roles.
 
-Finally, while there is a clear
+Furthermore, other intervention measures, such as school closures, are likely to have reduced transmission rates.
 
@@
 
-Data collection: SWP; conceptualisation: SWP and CV; analysis: SWP; first draft: SWP, JD. All authors contributed to the writing and approval of the final report.
+Data collection: SWP; analysis: SWP. All authors contributed to the writing and approval of the final report.
 
@@
 
-\section{Supplementary Materials}
+\pagebreak
+
+\renewcommand\thefigure{S\arabic{figure}}
+\setcounter{figure}{0}
+
+\section*{Supplementary Materials}
+
+\subsection*{Estimating delay distributions between onset and reporting}
+
+We estimated both forward and backward delay distributions.
 

### 2020-03-22 figure_report_delay.R

@@ -8,8 +8,12 @@ library(rstan)
 library(gridExtra)
 library(ggplot2); theme_set(theme_bw())
 load("report_delay_censor.rda")
+load("report_delay_backward.rda")
 
-covid_test <- read_xlsx("data/COVID19-Korea-2020-03-11.xlsx", na="NA", sheet=2) 
+covid_test <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=2)
+
+covid_test$negative[7] <- (covid_test$negative[6]+covid_test$negative[8])/2
+covid_test$negative[13] <- (covid_test$negative[12]+covid_test$negative[14])/2
 
 covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
   mutate(
@@ -99,17 +103,45 @@ estmean <- sapply(1:length(ee$Intercept), function(i) {
 })
 
 estdata <- data.frame(
-  date=seq.Date(from=as.Date("2020-01-10"), to=as.Date("2020-03-11"), by=1),
+  date=seq.Date(from=as.Date("2020-01-10"), to=as.Date("2020-03-10"), by=1),
   median=apply(estmean, 1, median),
   lwr=apply(estmean, 1, quantile, 0.025),
   upr=apply(estmean, 1, quantile, 0.975)
-) %>%
-  group_by(date) %>%
-  summarize(
-    median=median(median),
-    lwr=median(lwr),
-    upr=median(upr)
-  )
+)
+
+covid_delay3 <- covid_all %>%
+  filter(!is.na(date_onset), !is.na(date_confirm)) %>%
+  mutate(
+    date_confirm=as.Date(date_confirm),
+    delay=yday(date_confirm)-yday(date_onset),
+    day=yday(date_confirm),
+    day=day-min(day)
+  ) %>%
+  filter(yday(date_confirm) <= yday(as.Date("2020-03-11")))
+
+covid_delay_fake2 <- data.frame(
+  day=covid_delay3$day[covid_delay3$date_confirm==as.Date("2020-01-20")]:56,
+  delay=1
+)
+
+ex2 <- extract_draws(report_delay_backward, newdata=covid_delay_fake2)
+
+Xs2 <- ex2$dpars$mu$sm$fe$Xs
+
+Zs2 <- ex2$dpars$mu$sm$re$sday$Zs[[1]]
+
[…]

### 2020-03-22 figure_traffic_regression.R

@@ -1,121 +1,0 @@
-library(ggplot2); theme_set(theme_bw())
-
-load("R_t_daegu_censor.rda")
-load("R_t_seoul_censor.rda")
-load("traffic_daegu.rda")
-load("traffic_seoul.rda")
-
-traffic_daegu1 <- select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) %>%
-  mutate(
-    wday=wday(date)
-  ) %>%
-  filter(
-    date >= as.Date("2017-02-02")+3, date <= as.Date("2017-03-01")+3
-  )
-
-traffic_daegu2 <- select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) %>%
-  mutate(
-    wday=wday(date)
-  ) %>%
-  filter(
-    date >= as.Date("2018-02-02")+2, date <= as.Date("2018-03-01")+2
-  )
-
-traffic_daegu3 <- select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) %>%
-  mutate(
-    wday=wday(date)
-  ) %>%
-  filter(
-    date >= as.Date("2019-02-02")+1, date <= as.Date("2019-03-01")+1
-  )
-
-traffic_daegu4 <- select(ungroup(filter(traffic_daegu,year==2020)), -월, -일) %>%
-  mutate(
-    wday=wday(date)
-  ) %>%
-  filter(
-    date >= as.Date("2020-02-02"), date <= as.Date("2020-02-29")
-  )
-
-rt_daegu <- R_t_daegu_censor %>%
-  bind_rows(.id="sim") %>%
-  group_by(date) %>%
-  summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE),
-    location="Daegu"
-  ) %>%
-  filter(date <= as.Date("2020-02-29")) %>%
-  mutate(
-    traffic=traffic_daegu4$total/((traffic_daegu1$total + traffic_daegu2$total + traffic_daegu3$total)/3)
-  )
-
-traffic_seoul1 <- ungroup(filter(traffic_seoul,year==2017)) %>%
-  mutate(
-    wday=wday(date)
-  ) %>%
-  filter(
-    date >= as.Date("2017-02-02")+3, date <= as.Date("2017-03-11")+3
-  )
-
-traffic_seoul2 <- ungroup(filter(traffic_seoul,year==2018)) %>%
[…]

### 2020-03-22 infection_delay.R

@@ -1,23 +1,0 @@
-library(rstan)
-load("report_delay_censor.rda")
-load("pred_report_delay.rda")
-
-# pgamma(7.7, 175, 175/6.4) - pgamma(5.6, 175, 175/6.4)
-# pgamma(3.7, 35, 35/2.3) - pgamma(1.7, 35, 35/2.3)
-
-ee <- extract(report_delay_censor)
-
-for (i in 1:8000) {
-  mu_inc <- rgamma(1, 175, 175/6.4)
-  sd_inc <- rgamma(1, 35, 35/2.3)
-  
-  
-}
-
-plot(estmean[,1])
-
-ee$shape
-
-
-
-dnbinom()

### 2020-03-22 korea.tex
oops

@@
 	\textbf\newline{
-		Potential roles of social distancing in preventing the spread of coronavirus disease 2019 (COVID-19) in South Korea
+		Potential roles of social distancing in mitigating the spread of coronavirus disease 2019 (COVID-19) in South Korea
 	}
@@
 
-Since its first appearance in Wuhan, China, in December 2019, the novel coronavirus disease (COVID-19) has spread internationally, including to South Korea.
+Since its first appearance in Wuhan, China, in December 2019 \citep{pneumonia}, the novel coronavirus disease (COVID-19) has spread internationally, including to South Korea.
 The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
 The epidemic then took off in mid-February when COVID-19 began to spread within a church from a city of Daegu --- as of March 18, 2020, 8,413 cases have been confirmed, of which 60\% are related to the church \citep{kcdc}.
-While its success in controlling the epidemic has been widely attributed to its extensive testing \citep{science}, other factors, such as social distancing, are also likely to have played important roles.
+While its success in suppressing the epidemic has been widely attributed to its intensive testing \citep{fergusonimpact, science}, other factors, such as social distancing, are also likely to have played important roles.
 Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by using metro traffic data to compare epidemics in two geographically separated major cities in South Korea.
@@
 Solid lines represent the daily metro traffic volume in 2020 (red) and mean daily traffic volume between 2017--2019 (black).
-Daily traffic from previous years have been shifted by 1--3 days to align day of the week.
+Daily traffic from previous years have been shifted by 1--3 days to align day of the weeks.
 Vertical lines indicate Feb 18, 2020, when the first case was confirmed in Daegu.
@@
 Vertical lines indicate the date on which testing criteria expanded.
+Boxplots (C--D) represent the observed delays.
+Black lines and gray ribbons represent the median estimates of the mean delays and their associated 95\% credible intervals.
 \label{fig:dist}
@@
 We then estimated time-dependent \emph{backward} confirmation-to-onset delay distributions from the partial line list using a negative-binomial regression with \texttt{brms} package \citep{burkner2017brms} (\fref{dist}C).
-We combined the estimated posterior distribution of delay distributions with previously estimated incubation period distribution \citep{backer2020incubation} to obtain posterior samples for date of infection for each confirmed case.
-Likewise, we estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using a negative-binomial regression while accounting for right-censoring (\fref{dist}D).
-This allowed us to calculate the degree of censoring (i.e., median probability that a case infected on a given day will be reported after March 16, 2020).
+We combined the estimated posterior distribution of delay distributions with previously estimated incubation period distribution (Table 1) to obtain posterior samples for date of infection for each confirmed case.
+Likewise, we estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using the same negative-binomial regression model, while accounting for right-censoring (\fref{dist}D),
+to calculate the degree of censoring (i.e., median probability that a case infected on a given day will be reported after March 16, 2020).
 Details are provided in the Supplementary Materials.
+
+\begin{table}[t]
+\begin{center}
+\small
+\begin{tabular}{p{4cm}|p{4.5cm}|p{4.7cm}|l}
+ & Parameterization & Priors & Source \\
+\hline
+Incubation period\newline distribution & $\mathrm{Gamma}(\mu_I, \mu_I^2/\sigma^2)$ & $\mu_I\sim \mathrm{Gamma}(6.5\,\textrm{days}, 145)$\newline$\sigma\sim \mathrm{Gamma}(2.6, 25)$ & \citep{backer2020incubation} \\
+\hline
+Generation-interval\newline distribution & $\mathrm{NegativeBinomial}(\mu_G, \theta)$ & $\mu_G\sim\mathrm{Gamma}(5\,\textrm{days},62)$\newline$\theta\sim\mathrm{Gamma}(5,20)$ & \citep{ferretti2020quantifying, ganyani2020estimating} \\
+\hline
+\end{tabular}
+\end{center}
+\caption{
+\textbf{Assumed incubation and generation-interval distributions.}
+Gamma distributions are parameterized using its mean and shape.
+Negative binomial distributions are parameterized using its mean and dispersion.
+Priors are chosen such that the 95\% quantiles of prior means and standard deviations are consistent with previous estiamtes.
+}
+\end{table}
 
@@
 \end{equation}
-where $I_t$ is the censoring-adjusted incidence time series and $w_k$ is the assumed generation-interval distribution.
-We restricted the calculation of the time-dependent reproduction number between February 2, 2020 (14 days after the first confirmed case was imported) and March 10, 2020 (because the degree of censoring would be too large after this date).
+where $I_t$ is the censoring-adjusted incidence time series and $w_k$ is the generation-interval distribution randomly drawn from a prior distribution (Table 1).
+We restricted the calculation of $\mathcal R_t$ between February 2, 2020 (14 days after the first confirmed case was imported) and March 10, 2020 (because the degree of censoring would be too large after this date to estimate $\mathcal R_t$).
 
@@
 \textbf{Comparison of time-dependent reproduction number and normalized traffic in Daegu and Seoul.}
-Black lines represent the median estimates of $R_t$. 
-Gray ribbons represent the corresponding 95\% credible intervals.
+Black lines and gray ribbons represent the median estimates of $R_t$ and their corresponding 95\% credible intervals.
 Red lines represent the normalized traffic volume.
@@
 \fref{eff} compares the estimates of $\mathcal R_t$ in Daegu and Seoul.
-Estimates of $\mathcal R_t$ gradually decrease in Daegu and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}A).
-Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
-On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}B), suggesting that the degree of social distancing may have been insufficient to prevent the spread in Seoul;
-stronger distancing or further intervention will be necessary to decrease $\mathcal R_t$ below 1.
-Similar patterns in the estimates of $\mathcal R_t$ are found in adjacent provinces, providing support for the robustness of our analysis (Supplementary Materials).
+In Daegu, estimates of $\mathcal R_t$ gradually decrease and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}A).
+Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
+On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}B);
+while social distancing is likely to have assisted in mitigating the spread in Seoul, our analysis suggests that 50\% decrease in metro traffic was insufficient to suppress the spread.
+Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.
+Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces, providing support for the robustness of our analysis (Supplementary Materials).
 
@@
 The ongoing COVID-19 outbreak in South Korea provides a unique perspective to understanding and controlling the pandemic.
-Its experience provides evidence that the epidemic can be contained without draconian measures as in China.
-Our analysis provides indirect, but clear, evidence that social distancing is likely to have assisted in mitigating the epidemic in South Korea.
+Its experience provides evidence that the epidemic can be suppressed without draconian measures as in China \citep{kickbusch2020response}.
+Our analysis reveals potential roles of social distancing in mitigating the COVID-19 epidemic in South Korea.
 Even though social distancing alone may not be able to fully prevent the spread of the disease, its ability to flatten the epidemic curve (cf.
@@
 
-There are many other factors that allowed South Korea to control the COVID-19 outbreak.
-In particular, the large number of tests as well as the intensity of the testing are likely to have played major roles.
+Our study is not without limitations.
+In particular, we did not account for differences in the delay distributions or changes in the number of tests or detection rates among cities.
+The intensity of intervention is likely to vary across regions given that majority of COVID-19 cases in South Korea were reported from Daegu.
+We did not have the data to account for these factors.
+Nonetheless, the robustness of our findings is supported by the sensitivity analyses.
 
-Furthermore, other intervention measures, such as school closures, are likely to have reduced transmission rates.
+Our analysis focused on comparing metro traffic, which serves as a proxy for the degree of social distancing, with epidemiological dynamics in two cities.
+However, our analysis does not allow us to measure the direct ``effect'' of social distancing on epidemiological dynamics.
+Moreover, other measures, such as intensive testing of core transmission groups and school closure, are also likely to have affected the changes in $\mathcal R_t$ \citep{kcdc}.
+Future studies should consider quantifying contributions of different measures in preventing the spread.
 
-Caveats:
-\begin{itemize}
-  \item Only two cities...although two coupled provinces show strikingly similar patterns
-  \item Correlation not causation
-  \item Did not account for geographic heterogeneity in testing or etc. Estimation of delay distribution etc all based on national line list.
-\end{itemize}
+Finally, our studies highlights the importance of characterizing geographical heterogeneity in estimating epidemic potential.
+The recent decrease in the number of reported cases in South Korea is likely to be strongly correlated with the decrease in the number of cases in Daegu.
+Our analysis reveals that the epidemic may still persist in other regions, including Seoul and Gyeonggi-do.
+Unless the reproduction number can be reduced below 1 in all regions, small outbreaks may continue to occur in South Korea.
 
@@
 
-Data collection: SWP; analysis: SWP. All authors contributed to the writing and approval of the final report.
+Data collection: SWP; conceptualization: SWP, CV; analysis: SWP; first draft: SWP, JD. All authors contributed to the writing and approval of the final report.
 
@@
 
-We estimated both forward and backward delay distributions.
+We estimated both forward (onset-to-confirmation) and backward (confirmation-to-onset) delay distributions using Bayesian negative binomial regressions with log links.
+The backward delay distribution is inferred directly using the \texttt{brms} package \citep{burkner2017brms}.
+The time-dependent mean of the negative binomial distribution is modeled using splines.
+We assumed weakly informative priors on the fixed effects: normal distributions with mean of 0 and standard devition of 2;
+note that these distributions are priors on link scale.
+
+To estimate the forward delay distribution, we modified the stan code to account for right-censoring and ran the code using the \texttt{rstan} package \citep{rstan}.
+In particular, we modified the likelihood such that given a delay of $x_i$ days, symptom onset day $t_i$ and the day of measurement of $k$, the likelihood of observing the delay is given by:
+\begin{equation}
+\frac{f(x_i|\mu(t_i), \theta)}{F(k-t_i|\mu(t_i), \theta)},
+\end{equation}
+where $f$ is the negative binomial distribution with time-dependent mean $\mu(t_i)$ and dispersion parameter $\theta$.
+This likelihood accounts for the fact that the delay between symptom onset and confirmation cannot be longer than $k-t_i$ (otherwise, the case will be reported after the date of measurement).
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is lower than 1.01 for all parameters \citep{gelman1992inference}.
+
+\pagebreak
 
@@
 \caption{
-\textbf{Comparison of effective reproduction number in Daegu, Seoul, Gyeongsangbuk-do, and Gyeonggi-do.}
+\textbf{Comparison of effective reproduction number and the daily number of reported cases in Daegu, Seoul, Gyeongsangbuk-do, and Gyeonggi-do.}
+}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figure_R_t_daegu.pdf}
+\caption{
+\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Daegu.}
+}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figure_R_t_seoul.pdf}
+\caption{
+\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Seoul.}
+}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figure_R_t_gyeongbuk.pdf}
+\caption{
+\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Gyeongsangbuk-do.}
+}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figure_R_t_gyeonggi.pdf}
+\caption{
+\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Gyeonggi-do.}
 }

### 2020-03-22 figure_R_t_daegu.R

@@ -104,7 +104,8 @@ g1 <- ggplot(geo) +
   scale_color_manual(values=c(1, 2, 4)) +
   scale_fill_manual(values=c(1, 2, 4)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-19", "2020-03-14"))) +
-  scale_y_continuous("Reconstructed incidence", limits=c(0, 940), expand=c(0, 0),
+  scale_y_continuous("Reconstructed incidence", limits=c(0, 1100), expand=c(0, 0),
+                     breaks=c(0, 300, 600, 900),
                      sec.axis = sec_axis(~ .*1, name = "Daily number of reported cases")) +
   theme(
     panel.grid = element_blank(),
@@ -123,7 +124,8 @@ g2 <- ggplot(rt_all) +
   scale_color_manual(values=c(1, 2, 4)) +
   scale_fill_manual(values=c(1, 2, 4)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-19", "2020-03-14"))) +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 6.5), expand=c(0, 0)) +
+  scale_y_continuous("Effective reproduction number", limits=c(0, 11), expand=c(0, 0),
+                     breaks=c(0, 3, 6, 9)) +
   theme(
     panel.grid = element_blank(),
     panel.border = element_blank(),

### 2020-03-22 figure_R_t_gyeonggi.R

@@ -122,14 +122,14 @@ g2 <- ggplot(rt_all) +
   scale_color_manual(values=c(1, 2, 4)) +
   scale_fill_manual(values=c(1, 2, 4)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-19", "2020-03-14"))) +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 13), expand=c(0, 0)) +
+  scale_y_continuous("Effective reproduction number", limits=c(0, 25), expand=c(0, 0)) +
   theme(
     panel.grid = element_blank(),
     panel.border = element_blank(),
     axis.line = element_line(),
     legend.position = "none",
     legend.title = element_blank(),
-    plot.margin = margin(5, 39, 5, 29, unit="pt")
+    plot.margin = margin(5, 32, 5, 20, unit="pt")
   )
 
 gtot <- arrangeGrob(g1, g2, nrow=2)

### 2020-03-22 figure_R_t_seoul.R

@@ -121,15 +121,14 @@ g2 <- ggplot(rt_all) +
   scale_color_manual(values=c(1, 2, 4)) +
   scale_fill_manual(values=c(1, 2, 4)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-19", "2020-03-14"))) +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 10), expand=c(0, 0),
-                     breaks=c(0, 2, 4, 6, 8)) +
+  scale_y_continuous("Effective reproduction number", limits=c(0, 15), expand=c(0, 0)) +
   theme(
     panel.grid = element_blank(),
     panel.border = element_blank(),
     axis.line = element_line(),
     legend.position = "none",
     legend.title = element_blank(),
-    plot.margin = margin(5, 38, 5, 30, unit="pt")
+    plot.margin = margin(5, 38, 5, 25, unit="pt")
   )
 
 gseoul <- arrangeGrob(g1, g2, nrow=2)

### 2020-03-22 figure_compare_R_t.R

@@ -134,7 +134,7 @@ g1 <- ggplot(rt_daegu) +
   scale_color_manual(values=c(1, 2, 4)) +
   scale_fill_manual(values=c(1, 2, 4)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-20", "2020-03-16"))+c(0,0.5)) +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 8.5), expand=c(0, 0),
+  scale_y_continuous("Effective reproduction number", limits=c(0, 9), expand=c(0, 0),
                      sec.axis = sec_axis(~ .*1/6, name = "(Daily traffic, 2020)/(Mean daily traffic, 2017 - 2019)")) +
   ggtitle("A. Daegu") +
   theme(
@@ -159,7 +159,7 @@ g2 <- ggplot(rt_seoul) +
   geom_vline(xintercept=as.Date("2020-02-18"), lty=2) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-20", "2020-03-16"))+c(0,0.5)) +
   ggtitle("B. Seoul") +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 8.5), expand=c(0, 0),
+  scale_y_continuous("Effective reproduction number", limits=c(0, 9), expand=c(0, 0),
                      sec.axis = sec_axis(~ ./6, name = "(Daily traffic, 2020)/(Mean daily traffic, 2017 - 2019)")) +
   theme(
     panel.grid = element_blank(),

### 2020-03-22 reconstruct_time_series_daegu.R

@@ -106,7 +106,7 @@ for (i in 1:length(ee$b_Intercept)) {
   ## pgamma(4.2, 25, 25/2.6) - pgamma(1.8, 25, 25/2.6)
   sdinc <- rgamma(1, 25, 25/2.6)
   
-  shape <- meaninc^2/sdinc^2
+  shapeinc <- meaninc^2/sdinc^2
   
   meandelay <- estmean[,i]
   shape <- ee$shape[i]
@@ -115,7 +115,7 @@ for (i in 1:length(ee$b_Intercept)) {
     data_frame(
       confirm=as.Date("2020-01-20")+x-1,
       onset=confirm-rnbinom(round(cases[x]), mu=meandelay[x], size=shape),
-      infection=onset-floor(rgamma(round(cases[x]), shape=shape, rate=shape/meaninc))
+      infection=onset-floor(rgamma(round(cases[x]), shape=shapeinc, rate=shapeinc/meaninc))
     )
   }, simplify = FALSE) %>%
     bind_rows

### 2020-03-22 reconstruct_time_series_gyeonggi_detect.R

@@ -105,7 +105,7 @@ for (i in 1:length(ee$b_Intercept)) {
   ## pgamma(4.2, 25, 25/2.6) - pgamma(1.8, 25, 25/2.6)
   sdinc <- rgamma(1, 25, 25/2.6)
   
-  shape <- meaninc^2/sdinc^2
+  shapeinc <- meaninc^2/sdinc^2
   
   meandelay <- estmean[,i]
   shape <- ee$shape[i]
@@ -114,7 +114,7 @@ for (i in 1:length(ee$b_Intercept)) {
     data_frame(
       confirm=as.Date("2020-01-20")+x-1,
       onset=confirm-rnbinom(round(cases[x]), mu=meandelay[x], size=shape),
-      infection=onset-floor(rgamma(round(cases[x]), shape=shape, rate=shape/meaninc))
+      infection=onset-floor(rgamma(round(cases[x]), shape=shapeinc, rate=shapeinc/meaninc))
     )
   }, simplify = FALSE) %>%
     bind_rows

### 2020-03-22 reconstruct_time_series_gyeonggi_raw.R

@@ -101,7 +101,7 @@ for (i in 1:length(ee$b_Intercept)) {
   ## pgamma(4.2, 25, 25/2.6) - pgamma(1.8, 25, 25/2.6)
   sdinc <- rgamma(1, 25, 25/2.6)
   
-  shape <- meaninc^2/sdinc^2
+  shapeinc <- meaninc^2/sdinc^2
   
   meandelay <- estmean[,i]
   shape <- ee$shape[i]
@@ -110,7 +110,7 @@ for (i in 1:length(ee$b_Intercept)) {
     data_frame(
       confirm=as.Date("2020-01-20")+x-1,
       onset=confirm-rnbinom(round(cases[x]), mu=meandelay[x], size=shape),
-      infection=onset-floor(rgamma(round(cases[x]), shape=shape, rate=shape/meaninc))
+      infection=onset-floor(rgamma(round(cases[x]), shape=shapeinc, rate=shapeinc/meaninc))
     )
   }, simplify = FALSE) %>%
     bind_rows

### 2020-03-22 reconstruct_time_series_seoul_detect.R

@@ -105,7 +105,7 @@ for (i in 1:length(ee$b_Intercept)) {
   ## pgamma(4.2, 25, 25/2.6) - pgamma(1.8, 25, 25/2.6)
   sdinc <- rgamma(1, 25, 25/2.6)
   
-  shape <- meaninc^2/sdinc^2
+  shapeinc <- meaninc^2/sdinc^2
   
   meandelay <- estmean[,i]
   shape <- ee$shape[i]
@@ -114,7 +114,7 @@ for (i in 1:length(ee$b_Intercept)) {
     data_frame(
       confirm=as.Date("2020-01-20")+x-1,
       onset=confirm-rnbinom(cases[x], mu=meandelay[x], size=shape),
-      infection=onset-floor(rgamma(cases[x], shape=shape, rate=shape/meaninc))
+      infection=onset-floor(rgamma(cases[x], shape=shapeinc, rate=shapeinc/meaninc))
     )
   }, simplify = FALSE) %>%
     bind_rows

### 2020-03-22 reconstruct_time_series_seoul_raw.R

@@ -100,7 +100,7 @@ for (i in 1:length(ee$b_Intercept)) {
   ## pgamma(4.2, 25, 25/2.6) - pgamma(1.8, 25, 25/2.6)
   sdinc <- rgamma(1, 25, 25/2.6)
   
-  shape <- meaninc^2/sdinc^2
+  shapeinc <- meaninc^2/sdinc^2
   
   meandelay <- estmean[,i]
   shape <- ee$shape[i]
@@ -109,7 +109,7 @@ for (i in 1:length(ee$b_Intercept)) {
     data_frame(
       confirm=as.Date("2020-01-20")+x-1,
       onset=confirm-rnbinom(cases[x], mu=meandelay[x], size=shape),
-      infection=onset-floor(rgamma(cases[x], shape=shape, rate=shape/meaninc))
+      infection=onset-floor(rgamma(cases[x], shape=shapeinc, rate=shapeinc/meaninc))
     )
   }, simplify = FALSE) %>%
     bind_rows

### 2020-03-22 figure_R_t_daegu.R
typo

@@ -49,7 +49,7 @@ rt_all <- bind_rows(rt_censor, rt_censor_detect, rt_censor_raw) %>%
                                 "adjusted for detection rate + number of tests"))
   )
 
-geo <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA", sheet=3)
+geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3)
 
 geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
 

### 2020-03-22 figure_compare_R_t.R

@@ -29,7 +29,7 @@ seoul <- geo %>%
     cases=sum(Seoul, na.rm=TRUE)
   )
 
-traffic_daegu1 <- select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) %>%
+traffic_daegu1 <- dplyr::select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) %>%
   mutate(
     wday=wday(date)
   ) %>%
@@ -37,7 +37,7 @@ traffic_daegu1 <- select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) 
     date >= as.Date("2017-01-20")+3, date <= as.Date("2017-03-01")+3
   )
 
-traffic_daegu2 <- select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) %>%
+traffic_daegu2 <- dplyr::select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) %>%
   mutate(
     wday=wday(date)
   ) %>%
@@ -45,7 +45,7 @@ traffic_daegu2 <- select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) 
     date >= as.Date("2018-01-20")+2, date <= as.Date("2018-03-01")+2
   )
 
-traffic_daegu3 <- select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) %>%
+traffic_daegu3 <- dplyr::select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) %>%
   mutate(
     wday=wday(date)
   ) %>%
@@ -53,7 +53,7 @@ traffic_daegu3 <- select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) 
     date >= as.Date("2019-01-20")+1, date <= as.Date("2019-03-01")+1
   )
 
-traffic_daegu4 <- select(ungroup(filter(traffic_daegu,year==2020)), -월, -일) %>%
+traffic_daegu4 <- dplyr::select(ungroup(filter(traffic_daegu,year==2020)), -월, -일) %>%
   mutate(
     wday=wday(date)
   ) %>%

### 2020-03-22 korea.tex
typo
quick review followed by quick meeting with Daniel

@@
 The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
-The epidemic then took off in mid-February when COVID-19 began to spread within a church from a city of Daegu --- as of March 18, 2020, 8,413 cases have been confirmed, of which 60\% are related to the church \citep{kcdc}.
-While its success in suppressing the epidemic has been widely attributed to its intensive testing \citep{fergusonimpact, science}, other factors, such as social distancing, are also likely to have played important roles.
+The epidemic then took off in mid-February when COVID-19 began to spread within a church in the city of Daegu --- as of March 18, 2020, 8,413 cases were confirmed, of which 60\% were related to the church \citep{kcdc}.
+While Korea's success in suppressing the epidemic has been widely attributed to intensive testing \citep{fergusonimpact, science}, other factors, such as social distancing, are also likely to have played important roles.
 Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by using metro traffic data to compare epidemics in two geographically separated major cities in South Korea.
@@
 
-We analyzed epidemiological data, collected bewteen January 20--March 16, 2020, describing the COVID-19 outbreak in South Korea.
+We analyzed epidemiological data, collected between January 20--March 16, 2020, describing the COVID-19 outbreak in South Korea.
 Daily number of reported cases in each geographic region was transcribed from press releases by Korea Centers for Disease Control and Prevention (KCDC) \cite{kcdc}.
@@
 Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from \url{data.go.kr} and \url{data.seoul.go.kr}, respectively.
-We compared the the daily number of individuals who got on the subway --- including a monorail in Daegu --- across all stations (\fref{data}).
-Soon after the first church-related case was confirmed in Daegu on Feb 18, 2020, the daily traffic volume decreased by about 80\% and 50\% compared to previous years in Daegu and Seoul, respectively.
+We tabulated the total number of individuals who got on the subway or monorail (\fref{data}).
+Soon after the first church-related case was confirmed in Daegu on Feb 18, 2020, the daily traffic volume decreased by about 80\% and 50\% in Daegu and Seoul, respectively.
+\jd{Cut a lot of words here; I think it's fine, but see wyt.}
 
@@
 
+We are assuming that during the early period of an epidemic if we follow a consistent protocol of identifying and deciding whom to test, the positivity should stay about the same and that's the justification for the first correction.
+
+\jd{Flow needs to be cleaned up here, but I'm not letting myself take the time.}
 In order to estimate the time-dependent reproduction number $R_t$ (i.e., the expected number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), 
 we first accounted for changes in testing criteria (\fref{dist}B) by multiplying the daily number of reported cases by the relative detection rate for each criterion (i.e., a proportion of cases tested positive based on a criterion divided by the mean detection rate).
+\jd{What do you mean by detection rate?}
 Likewise, we accounted for changes in the number of tests completed on each day (\fref{dist}A) by dividing the scaled number of reported cases by the relative number of tests completed on each day; this step was performed separately for each testing criterion because widening of criteria necessarily increases the number of tests.
+The number of negative cases was not reported on January 25 and 31, 2020; we took the average of cumulative negative cases from one day before and after these dates instead to impute missing values.
+\jd{It seems like it would be better to use the 
 A sensitivity analysis showed that qualitative patterns of inferred $R_t$ are robust to these adjustments (Supplementary Materials).
@@
 On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}B);
-while social distancing is likely to have assisted in mitigating the spread in Seoul, our analysis suggests that 50\% decrease in metro traffic was insufficient to suppress the spread.
+our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less effective.
 Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.
@@
 The ongoing COVID-19 outbreak in South Korea provides a unique perspective to understanding and controlling the pandemic.
-Its experience provides evidence that the epidemic can be suppressed without draconian measures as in China \citep{kickbusch2020response}.
+Its experience provides evidence that the epidemic can be suppressed without less extreme measures than those take by China \citep{kickbusch2020response}.
 Our analysis reveals potential roles of social distancing in mitigating the COVID-19 epidemic in South Korea.
@@
 Our analysis focused on comparing metro traffic, which serves as a proxy for the degree of social distancing, with epidemiological dynamics in two cities.
-However, our analysis does not allow us to measure the direct ``effect'' of social distancing on epidemiological dynamics.
+This analysis is indirect; we are not able to directly estimate the effect of social distancing on epidemiological dynamics.
 Moreover, other measures, such as intensive testing of core transmission groups and school closure, are also likely to have affected the changes in $\mathcal R_t$ \citep{kcdc}.
@@
 
-Finally, our studies highlights the importance of characterizing geographical heterogeneity in estimating epidemic potential.
+Finally, our studies highlights the importance of considering geographical heterogeneity in estimating epidemic potential.
 The recent decrease in the number of reported cases in South Korea is likely to be strongly correlated with the decrease in the number of cases in Daegu.

### 2020-03-23 korea.tex
tweaks

@@
 Soon after the first church-related case was confirmed in Daegu on Feb 18, 2020, the daily traffic volume decreased by about 80\% and 50\% in Daegu and Seoul, respectively.
-\jd{Cut a lot of words here; I think it's fine, but see wyt.}
 
@@
 
-We are assuming that during the early period of an epidemic if we follow a consistent protocol of identifying and deciding whom to test, the positivity should stay about the same and that's the justification for the first correction.
-
-\jd{Flow needs to be cleaned up here, but I'm not letting myself take the time.}
-In order to estimate the time-dependent reproduction number $R_t$ (i.e., the expected number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), 
-we first accounted for changes in testing criteria (\fref{dist}B) by multiplying the daily number of reported cases by the relative detection rate for each criterion (i.e., a proportion of cases tested positive based on a criterion divided by the mean detection rate).
-\jd{What do you mean by detection rate?}
+In order to estimate the time-dependent reproduction number $R_t$ (i.e., the expected number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), we reconstructed the incidence time series (i.e., the number of infected individuals on each day).
+First, we accounted for changes in testing criteria, which occurred 4 times between January 20--March 16, 2020, by assuming that the proportion positive should remain roughly constant if we follow a consistent protocol of identifying and deciding whom to test.
+To do so, we calculated the relative proportion of positive cases during each period (divided by the the between-period mean) and multiplied the daily number of reported cases by the relative proportions of the corresponding criterion.
 Likewise, we accounted for changes in the number of tests completed on each day (\fref{dist}A) by dividing the scaled number of reported cases by the relative number of tests completed on each day; this step was performed separately for each testing criterion because widening of criteria necessarily increases the number of tests.
 The number of negative cases was not reported on January 25 and 31, 2020; we took the average of cumulative negative cases from one day before and after these dates instead to impute missing values.
-\jd{It seems like it would be better to use the 
 A sensitivity analysis showed that qualitative patterns of inferred $R_t$ are robust to these adjustments (Supplementary Materials).
@@
 
-For each posterior sample of the reconstructed incidence time series (i.e., the number of infected individuals on day $t$), we divided incidence by $1-\textrm{degree of censoring}$ and estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
+For each posterior sample of the reconstructed incidence time series, we divided incidence by $1-\textrm{degree of censoring}$ and estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
 \begin{equation}

### 2020-03-23 korea.tex
more tweaks

@@
 Likewise, we estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using the same negative-binomial regression model, while accounting for right-censoring (\fref{dist}D),
-to calculate the degree of censoring (i.e., median probability that a case infected on a given day will be reported after March 16, 2020).
+to calculate the median probability that a case infected on a given day will be reported before March 16, 2020.
 Details are provided in the Supplementary Materials.
@@
 Negative binomial distributions are parameterized using its mean and dispersion.
-Priors are chosen such that the 95\% quantiles of prior means and standard deviations are consistent with previous estiamtes.
+Priors are chosen such that the 95\% quantiles of prior means and standard deviations are consistent with previous estimates.
 }
@@
 
-For each posterior sample of the reconstructed incidence time series, we divided incidence by $1-\textrm{degree of censoring}$ and estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
+For each posterior sample of the reconstructed incidence time series, we divided daily incidence by the probability that they would have been reported before March 16, 2020, and estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
 \begin{equation}
@@
 The time-dependent mean of the negative binomial distribution is modeled using splines.
-We assumed weakly informative priors on the fixed effects: normal distributions with mean of 0 and standard devition of 2;
+We assumed weakly informative priors on the fixed effects: normal distributions with mean of 0 and standard deviation of 2;
 note that these distributions are priors on link scale.

### 2020-03-23 korea.tex
k im gonna send it off

@@
 Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.
-Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces, providing support for the robustness of our analysis (Supplementary Materials).
+Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces (Gyeongsangbuk-do and Gyeonggi-do), providing support for the robustness of our analysis (Supplementary Materials).
 
@@
 The ongoing COVID-19 outbreak in South Korea provides a unique perspective to understanding and controlling the pandemic.
-Its experience provides evidence that the epidemic can be suppressed without less extreme measures than those take by China \citep{kickbusch2020response}.
+Its experience provides evidence that the epidemic can be suppressed with less extreme measures than those take by China \citep{kickbusch2020response}.
 Our analysis reveals potential roles of social distancing in mitigating the COVID-19 epidemic in South Korea.
@@
 Our study is not without limitations.
-In particular, we did not account for differences in the delay distributions or changes in the number of tests or detection rates among cities.
+In particular, we did not account for differences in the delay distributions or changes in the number of tests among cities.
 The intensity of intervention is likely to vary across regions given that majority of COVID-19 cases in South Korea were reported from Daegu.
@@
 
-Finally, our studies highlights the importance of considering geographical heterogeneity in estimating epidemic potential.
+Finally, our study highlights the importance of considering geographical heterogeneity in estimating epidemic potential.
 The recent decrease in the number of reported cases in South Korea is likely to be strongly correlated with the decrease in the number of cases in Daegu.

### 2020-03-23 korea.tex
tweaks

@@
 Kaiyuan Sun\textsuperscript{2}
-C\'ecile Viboud\textsuperscript{2}
-Jonathan Dushoff\textsuperscript{3,4,5}
-Bryan T.\ Grenfell\textsuperscript{1,2,6}
+Jonathan Dushoff\textsuperscript{3,4,5,$\dagger$}
+Bryan T.\ Grenfell\textsuperscript{1,2,6,$\dagger$}
+C\'ecile Viboud\textsuperscript{2,$\dagger$}
 \\
@@
 
+$\dagger$These authors contributed equally\\
 *Corresponding author: swp2@princeton.edu
@@
 The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
-The epidemic then took off in mid-February when COVID-19 began to spread within a church in the city of Daegu --- as of March 18, 2020, 8,413 cases were confirmed, of which 60\% were related to the church \citep{kcdc}.
-While Korea's success in suppressing the epidemic has been widely attributed to intensive testing \citep{fergusonimpact, science}, other factors, such as social distancing, are also likely to have played important roles.
+The epidemic then took off in mid-February when COVID-19 began to spread within a church in the city of Daegu.
+The chains of transmission that began from this cluster distinguish the epidemic in South Korea from that in any other countries:
+As of March 18, 2020, 8,413 cases were confirmed, of which 60\% were related to the church and 27\% were in their 20's \citep{kcdc}.
+Nonetheless, South Korea's intensive testing using novel tracing techniques allowed it to rapidly identify cases and reduce the spread \citep{fergusonimpact, tracingkorea, science}.
 Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by using metro traffic data to compare epidemics in two geographically separated major cities in South Korea.
@@
 
-\begin{figure}
+\begin{figure}[!ht]
 \includegraphics[width=\textwidth]{figure_report_delay.pdf}
@@
 
-We then estimated time-dependent \emph{backward} confirmation-to-onset delay distributions from the partial line list using a negative-binomial regression with \texttt{brms} package \citep{burkner2017brms} (\fref{dist}C).
+We then estimated time-dependent \emph{backward} confirmation-to-onset delay distributions from the partial line list using a negative-binomial regression with the \texttt{brms} package \citep{burkner2017brms} (\fref{dist}C).
 We combined the estimated posterior distribution of delay distributions with previously estimated incubation period distribution (Table 1) to obtain posterior samples for date of infection for each confirmed case.
@@
 to calculate the median probability that a case infected on a given day will be reported before March 16, 2020.
+We note that changes in the mean decrease delay (\fref{dist}C) reflect changes in incidence of symptomatic cases, whereas the decrease in the mean forward delay (\fref{dist}D) reflects the increase in the intensity of testing.
 Details are provided in the Supplementary Materials.
@@
 Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
+The initial decrease in $\mathcal R_t$ may reflect the possibility that a large proportion of church-related transmission could have occurred much earlier; according to the KCDC press release, the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.
+
 On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}B);
@@
 Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.
+While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.49$; 95\% CI: 0.20--0.70), the strength of correlations are likely to be strongly conflated because they do not account for the effects of any other measures that could have affected $\mathcal R_t$.
 Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces (Gyeongsangbuk-do and Gyeonggi-do), providing support for the robustness of our analysis (Supplementary Materials).

### 2020-03-23 figure_compare_R_t.R

@@ -123,6 +123,13 @@ seoul_traffic <- data.frame(
   traffic=traffic_seoul4$total/((traffic_seoul1$total + traffic_seoul2$total + traffic_seoul3$total)/3)
 )
 
+daegu_merge <- merge(rt_daegu, daegu_traffic)
+cor.test(daegu_merge$median, daegu_merge$traffic)
+
+seoul_merge <- merge(rt_seoul, seoul_traffic)
+cor.test(seoul_merge$median, seoul_merge$traffic)
+
+
 g1 <- ggplot(rt_daegu) +
   geom_bar(data=daegu, aes(date_report, cases/max(daegu$cases)), stat="identity", alpha=0.3) + 
   geom_hline(yintercept=6, lty=2, col=2) + 

### 2020-03-23 korea.tex
tweak

@@
 to calculate the median probability that a case infected on a given day will be reported before March 16, 2020.
-We note that changes in the mean decrease delay (\fref{dist}C) reflect changes in incidence of symptomatic cases, whereas the decrease in the mean forward delay (\fref{dist}D) reflects the increase in the intensity of testing.
+We note that changes in the mean backward delay (\fref{dist}C) reflect changes in incidence of symptomatic cases, whereas the decrease in the mean forward delay (\fref{dist}D) reflects the increase in the intensity of testing.
 Details are provided in the Supplementary Materials.

### 2020-03-23 korea.tex
done

@@
 Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.
-While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.49$; 95\% CI: 0.20--0.70), the strength of correlations are likely to be strongly conflated because they do not account for the effects of any other measures that could have affected $\mathcal R_t$.
+While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.49$; 95\% CI: 0.20--0.70), the strength of correlations are likely to be strongly conflated because they do not account for the effects of any other measures that could have affected $\mathcal R_t$;
+we do not find any signatures of lags (Supplementary Materials).
 Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces (Gyeongsangbuk-do and Gyeonggi-do), providing support for the robustness of our analysis (Supplementary Materials).
@@
 
+\pagebreak
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figure_cross.pdf}
+\caption{
+\textbf{Cross correlation between the normalized traffic volume and the median estimates of $\mathcal R_t$.}
+}
+\end{figure}
+
 \end{document}

### 2020-03-24 korea.tex
still working on it...

@@
 Kaiyuan Sun\textsuperscript{2}
-Jonathan Dushoff\textsuperscript{3,4,5,$\dagger$}
-Bryan T.\ Grenfell\textsuperscript{1,2,6,$\dagger$}
 C\'ecile Viboud\textsuperscript{2,$\dagger$}
+Bryan T.\ Grenfell\textsuperscript{1,2,3,$\dagger$}
+Jonathan Dushoff\textsuperscript{4,5,6,$\dagger$}
 \\
@@
 \\
-\textbf{3} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
+\textbf{3} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
 \\
-\textbf{4} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
+\textbf{4} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
 \\
-\textbf{5} Department of Biology, McMaster University, Hamilton, ON, Canada
+\textbf{5} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
 \\
-\textbf{6} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
+\textbf{6} Department of Biology, McMaster University, Hamilton, ON, Canada
 \\
@@
 $\dagger$These authors contributed equally\\
-*Corresponding author: swp2@princeton.edu
+*Corresponding author: swp2@princeton.edu\\
+This article does not necessarily represent the views of the NIH or the US government.
 \end{flushleft}
@@
 As of March 18, 2020, 8,413 cases were confirmed, of which 60\% were related to the church and 27\% were in their 20's \citep{kcdc}.
-Nonetheless, South Korea's intensive testing using novel tracing techniques allowed it to rapidly identify cases and reduce the spread \citep{fergusonimpact, tracingkorea, science}.
+Nonetheless, South Korea's intensive testing using novel contact tracing techniques allowed rapid identification and isolation cases and reduction of onward transmission \citep{fergusonimpact, tracingkorea, science}.
 Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by using metro traffic data to compare epidemics in two geographically separated major cities in South Korea.
@@
 
-We analyzed epidemiological data, collected between January 20--March 16, 2020, describing the COVID-19 outbreak in South Korea.
+We analyzed epidemiological data describing the COVID-19 outbreak in South Korea between January 20--March 16, 2020.
 Daily number of reported cases in each geographic region was transcribed from press releases by Korea Centers for Disease Control and Prevention (KCDC) \cite{kcdc}.
@@
 We compared epidemiological dynamics of COVID-19 from two cities in which the biggest number of COVID-19 cases have been reported: Daegu and Seoul.
-Between January 20--March 16, 2020, 6,083 cases were reported from Daegu and 248 from Seoul.
+Between January 20--March 16, 2020, 6,083 cases from Daegu and 248 from Seoul were reported by the KCDC (Supplementary Materials).
 Unlike the epidemic in Daegu, which is characterized by a single, large peak followed by a gradual decrease, the epidemic in Seoul consists of several small outbreaks (\fref{data}).
@@
 Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from \url{data.go.kr} and \url{data.seoul.go.kr}, respectively.
-We tabulated the total number of individuals who got on the subway or monorail (\fref{data}).
+We tabulated the total number of individuals who accessed subway or monorail (\fref{data});
+we restricted our analysis to subway lines 1--9 in Seoul.
 Soon after the first church-related case was confirmed in Daegu on Feb 18, 2020, the daily traffic volume decreased by about 80\% and 50\% in Daegu and Seoul, respectively.
@@
 The ongoing COVID-19 outbreak in South Korea provides a unique perspective to understanding and controlling the pandemic.
-Its experience provides evidence that the epidemic can be suppressed with less extreme measures than those take by China \citep{kickbusch2020response}.
+Its experience provides evidence that the epidemic can be suppressed with less extreme measures than those taken by China \citep{kickbusch2020response}.
 Our analysis reveals potential roles of social distancing in mitigating the COVID-19 epidemic in South Korea.
@@
 We did not have the data to account for these factors.
+We also did not distinguish the differences between local and imported cases, which may overestimate $\mathcal R_t$.
 Nonetheless, the robustness of our findings is supported by the sensitivity analyses.
@@
 \section*{Supplementary Materials}
+
+\subsection*{Epidemiological data}
+
+The daily number of reported cases from each region was translated and trascribed from the KCDC press release.
+
+;
+the reported number of cases by each city's government may be different as some cases may be transferred after they are confirmed.
 

### 2020-03-24 figure_compare_report.R

@@ -152,4 +152,4 @@ g2 <- ggplot(filter(traffic_seoul_all, year==2020)) +
 
 gtot <- arrangeGrob(g1, g2, nrow=1, widths=c(1.1, 1))
 
-ggsave("figure_compare_report.pdf", gtot, width=8, height=4)
+ggsave("figure_compare_report.pdf", gtot, width=8, height=3.5)

### 2020-03-24 report_delay_backward.R

@@ -5,7 +5,7 @@ library(lubridate)
 library(brms)
 library(rstan)
 
-covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA") %>%
   mutate(
     age=as.numeric(age),
     date_onset=as.Date(date_onset)
@@ -13,7 +13,7 @@ covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
 
 casenum <- covid_line$case[!is.na(covid_line$case)]
 
-covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=1, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -21,7 +21,7 @@ covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1,
     !(case %in% casenum)
   )
 
-covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=2, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -29,7 +29,7 @@ covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet
     !(case %in% casenum)
   )
 
-covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=3, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -37,7 +37,7 @@ covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3,
     !(case %in% casenum)
   )
 
[…]

### 2020-03-24 report_delay_censor.R

@@ -5,7 +5,7 @@ library(lubridate)
 library(brms)
 library(rstan)
 
-covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA") %>%
   mutate(
     age=as.numeric(age),
     date_onset=as.Date(date_onset)
@@ -13,7 +13,7 @@ covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
 
 casenum <- covid_line$case[!is.na(covid_line$case)]
 
-covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=1, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -21,7 +21,7 @@ covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1,
     !(case %in% casenum)
   )
 
-covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=2, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -29,7 +29,7 @@ covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet
     !(case %in% casenum)
   )
 
-covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=3, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -37,7 +37,7 @@ covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3,
     !(case %in% casenum)
   )
 
[…]

### 2020-03-24 traffic_seoul.R

@@ -5,7 +5,7 @@ library(ggplot2); theme_set(theme_bw())
 library(readxl)
 
 whichfilter <- c("1호선", "2호선", "3호선", "4호선", "5호선", "6호선", "7호선", "8호선", "9호선",
-                 "9호선2~3단계", "우이신설선", "9호선2단계")
+                 "9호선2~3단계", "9호선2단계")
 
 ss2020 <- read.csv("seoul_subway/서울시 지하철호선별 역별 승하차 인원 정보.csv", fileEncoding="EUC-KR") 
 

### 2020-03-24 korea.tex
k

@@
 The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
-The epidemic then took off in mid-February when COVID-19 began to spread within a church in the city of Daegu.
+In February, the disease spread rapidly within a church in the city of Daegu.
 The chains of transmission that began from this cluster distinguish the epidemic in South Korea from that in any other countries:
@@
 In order to estimate the time-dependent reproduction number $R_t$ (i.e., the expected number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), we reconstructed the incidence time series (i.e., the number of infected individuals on each day).
-First, we accounted for changes in testing criteria, which occurred 4 times between January 20--March 16, 2020, by assuming that the proportion positive should remain roughly constant if we follow a consistent protocol of identifying and deciding whom to test.
-To do so, we calculated the relative proportion of positive cases during each period (divided by the the between-period mean) and multiplied the daily number of reported cases by the relative proportions of the corresponding criterion.
-Likewise, we accounted for changes in the number of tests completed on each day (\fref{dist}A) by dividing the scaled number of reported cases by the relative number of tests completed on each day; this step was performed separately for each testing criterion because widening of criteria necessarily increases the number of tests.
-The number of negative cases was not reported on January 25 and 31, 2020; we took the average of cumulative negative cases from one day before and after these dates instead to impute missing values.
-A sensitivity analysis showed that qualitative patterns of inferred $R_t$ are robust to these adjustments (Supplementary Materials).
+We first adjusted the number of reported cases to account for changes in testing criteria (\fref{dist}), which occurred 4 times between January 20--March 16, 2020, as well as changes in the number of tests (Supplementary Materials).
+A sensitivity analysis showed that qualitative patterns of inferred $R_t$ are robust to these adjustments.
 
@@
 where $I_t$ is the censoring-adjusted incidence time series and $w_k$ is the generation-interval distribution randomly drawn from a prior distribution (Table 1).
+We weighted each sample of $\mathcal R_t$ by a gamma probability distribution with a mean of 2.6 and a standard deviation of 2 to reflect prior knowledge \citep{tempvar} and took weighted quantiles to calculate the medians and associated 95\% credible intervals.
 We restricted the calculation of $\mathcal R_t$ between February 2, 2020 (14 days after the first confirmed case was imported) and March 10, 2020 (because the degree of censoring would be too large after this date to estimate $\mathcal R_t$).
@@
 Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.
-While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.49$; 95\% CI: 0.20--0.70), the strength of correlations are likely to be strongly conflated because they do not account for the effects of any other measures that could have affected $\mathcal R_t$;
+While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.55$; 95\% CI: 0.27--0.74), the strength of correlations are likely to be strongly conflated because they do not account for the effects of any other measures that could have affected $\mathcal R_t$;
 we do not find any signatures of lags (Supplementary Materials).
@@
 We did not have the data to account for these factors.
-We also did not distinguish the differences between local and imported cases, which may overestimate $\mathcal R_t$.
+We also did not distinguish local and imported cases, which may overestimate $\mathcal R_t$.
 Nonetheless, the robustness of our findings is supported by the sensitivity analyses.
@@
 
-The daily number of reported cases from each region was translated and trascribed from the KCDC press release.
+The daily number of reported cases from each region was translated and transcribed from the KCDC press release \citep{kcdc}.
+Following the KCDC's protocol, the daily number of reported cases prior to February 20, 2020, reflects the number of confirmed cases on each day.
+Between February 21 -- March 1, 2020, the daily number of reported cases reflects the number of reported cases within the last 24 hours (9 AM to 9 AM).
+On March 2, 2020, the daily number of reported cases reflects the number of cases that were reported between 9 AM March 1, 2020, and 12 AM March 2, 2020.
+Since then, the daily number of reported cases reflects the number of reported cases within the last 24 hours (12 AM to 12 AM).
+The daily number of reported cases by the KCDC may be different from the reports by each city's government as some cases may be transferred after they are confirmed.
 
-;
-the reported number of cases by each city's government may be different as some cases may be transferred after they are confirmed.
+\subsection*{Time series adjustments}
+
+First, we accounted for changes in testing criteria, which occurred 4 times between January 20--March 16, 2020, by assuming that the proportion positive should remain roughly constant if we follow a consistent protocol of identifying and deciding whom to test.
+To do so, we calculated the relative proportion of positive cases during each period (divided by the the between-period mean) and multiplied the daily number of reported cases by the relative proportions of the corresponding criterion.
+Likewise, we accounted for changes in the number of tests completed on each day by dividing the scaled number of reported cases by the relative number of tests completed on each day; this step was performed separately for each testing criterion because widening of criteria necessarily increases the number of tests.
+The number of negative cases was not reported on January 25 and 31, 2020; we took the average of cumulative negative cases from one day before and after these dates instead to impute missing values.
 

### 2020-03-24 figure_R_t_daegu.R

@@ -1,18 +1,22 @@
 library(readxl)
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
+source("wquant.R")
 
 load("R_t_daegu_censor.rda")
 load("R_t_daegu_censor_detect.rda")
 load("R_t_daegu_censor_raw.rda")
 
+R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
+
 rt_censor <- R_t_daegu_censor %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE)
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
+    upr=wquant(IRt, weights=weight,0.975)
   ) %>%
   mutate(
     group="adjusted for detection rate + number of tests"
@@ -21,10 +25,11 @@ rt_censor <- R_t_daegu_censor %>%
 rt_censor_detect <- R_t_daegu_censor_detect %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE)
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
+    upr=wquant(IRt, weights=weight,0.975)
   ) %>%
   mutate(
     group="adjusted for detection rate"
@@ -33,10 +38,11 @@ rt_censor_detect <- R_t_daegu_censor_detect %>%
 rt_censor_raw <- R_t_daegu_censor_raw %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE)
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
[…]

### 2020-03-24 figure_R_t_gyeongbuk.R

@@ -1,17 +1,22 @@
+library(readxl)
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
+source("wquant.R")
 
 load("R_t_gyeongbuk_censor.rda")
 load("R_t_gyeongbuk_censor_detect.rda")
 load("R_t_gyeongbuk_censor_raw.rda")
 
+R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
+
 rt_censor <- R_t_gyeongbuk_censor %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE)
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
+    upr=wquant(IRt, weights=weight,0.975)
   ) %>%
   mutate(
     group="adjusted for detection rate + number of tests"
@@ -20,10 +25,11 @@ rt_censor <- R_t_gyeongbuk_censor %>%
 rt_censor_detect <- R_t_gyeongbuk_censor_detect %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE)
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
+    upr=wquant(IRt, weights=weight,0.975)
   ) %>%
   mutate(
     group="adjusted for detection rate"
@@ -32,10 +38,11 @@ rt_censor_detect <- R_t_gyeongbuk_censor_detect %>%
 rt_censor_raw <- R_t_gyeongbuk_censor_raw %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
[…]

### 2020-03-24 figure_R_t_gyeonggi.R

@@ -1,17 +1,22 @@
+library(readxl)
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
+source("wquant.R")
 
 load("R_t_gyeonggi_censor.rda")
 load("R_t_gyeonggi_censor_detect.rda")
 load("R_t_gyeonggi_censor_raw.rda")
 
+R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
+
 rt_censor <- R_t_gyeonggi_censor %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE)
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
+    upr=wquant(IRt, weights=weight,0.975)
   ) %>%
   mutate(
     group="adjusted for detection rate + number of tests"
@@ -20,10 +25,11 @@ rt_censor <- R_t_gyeonggi_censor %>%
 rt_censor_detect <- R_t_gyeonggi_censor_detect %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE)
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
+    upr=wquant(IRt, weights=weight,0.975)
   ) %>%
   mutate(
     group="adjusted for detection rate"
@@ -32,10 +38,11 @@ rt_censor_detect <- R_t_gyeonggi_censor_detect %>%
 rt_censor_raw <- R_t_gyeonggi_censor_raw %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE)
+    median=wquant(IRt, weights=weight,0.5),
[…]

### 2020-03-24 figure_cross.R

@@ -4,14 +4,11 @@ library(lubridate)
 library(dplyr)
 library(tidyr)
 library(readxl)
-source("wquant.R")
 
 load("R_t_daegu_censor.rda")
 load("R_t_seoul_censor.rda")
 load("traffic_daegu.rda")
 load("traffic_seoul.rda")
-
-R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
 
 geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
   mutate(
@@ -67,11 +64,10 @@ traffic_daegu4 <- dplyr::select(ungroup(filter(traffic_daegu,year==2020)), -월,
 rt_daegu <- R_t_daegu_censor %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
-  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=wquant(IRt, weights=weight,0.5),
-    lwr=wquant(IRt, weights=weight,0.025),
-    upr=wquant(IRt, weights=weight,0.975),
+    median=median(IRt, na.rm=TRUE),
+    lwr=quantile(IRt, 0.025, na.rm=TRUE),
+    upr=quantile(IRt, 0.975, na.rm=TRUE),
     location="Daegu"
   )
 
@@ -115,11 +111,10 @@ traffic_seoul4 <- ungroup(filter(traffic_seoul,year==2020)) %>%
 rt_seoul <- R_t_seoul_censor %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
-  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=wquant(IRt, weights=weight,0.5),
-    lwr=wquant(IRt, weights=weight,0.025),
-    upr=wquant(IRt, weights=weight,0.975),
+    median=median(IRt, na.rm=TRUE),
+    lwr=quantile(IRt, 0.025, na.rm=TRUE),
+    upr=quantile(IRt, 0.975, na.rm=TRUE),
     location="Seoul"
   )
 
@@ -128,62 +123,50 @@ seoul_traffic <- data.frame(
   traffic=traffic_seoul4$total/((traffic_seoul1$total + traffic_seoul2$total + traffic_seoul3$total)/3)
 )
 
-daegu_merge <- merge(rt_daegu, daegu_traffic)
-cor.test(daegu_merge$median, daegu_merge$traffic)
+c1 <- ccf(
[…]

### 2020-03-24 figure_report_delay.R

@@ -15,7 +15,7 @@ covid_test <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=2)
 covid_test$negative[7] <- (covid_test$negative[6]+covid_test$negative[8])/2
 covid_test$negative[13] <- (covid_test$negative[12]+covid_test$negative[14])/2
 
-covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
+covid_line <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA") %>%
   mutate(
     age=as.numeric(age),
     date_onset=as.Date(date_onset)
@@ -23,7 +23,7 @@ covid_line <- read_xlsx("data/COVID19-Korea-2020-03-13.xlsx", na="NA") %>%
 
 casenum <- covid_line$case[!is.na(covid_line$case)]
 
-covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=1, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -31,7 +31,7 @@ covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1,
     !(case %in% casenum)
   )
 
-covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=2, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -39,7 +39,7 @@ covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet
     !(case %in% casenum)
   )
 
-covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=3, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
[…]

### 2020-03-24 reconstruct_time_series_gyeongbuk_detect.R

@@ -16,7 +16,7 @@ covid_line <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA") %>%
 
 casenum <- covid_line$case[!is.na(covid_line$case)]
 
-covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1, na="NA") %>%
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=1, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -24,7 +24,7 @@ covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=1,
     !(case %in% casenum)
   )
 
-covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=2, na="NA") %>%
+covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=2, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -32,7 +32,7 @@ covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet
     !(case %in% casenum)
   )
 
-covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3, na="NA") %>%
+covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=3, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -40,7 +40,7 @@ covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=3,
     !(case %in% casenum)
   )
 
-covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-13.xlsx", sheet=4, na="NA") %>%
+covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=4, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -57,7 +57,7 @@ covid_delay <- covid_all %>%
     day=yday(date_confirm),
     day=day-min(day)
[…]

### 2020-03-24 korea.tex
getting a bit too long!!!

@@
 to calculate the median probability that a case infected on a given day will be reported before March 16, 2020.
-We note that changes in the mean backward delay (\fref{dist}C) reflect changes in incidence of symptomatic cases, whereas the decrease in the mean forward delay (\fref{dist}D) reflects the increase in the intensity of testing.
-Details are provided in the Supplementary Materials.
+Changes in the mean backward delay (\fref{dist}C) reflect changes in incidence of symptomatic cases, whereas the decrease in the mean forward delay (\fref{dist}D) reflects the improvement in the accuracy of case identification \citep{lai2020effect}.
+Implementation details are provided in the Supplementary Materials.
 
@@
 While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.55$; 95\% CI: 0.27--0.74), the strength of correlations are likely to be strongly conflated because they do not account for the effects of any other measures that could have affected $\mathcal R_t$;
-we do not find any signatures of lags (Supplementary Materials).
+in particular, the increase in the number of tests (\fref{dist}A) as well as the improvement in the accuracy of case identification (\fref{dist}D) are likely to have played important roles in reducing $\mathcal R_t$.
+We do not find any signatures of lags (Supplementary Materials).
 Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces (Gyeongsangbuk-do and Gyeonggi-do), providing support for the robustness of our analysis (Supplementary Materials).
@@
 We did not have the data to account for these factors.
-We also did not distinguish local and imported cases, which may overestimate $\mathcal R_t$.
 Nonetheless, the robustness of our findings is supported by the sensitivity analyses.
+We also did not distinguish local and imported cases, which may overestimate $\mathcal R_t$ \citep{thompson2019improved}.
 
@@
 Our analysis reveals that the epidemic may still persist in other regions, including Seoul and Gyeonggi-do.
+Persistent reportings of COVID-19 cases in Seoul and Gyeonggi-do (around 10 cases almost everyday) between March 11--22, 2020, provide further support for our analysis \citep{kcdc}.
 Unless the reproduction number can be reduced below 1 in all regions, small outbreaks may continue to occur in South Korea.

### 2020-03-24 korea.tex
k done for tonight

@@
 Kaiyuan Sun\textsuperscript{2}
-C\'ecile Viboud\textsuperscript{2,$\dagger$}
-Bryan T.\ Grenfell\textsuperscript{1,2,3,$\dagger$}
-Jonathan Dushoff\textsuperscript{4,5,6,$\dagger$}
+C\'ecile Viboud\textsuperscript{2}
+Bryan T.\ Grenfell\textsuperscript{1,2,3}
+Jonathan Dushoff\textsuperscript{4,5,6}
 \\
@@
 
-$\dagger$These authors contributed equally\\
 *Corresponding author: swp2@princeton.edu\\
@@
 
-In order to estimate the time-dependent reproduction number $R_t$ (i.e., the expected number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), we reconstructed the incidence time series (i.e., the number of infected individuals on each day).
+In order to estimate the time-dependent reproduction number $R_t$ (i.e., the expected number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), we reconstructed the incidence time series (i.e., the number of infected individuals on each day) from the daily number of reported cases by the KCDC \citep{kcdc}.
 We first adjusted the number of reported cases to account for changes in testing criteria (\fref{dist}), which occurred 4 times between January 20--March 16, 2020, as well as changes in the number of tests (Supplementary Materials).
@@
 \caption{
-\textbf{Comparison of time-dependent reproduction number and normalized traffic in Daegu and Seoul.}
-Black lines and gray ribbons represent the median estimates of $R_t$ and their corresponding 95\% credible intervals.
+\textbf{Comparison of reconstructed incidence and time-dependent reproduction number in Daegu and Seoul.}
+Black lines and gray ribbons represent the median estimates of reconstructed incidence (A,B) and $R_t$ (C,D) and their corresponding 95\% credible intervals.
 Red lines represent the normalized traffic volume.
@@
 
-\fref{eff} compares the estimates of $\mathcal R_t$ in Daegu and Seoul.
-In Daegu, estimates of $\mathcal R_t$ gradually decrease and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}A).
-Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
-The initial decrease in $\mathcal R_t$ may reflect the possibility that a large proportion of church-related transmission could have occurred much earlier; according to the KCDC press release, the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.
+\fref{eff} compares the reconstructed incidence (A,B) and estimates of $\mathcal R_t$ (C,D) in Daegu and Seoul.
+In Daegu, estimates of $\mathcal R_t$ gradually decrease and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}C).
+The initial decrease in $\mathcal R_t$ may reflect the possibility that a large proportion of church-related transmission could have occurred much earlier and therefore caused behavior change within the church; for example, the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.
+% Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
 
-On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}B);
+On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}D);
 our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less effective.
@@
 We also did not distinguish local and imported cases, which may overestimate $\mathcal R_t$ \citep{thompson2019improved}.
+\swp{The analysis of the Seoul outbreak which accounts for imported cases is presented in Supplementary Materials?}
 
@@
 Our analysis reveals that the epidemic may still persist in other regions, including Seoul and Gyeonggi-do.
-Persistent reportings of COVID-19 cases in Seoul and Gyeonggi-do (around 10 cases almost everyday) between March 11--22, 2020, provide further support for our analysis \citep{kcdc}.
+Persistent reportings of COVID-19 cases in Seoul and Gyeonggi-do (around 10 cases almost everyday) between March 11--22, 2020, provide further support for our conclusion \citep{kcdc}.
 Unless the reproduction number can be reduced below 1 in all regions, small outbreaks may continue to occur in South Korea.

### 2020-03-24 figure_compare_R_t.R

@@ -7,7 +7,27 @@ library(readxl)
 source("wquant.R")
 
 load("R_t_daegu_censor.rda")
+
+reconstruct_daegu <- reconstruct_censor %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(case),
+    lwr=quantile(case, 0.025),
+    upr=quantile(case, 0.975)
+  )
+
 load("R_t_seoul_censor.rda")
+
+reconstruct_seoul <- reconstruct_censor %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(case),
+    lwr=quantile(case, 0.025),
+    upr=quantile(case, 0.975)
+  )
+
 load("traffic_daegu.rda")
 load("traffic_seoul.rda")
 
@@ -134,7 +154,43 @@ cor.test(daegu_merge$median, daegu_merge$traffic)
 seoul_merge <- merge(rt_seoul, seoul_traffic)
 cor.test(seoul_merge$median, seoul_merge$traffic)
 
-g1 <- ggplot(rt_daegu) +
+g1 <- ggplot(reconstruct_daegu) +
+  geom_bar(data=daegu, aes(date_report, cases), stat="identity", alpha=0.3) +
+  geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.3) +
+  geom_line(aes(date, median), lwd=1) +
+  geom_vline(xintercept=as.Date("2020-02-18"), lty=2) +
+  scale_color_manual(values=c(1, 2, 4)) +
+  scale_fill_manual(values=c(1, 2, 4)) +
+  scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-20", "2020-03-16"))+c(0,1)) +
+  scale_y_continuous("Reconstructed incidence", expand=c(0, 0)) +
+  ggtitle("A. Daegu") +
+  theme(
+    panel.grid = element_blank(),
+    panel.border = element_blank(),
+    axis.line = element_line(),
+    legend.position = "none",
+    legend.title = element_blank()
+  )
+
+g2 <- ggplot(reconstruct_seoul) +
+  geom_bar(data=seoul, aes(date_report, cases), stat="identity", alpha=0.3) +
+  geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.3) +
+  geom_line(aes(date, median), lwd=1) +
[…]

### 2020-03-24 korea.tex
k no more

@@
 \end{flushleft}
+
+\section*{Abstract - 75 word limit; currently 70.}
+
+On January 20, 2020, the first COVID-19 case was confirmed in South Korea.
+As of March 18, 2020, 8,413 cases were confirmed but the number of cases has been consistently decreasing;
+the reason for the decrease has been widely attributed to its intensive testing.
+We report on the potential roles of social distancing in reducing transmission in South Korea.
+Our analysis suggests that transmission may still persist in some regions.
 

### 2020-03-24 korea.tex
format

@@
 
-On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}D);
-our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less effective.
+On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}D).
+Our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less effective.
 Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.
+
 While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.55$; 95\% CI: 0.27--0.74), the strength of correlations are likely to be strongly conflated because they do not account for the effects of any other measures that could have affected $\mathcal R_t$;

### 2020-03-24 korea.tex
k

@@
 Our analysis focused on comparing metro traffic, which serves as a proxy for the degree of social distancing, with epidemiological dynamics in two cities.
-This analysis is indirect; we are not able to directly estimate the effect of social distancing on epidemiological dynamics.
+The 80\% decrease in traffic volume suggests that the strength of social distancing in Daegu may be comparable to that in Wuhan, China \citep{lai2020effect}.
+However, this analysis is indirect; we are not able to directly estimate the effect of social distancing on epidemiological dynamics.
 Moreover, other measures, such as intensive testing of core transmission groups and school closure, are also likely to have affected the changes in $\mathcal R_t$ \citep{kcdc}.

### 2020-03-24 korea.tex
one more...
Some questions here for Daniel.

@@
 On January 20, 2020, the first COVID-19 case was confirmed in South Korea.
-As of March 18, 2020, 8,413 cases were confirmed but the number of cases has been consistently decreasing;
-the reason for the decrease has been widely attributed to its intensive testing.
-We report on the potential roles of social distancing in reducing transmission in South Korea.
-Our analysis suggests that transmission may still persist in some regions.
+After a rapid outbreak, the number of incident cases has been consistently decreasing since XX Feb/Mar;
+this decrease has been widely attributed to its intensive testing.
+We report here on the likely role of social distancing in reducing transmission in South Korea.
+Our analysis suggests that transmission may still be persisting in some regions.
 
@@
 
-Since its first appearance in Wuhan, China, in December 2019 \citep{pneumonia}, the novel coronavirus disease (COVID-19) has spread internationally, including to South Korea.
+Since its first appearance in Wuhan, China, in December 2019 \citep{pneumonia}, COVID-19 disease has spread internationally, including to South Korea.
 The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
-In February, the disease spread rapidly within a church in the city of Daegu.
+In February, the disease spread rapidly within a church community in the city of Daegu.
 The chains of transmission that began from this cluster distinguish the epidemic in South Korea from that in any other countries:
-As of March 18, 2020, 8,413 cases were confirmed, of which 60\% were related to the church and 27\% were in their 20's \citep{kcdc}.
+As of March 18, 2020, 8,413 cases were confirmed, of which 60\% were related to the church and 27\% were in their 20s \citep{kcdc}.
 Nonetheless, South Korea's intensive testing using novel contact tracing techniques allowed rapid identification and isolation cases and reduction of onward transmission \citep{fergusonimpact, tracingkorea, science}.
@@
 
-We compared epidemiological dynamics of COVID-19 from two cities in which the biggest number of COVID-19 cases have been reported: Daegu and Seoul.
+We compared epidemiological dynamics of COVID-19 from two cities in which the largest number of COVID-19 cases have been reported: Daegu and Seoul.
 Between January 20--March 16, 2020, 6,083 cases from Daegu and 248 from Seoul were reported by the KCDC (Supplementary Materials).
@@
 Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from \url{data.go.kr} and \url{data.seoul.go.kr}, respectively.
+\jd{Clarify a tiny bit:}
 We tabulated the total number of individuals who accessed subway or monorail (\fref{data});
@@
 
-In order to estimate the time-dependent reproduction number $R_t$ (i.e., the expected number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), we reconstructed the incidence time series (i.e., the number of infected individuals on each day) from the daily number of reported cases by the KCDC \citep{kcdc}.
-We first adjusted the number of reported cases to account for changes in testing criteria (\fref{dist}), which occurred 4 times between January 20--March 16, 2020, as well as changes in the number of tests (Supplementary Materials).
+\jd{Grey ribbons should still be lighter; we want eyes drawn first to the estimate.}
+
+We estimated the time-dependent reproduction number $R_t$ (the average number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), we estimated daily incidence of infection  from the daily number of reported cases by the KCDC \citep{kcdc}.
+We adjusted the number of reported cases to account for changes in testing criteria (\fref{dist}), which occurred 4 times between January 20--March 16, 2020 and changes in the number of tests (Supplementary Materials).
 A sensitivity analysis showed that qualitative patterns of inferred $R_t$ are robust to these adjustments.
 
+\jd{Need to discuss how to present this; we don't want it to get in the way.}
 We then estimated time-dependent \emph{backward} confirmation-to-onset delay distributions from the partial line list using a negative-binomial regression with the \texttt{brms} package \citep{burkner2017brms} (\fref{dist}C).
@@
 
-For each posterior sample of the reconstructed incidence time series, we divided daily incidence by the probability that they would have been reported before March 16, 2020, and estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
+\jd{This looks like backward R, which is fine, but not what we said we would do.}
+For each posterior sample of the reconstructed incidence time series, we divided daily incidence by the probability that they would have been reported before March 16, 2020, \jd{Can't we make this correction part of the reconstruction step?} and estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
 \begin{equation}
@@
 We weighted each sample of $\mathcal R_t$ by a gamma probability distribution with a mean of 2.6 and a standard deviation of 2 to reflect prior knowledge \citep{tempvar} and took weighted quantiles to calculate the medians and associated 95\% credible intervals.
-We restricted the calculation of $\mathcal R_t$ between February 2, 2020 (14 days after the first confirmed case was imported) and March 10, 2020 (because the degree of censoring would be too large after this date to estimate $\mathcal R_t$).
+\jd{Is this in fact good? Would smoothing be better?}
+We estimated $\mathcal R_t$ between February 2, 2020 (14 days after the first confirmed case was imported) and March 10, 2020 (after this point the effects of censoring are too strong for reliable estimates).
 
@@
 The intensity of intervention is likely to vary across regions given that majority of COVID-19 cases in South Korea were reported from Daegu.
-We did not have the data to account for these factors.
+We did not have sufficient data to account for these factors and therefore applied identical methods for adjusting the time series and estimating $\mathcal R_t$ across 4 regions.
 Nonetheless, the robustness of our findings is supported by the sensitivity analyses.

### 2020-03-24 korea.tex
Once through main text

@@
 
-%% \newcommand{\comment}{\showcomment}
-\newcommand{\comment}{\nocomment}
+\newcommand{\comment}{\showcomment}
+%% \newcommand{\comment}{\nocomment}
 
@@
 In Daegu, estimates of $\mathcal R_t$ gradually decrease and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}C).
-The initial decrease in $\mathcal R_t$ may reflect the possibility that a large proportion of church-related transmission could have occurred much earlier and therefore caused behavior change within the church; for example, the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.
+The initial decrease in $\mathcal R_t$ may reflect behavior change within the church; the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.
 % Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
@@
 
-While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.55$; 95\% CI: 0.27--0.74), the strength of correlations are likely to be strongly conflated because they do not account for the effects of any other measures that could have affected $\mathcal R_t$;
-in particular, the increase in the number of tests (\fref{dist}A) as well as the improvement in the accuracy of case identification (\fref{dist}D) are likely to have played important roles in reducing $\mathcal R_t$.
-We do not find any signatures of lags (Supplementary Materials).
+While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.55$; 95\% CI: 0.27--0.74), these correlations are conflated by time trends, and are also likely conflated by other measures that could have affected $\mathcal R_t$ --
+in particular, the increase in the number of tests (\fref{dist}A) as well as the improvement in the accuracy of case identification (\fref{dist}D).
+We do not find clear signatures of lagged correlation (Supplementary Materials).
 Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces (Gyeongsangbuk-do and Gyeonggi-do), providing support for the robustness of our analysis (Supplementary Materials).
@@
 
-The ongoing COVID-19 outbreak in South Korea provides a unique perspective to understanding and controlling the pandemic.
-Its experience provides evidence that the epidemic can be suppressed with less extreme measures than those taken by China \citep{kickbusch2020response}.
+The South Korean experience with COVID-19 provides evidence that epidemics can be suppressed with less extreme measures than those taken by China \citep{kickbusch2020response}.
 Our analysis reveals potential roles of social distancing in mitigating the COVID-19 epidemic in South Korea.
@@
 Finally, our study highlights the importance of considering geographical heterogeneity in estimating epidemic potential.
-The recent decrease in the number of reported cases in South Korea is likely to be strongly correlated with the decrease in the number of cases in Daegu.
-Our analysis reveals that the epidemic may still persist in other regions, including Seoul and Gyeonggi-do.
-Persistent reportings of COVID-19 cases in Seoul and Gyeonggi-do (around 10 cases almost everyday) between March 11--22, 2020, provide further support for our conclusion \citep{kcdc}.
+The recent decrease in the number of reported cases in South Korea is driven by the sharp decrease in Daegu.
+Our analysis reveals that the epidemic may still persist in other regions, including Seoul and Gyeonggi-do;
+reports from Seoul and Gyeonggi-do (around 10 cases almost every day between March 11--22, 2020) provide further support for our conclusion \citep{kcdc}.
 Unless the reproduction number can be reduced below 1 in all regions, small outbreaks may continue to occur in South Korea.

### 2020-03-24 korea.tex
Meeting with Daniel

@@
 Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from \url{data.go.kr} and \url{data.seoul.go.kr}, respectively.
-\jd{Clarify a tiny bit:}
-We tabulated the total number of individuals who accessed subway or monorail (\fref{data});
-we restricted our analysis to subway lines 1--9 in Seoul.
+We tabulated the total number of individuals who accessed the subway or monorail (\fref{data}) (using Seoul lines 1--9, and Daegu lines 1--3).
 Soon after the first church-related case was confirmed in Daegu on Feb 18, 2020, the daily traffic volume decreased by about 80\% and 50\% in Daegu and Seoul, respectively.
@@
 
-\jd{Need to discuss how to present this; we don't want it to get in the way.}
+\jd{Daniel to move, me to clean up both text and supp.}
 We then estimated time-dependent \emph{backward} confirmation-to-onset delay distributions from the partial line list using a negative-binomial regression with the \texttt{brms} package \citep{burkner2017brms} (\fref{dist}C).
@@
 
-\jd{This looks like backward R, which is fine, but not what we said we would do.}
-For each posterior sample of the reconstructed incidence time series, we divided daily incidence by the probability that they would have been reported before March 16, 2020, \jd{Can't we make this correction part of the reconstruction step?} and estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
+\jd{Me to change backward/forward R description above, Daniel to approve.}
+For each posterior sample of the reconstructed incidence time series, we divided daily incidence by the probability that they would have been reported before March 16, 2020, \jd{Make this correction part of the reconstruction step?} and estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
 \begin{equation}
@@
 We weighted each sample of $\mathcal R_t$ by a gamma probability distribution with a mean of 2.6 and a standard deviation of 2 to reflect prior knowledge \citep{tempvar} and took weighted quantiles to calculate the medians and associated 95\% credible intervals.
-\jd{Is this in fact good? Would smoothing be better?}
 We estimated $\mathcal R_t$ between February 2, 2020 (14 days after the first confirmed case was imported) and March 10, 2020 (after this point the effects of censoring are too strong for reliable estimates).

### 2020-03-24 korea.tex
too much at once!! hope github doesn't complain

@@
 
-\section*{Abstract - 75 word limit; currently 70.}
+\section*{Abstract - 75 word limit; currently 67.}
 
 On January 20, 2020, the first COVID-19 case was confirmed in South Korea.
-After a rapid outbreak, the number of incident cases has been consistently decreasing since XX Feb/Mar;
+After a rapid outbreak, the number of incident cases has been consistently decreasing since early March;
 this decrease has been widely attributed to its intensive testing.
@@
 
+\section*{Contribution}
+
+Data collection: SWP; conceptualization: SWP, CV; analysis: SWP; first draft: SWP, JD. All authors contributed to the writing and approval of the final report.
+
+\section*{Data availability}
+
+All transcribed data and original reports used in this analysis are stored in a publicly available GitHub repository: \url{https://github.com/parksw3/COVID19-Korea}.
+All code and traffic data are stored in a separate GitHub repository: \url{https://github.com/parksw3/Korea-analysis}.
+
 \pagebreak
+
+\section{Background}
 
@@
 Partial line lists were transcribed from press releases from KCDC and various local and provincial governments.
-All data and original reports are stored in a publicly available GitHub repository: \url{https://github.com/parksw3/COVID19-Korea}.
+All data and original reports are stored in a publicly available GitHub repository.
 
@@
 We compared epidemiological dynamics of COVID-19 from two cities in which the largest number of COVID-19 cases have been reported: Daegu and Seoul.
-Between January 20--March 16, 2020, 6,083 cases from Daegu and 248 from Seoul were reported by the KCDC (Supplementary Materials).
-Unlike the epidemic in Daegu, which is characterized by a single, large peak followed by a gradual decrease, the epidemic in Seoul consists of several small outbreaks (\fref{data}).
+Between January 20--March 16, 2020, 6,083 cases from Daegu and 248 from Seoul were reported by the KCDC.
+Unlike the epidemic in Daegu, which is characterized by a single, large peak followed by a gradual decrease, the epidemic in Seoul consists of several small outbreaks (\fref{data}; see Supplementary Materials for details).
 
@@
 
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figure_report_delay.pdf}
-\caption{
-\textbf{Changes in the number of tests and delay distributions over time.}
-}
-Vertical lines indicate the date on which testing criteria expanded.
-Boxplots (C--D) represent the observed delays.
-Black lines and gray ribbons represent the median estimates of the mean delays and their associated 95\% credible intervals.
-\label{fig:dist}
-\end{figure}
-
-\jd{Grey ribbons should still be lighter; we want eyes drawn first to the estimate.}
-
-We estimated the time-dependent reproduction number $R_t$ (the average number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), we estimated daily incidence of infection  from the daily number of reported cases by the KCDC \citep{kcdc}.
-We adjusted the number of reported cases to account for changes in testing criteria (\fref{dist}), which occurred 4 times between January 20--March 16, 2020 and changes in the number of tests (Supplementary Materials).
-A sensitivity analysis showed that qualitative patterns of inferred $R_t$ are robust to these adjustments.
-
-\jd{Daniel to move, me to clean up both text and supp.}
-We then estimated time-dependent \emph{backward} confirmation-to-onset delay distributions from the partial line list using a negative-binomial regression with the \texttt{brms} package \citep{burkner2017brms} (\fref{dist}C).
-We combined the estimated posterior distribution of delay distributions with previously estimated incubation period distribution (Table 1) to obtain posterior samples for date of infection for each confirmed case.
-Likewise, we estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using the same negative-binomial regression model, while accounting for right-censoring (\fref{dist}D),
-to calculate the median probability that a case infected on a given day will be reported before March 16, 2020.
-Changes in the mean backward delay (\fref{dist}C) reflect changes in incidence of symptomatic cases, whereas the decrease in the mean forward delay (\fref{dist}D) reflects the improvement in the accuracy of case identification \citep{lai2020effect}.
+To estimate the time-dependent reproduction number $R_t$ (the average number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), we first estimated daily incidence of infection from the daily number of reported cases by the KCDC \citep{kcdc}.
+We adjusted the number of reported cases to account for changes in testing criteria, which occurred 4 times between January 20--March 16, 2020.
+Then, we inferred onset-to-confirmation delay distributions from the partial line list;
+we combined the inferred delay distribution with previously estimated incubation period distribution (Table 1) to obtain probability distributions for date of infection for each reported case.
+We accounted for right-censoring by dividing the daily incidence by the probability that a case infected on a given day would have been reported before March 16, 2020.
 Implementation details are provided in the Supplementary Materials.
@@
 \jd{Me to change backward/forward R description above, Daniel to approve.}
-For each posterior sample of the reconstructed incidence time series, we divided daily incidence by the probability that they would have been reported before March 16, 2020, \jd{Make this correction part of the reconstruction step?} and estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
+We estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
 \begin{equation}
@@
 \end{equation}
-where $I_t$ is the censoring-adjusted incidence time series and $w_k$ is the generation-interval distribution randomly drawn from a prior distribution (Table 1).
+where $I_t$ is the reconstructed incidence time series (i.e., the number of infected cases on day $t$) and $w_k$ is the generation-interval distribution randomly drawn from a prior distribution (Table 1).
 We weighted each sample of $\mathcal R_t$ by a gamma probability distribution with a mean of 2.6 and a standard deviation of 2 to reflect prior knowledge \citep{tempvar} and took weighted quantiles to calculate the medians and associated 95\% credible intervals.
@@
 Black lines and gray ribbons represent the median estimates of reconstructed incidence (A,B) and $R_t$ (C,D) and their corresponding 95\% credible intervals.
+Bar plots show the number of reported cases.
 Red lines represent the normalized traffic volume.
@@
 \fref{eff} compares the reconstructed incidence (A,B) and estimates of $\mathcal R_t$ (C,D) in Daegu and Seoul.
-In Daegu, estimates of $\mathcal R_t$ gradually decrease and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}C).
+In Daegu, incidence peaked shortly after the first case was confirmed and decreased (\fref{eff}A).
+Likewise, the estimates of $\mathcal R_t$ gradually decrease and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}C).
 The initial decrease in $\mathcal R_t$ may reflect behavior change within the church; the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.
-% Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
+Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
 
-On the other hand, estimates of $\mathcal R_t$ remain around 1 in Seoul (\fref{eff}D).
+In Seoul, incidence decreases slightly as traffic decreases but estimates of $\mathcal R_t$ remain around 1 throughout (\fref{eff}D).
 Our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less effective.
@@
 
-While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.87$; 95\% CI: 0.74--0.94) and Seoul ($r=0.55$; 95\% CI: 0.27--0.74), these correlations are conflated by time trends, and are also likely conflated by other measures that could have affected $\mathcal R_t$ --
-in particular, the increase in the number of tests (\fref{dist}A) as well as the improvement in the accuracy of case identification (\fref{dist}D).
-We do not find clear signatures of lagged correlation (Supplementary Materials).
+While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.90$; 95\% CI: 0.79--0.95) and Seoul ($r=0.76$; 95\% CI: 0.57--0.87), these correlations are conflated by time trends, and are also likely conflated by other measures that could have affected $\mathcal R_t$ --
+in particular, the increase in the number of tests as well as the improvement in the accuracy of case identification (Supplementary Materials).
+We do not find clear signatures of lagged correlation between $\mathcal R_t$ and traffic volume (Supplementary Materials).
 Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces (Gyeongsangbuk-do and Gyeonggi-do), providing support for the robustness of our analysis (Supplementary Materials).
@@
 Even though social distancing alone may not be able to fully prevent the spread of the disease, its ability to flatten the epidemic curve (cf.
-\fref{eff}B) reduces burden for healthcare system and provides time to plan for the future \citep{anderson2020will}.
+\fref{eff}B,D) reduces burden for healthcare system and provides time to plan for the future \citep{anderson2020will}.
 
@@
 
-\section*{Contribution}
-
-Data collection: SWP; conceptualization: SWP, CV; analysis: SWP; first draft: SWP, JD. All authors contributed to the writing and approval of the final report.
-
 \pagebreak
@@
 
-\renewcommand\thefigure{S\arabic{figure}}
-\setcounter{figure}{0}
-
-\section*{Supplementary Materials}
-
-\subsection*{Epidemiological data}
-
-The daily number of reported cases from each region was translated and transcribed from the KCDC press release \citep{kcdc}.
-Following the KCDC's protocol, the daily number of reported cases prior to February 20, 2020, reflects the number of confirmed cases on each day.
-Between February 21 -- March 1, 2020, the daily number of reported cases reflects the number of reported cases within the last 24 hours (9 AM to 9 AM).
-On March 2, 2020, the daily number of reported cases reflects the number of cases that were reported between 9 AM March 1, 2020, and 12 AM March 2, 2020.
-Since then, the daily number of reported cases reflects the number of reported cases within the last 24 hours (12 AM to 12 AM).
-The daily number of reported cases by the KCDC may be different from the reports by each city's government as some cases may be transferred after they are confirmed.
-
-\subsection*{Time series adjustments}
-
-First, we accounted for changes in testing criteria, which occurred 4 times between January 20--March 16, 2020, by assuming that the proportion positive should remain roughly constant if we follow a consistent protocol of identifying and deciding whom to test.
-To do so, we calculated the relative proportion of positive cases during each period (divided by the the between-period mean) and multiplied the daily number of reported cases by the relative proportions of the corresponding criterion.
-Likewise, we accounted for changes in the number of tests completed on each day by dividing the scaled number of reported cases by the relative number of tests completed on each day; this step was performed separately for each testing criterion because widening of criteria necessarily increases the number of tests.
-The number of negative cases was not reported on January 25 and 31, 2020; we took the average of cumulative negative cases from one day before and after these dates instead to impute missing values.
-
-\subsection*{Estimating delay distributions between onset and reporting}
-
-We estimated both forward (onset-to-confirmation) and backward (confirmation-to-onset) delay distributions using Bayesian negative binomial regressions with log links.
-The backward delay distribution is inferred directly using the \texttt{brms} package \citep{burkner2017brms}.
-The time-dependent mean of the negative binomial distribution is modeled using splines.
-We assumed weakly informative priors on the fixed effects: normal distributions with mean of 0 and standard deviation of 2;
-note that these distributions are priors on link scale.
-
-To estimate the forward delay distribution, we modified the stan code to account for right-censoring and ran the code using the \texttt{rstan} package \citep{rstan}.
-In particular, we modified the likelihood such that given a delay of $x_i$ days, symptom onset day $t_i$ and the day of measurement of $k$, the likelihood of observing the delay is given by:
-\begin{equation}
-\frac{f(x_i|\mu(t_i), \theta)}{F(k-t_i|\mu(t_i), \theta)},
-\end{equation}
-where $f$ is the negative binomial distribution with time-dependent mean $\mu(t_i)$ and dispersion parameter $\theta$.
-This likelihood accounts for the fact that the delay between symptom onset and confirmation cannot be longer than $k-t_i$ (otherwise, the case will be reported after the date of measurement).
-Convergence is assessed by ensuring that the Gelman-Rubin statistic is lower than 1.01 for all parameters \citep{gelman1992inference}.
-
-\pagebreak
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figure_R_t_all.pdf}
-\caption{
-\textbf{Comparison of effective reproduction number and the daily number of reported cases in Daegu, Seoul, Gyeongsangbuk-do, and Gyeonggi-do.}
-}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figure_R_t_daegu.pdf}
-\caption{
-\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Daegu.}
-}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figure_R_t_seoul.pdf}
-\caption{
-\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Seoul.}
-}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figure_R_t_gyeongbuk.pdf}
-\caption{
-\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Gyeongsangbuk-do.}
-}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figure_R_t_gyeonggi.pdf}
-\caption{
-\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Gyeonggi-do.}
-}
-\end{figure}
-
-\pagebreak
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figure_cross.pdf}
-\caption{
-\textbf{Cross correlation between the normalized traffic volume and the median estimates of $\mathcal R_t$.}
-}
-\end{figure}
+\input{supp.tex}
 

### 2020-03-24 supp.tex
too much at once!! hope github doesn't complain

\renewcommand\thefigure{S\arabic{figure}}
\setcounter{figure}{0}

\section*{Supplementary Materials}

\subsection*{Epidemiological data}

The daily number of reported cases from each region was translated and transcribed from the KCDC press release \citep{kcdc}.
Following the KCDC's protocol, the daily number of reported cases prior to February 20, 2020, reflects the number of confirmed cases on each day.
Between February 21 -- March 1, 2020, the daily number of reported cases reflects the number of reported cases within the last 24 hours (9 AM to 9 AM).
On March 2, 2020, the daily number of reported cases reflects the number of cases that were reported between 9 AM March 1, 2020, and 12 AM March 2, 2020.
Since then, the daily number of reported cases reflects the number of reported cases within the last 24 hours (12 AM to 12 AM).
The daily number of reported cases by the KCDC may be different from the reports by each city's government as some cases may be transferred after they are confirmed.

\subsection*{Reconstruction of incidence time series}

Testing criteria expanded 4 times between January 20--March 16, 2020: January 28, February 7, February 20, and March 2, 2020.
We accounted for these changes by assuming that the proportion positive should remain roughly constant if we follow a consistent protocol of identifying and deciding whom to test.
To do so, we calculated the relative proportion of positive cases during each period (divided by the the between-period mean) and multiplied the daily number of reported cases by the relative proportions of the corresponding criterion.
A sensitivity analysis showed that results are robust to these adjustments.

We then estimated time-dependent \emph{backward} onset-to-confirmation delay distributions from the partial line list. We asked: Given a cohort of individuals who were reported on the same day, what is the probability distribution of the onset-to-confirmation delay?
The backward delay distribution depends on changes in the number of symptomatic cases ---
e.g., when the number of symptomatic cases are increasing, the backward delay distribution is likely to be shorter because individuals are more likely to have developed symptoms recently.
The backward delay distribution was inferred using a negative-binomial regression with log-link using the \texttt{brms} package \citep{burkner2017brms}.
Time-dependent mean of the negative binomial distribution is modeled using splines.
We assumed weakly informative priors on the fixed effects: normal distributions with mean of 0 and standard deviation of 2;
these distributions are priors on link scale.

For each posterior sample of the backward delay distribution, we drew a random sample of onset-to-confirmation delay and incubation period for each confirmed case. 
This allowed us to obtain posterior samples of possible infection dates for each case,
which was then converted into posterior samples of incidence time series.

To account for right-censoring in the reported cases, we also estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using the same negative-binomial regression model.
Here, we asked: Given a cohort of individuals who became symptomatic on the same day, what is the probability distribution of the onset-to-confirmation delay?
The forward delay distribution reflects the changes in the accuracy of case identification --- e.g., a decrease in the delay reflects improvement in accuracy.

To estimate the forward delay distribution, we modified the stan code from the negative-binomial regression that we used to infer the backward delay distribution to account for right-censoring (in the observed delays) and ran the code using the \texttt{rstan} package \citep{rstan}.
In particular, we modified the likelihood such that given a delay of $x_i$ days, symptom onset day $t_i$ and the day of measurement of $k$, the likelihood of observing the delay is given by:
\begin{equation}
\frac{f(x_i|\mu(t_i), \theta)}{F(k-t_i|\mu(t_i), \theta)},
\end{equation}
where $f$ is the negative binomial distribution with time-dependent mean $\mu(t_i)$ and dispersion parameter $\theta$. This likelihood accounts for the fact that the delay between symptom onset and confirmation cannot be longer than $k-t_i$ (otherwise, the case will be reported after the date of measurement). Convergence is assessed by the lack of warning messages from the \texttt{rstan} package \citep{rstan}.

For each combination of date of infection and a posterior sample of the forward delay distribution, we drew 1000 samples of incubation periods and onset-to-confirmation delays and calculated the median probability that an individual infected on a given day will be confirmed before March 16, 2020.
We divided the daily number of

We combined the forward delay distribution with the , we calculated the median probability that a

to calculate the median probability that a case infected on a given day will be reported before March 16, 2020.
Changes in the mean backward delay (\fref{dist}C) reflect changes in incidence of symptomatic cases, whereas the decrease in the mean forward delay (\fref{dist}D) reflects the improvement in the accuracy of case identification \citep{lai2020effect}.

\pagebreak

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_report_delay.pdf}
\caption{
\textbf{Changes in the number of tests and delay distributions over time.}
}
Vertical lines indicate the date on which testing criteria expanded.
Boxplots (C--D) represent the observed delays.
Black lines and gray ribbons represent the median estimates of the mean delays and their associated 95\% credible intervals.
\end{figure}

\pagebreak

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_R_t_all.pdf}
\caption{
\textbf{Comparison of effective reproduction number and the daily number of reported cases in Daegu, Seoul, Gyeongsangbuk-do, and Gyeonggi-do.}
}
\end{figure}

\pagebreak

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_R_t_daegu.pdf}
\caption{
\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Daegu.}
}
\end{figure}

\pagebreak

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_R_t_seoul.pdf}
\caption{
\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Seoul.}
}
\end{figure}

\pagebreak

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_R_t_gyeongbuk.pdf}
\caption{
\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Gyeongsangbuk-do.}
}
\end{figure}

\pagebreak

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_R_t_gyeonggi.pdf}
\caption{
\textbf{Sensitivity analysis of estimates of $\mathcal R_t$ in Gyeonggi-do.}
}
\end{figure}

\pagebreak

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figure_cross.pdf}
\caption{
\textbf{Cross correlation between the normalized traffic volume and the median estimates of $\mathcal R_t$.}
}
\end{figure}

### 2020-03-24 R_t_daegu_censor.R

@@ -1,50 +1,0 @@
-library(dplyr)
-load("reconstruct_time_series_daegu.rda")
-load("degree_censor.rda")
-
-reconstruct_censor <- R_t_daegu_censor <- vector('list', length(reconstruct_time_series_daegu))
-
-set.seed(101)
-for (i in 1:length(reconstruct_time_series_daegu)) {
-  rr <- reconstruct_time_series_daegu[[i]]
-  
-  # pgamma(7, 62, 62/5) - pgamma(4, 62, 62/5) approx .95
-  
-  meangen <- rgamma(1, 62, 62/5)
-  size <- rgamma(1, 20, 20/5)
-  
-  gen <- dnbinom(1:14, mu=meangen, size=size)
-  gen <- gen/sum(gen)
-  
-  recondata <- rr %>%
-    group_by(infection) %>%
-    summarize(
-      case=length(infection)
-    ) %>%
-    rename(
-      date=infection
-    ) %>%
-    filter(date >= as.Date("2020-01-19"), date <= as.Date("2020-03-10"))
-  
-  dd <- data.frame(
-    date=seq.Date(as.Date("2020-01-19"), as.Date("2020-03-10"), 1),
-    case=0
-  )
-  
-  dd$case[match(recondata$date, dd$date)] <- recondata$case
-  dd$case <- dd$case/(1-degree_censor$median)
-  
-  reconstruct_censor[[i]] <- dd
-  
-  IRt <- tail(dd$case, -14)/sapply(1:(length(dd$case)-14), function(x) sum(dd$case[x:(x+13)]*rev(gen)))
-  
-  Rt <- sapply(1:(length(IRt)-13), function(x) sum(IRt[x:(x+13)]*gen))
-  
-  R_t_daegu_censor[[i]] <- data.frame(
-    date=seq.Date(as.Date("2020-01-19")+14, as.Date("2020-03-10"), 1),
-    IRt=IRt,
-    Rt=c(Rt, rep(NA, 13))
-  )
-}
-
-save("R_t_daegu_censor", "reconstruct_censor", file="R_t_daegu_censor.rda")

### 2020-03-24 R_t_gyeonggi_censor.R

@@ -1,50 +1,0 @@
-library(dplyr)
-load("reconstruct_time_series_gyeonggi.rda")
-load("degree_censor.rda")
-
-reconstruct_censor <- R_t_gyeonggi_censor <- vector('list', length(reconstruct_time_series_gyeonggi))
-
-set.seed(101)
-for (i in 1:length(reconstruct_time_series_gyeonggi)) {
-  rr <- reconstruct_time_series_gyeonggi[[i]]
-  
-  # pgamma(7, 62, 62/5) - pgamma(4, 62, 62/5) approx .95
-  
-  meangen <- rgamma(1, 62, 62/5)
-  size <- rgamma(1, 20, 20/5)
-  
-  gen <- dnbinom(1:14, mu=meangen, size=size)
-  gen <- gen/sum(gen)
-  
-  recondata <- rr %>%
-    group_by(infection) %>%
-    summarize(
-      case=length(infection)
-    ) %>%
-    rename(
-      date=infection
-    ) %>%
-    filter(date >= as.Date("2020-01-19"), date <= as.Date("2020-03-10"))
-  
-  dd <- data.frame(
-    date=seq.Date(as.Date("2020-01-19"), as.Date("2020-03-10"), 1),
-    case=0
-  )
-  
-  dd$case[match(recondata$date, dd$date)] <- recondata$case
-  dd$case <- dd$case/(1-degree_censor$median)
-  
-  reconstruct_censor[[i]] <- dd
-  
-  IRt <- tail(dd$case, -14)/sapply(1:(length(dd$case)-14), function(x) sum(dd$case[x:(x+13)]*rev(gen)))
-  
-  Rt <- sapply(1:(length(IRt)-13), function(x) sum(IRt[x:(x+13)]*gen))
-  
-  R_t_gyeonggi_censor[[i]] <- data.frame(
-    date=seq.Date(as.Date("2020-01-19")+14, as.Date("2020-03-10"), 1),
-    IRt=IRt,
-    Rt=c(Rt, rep(NA, 13))
-  )
-}
-
-save("R_t_gyeonggi_censor", "reconstruct_censor", file="R_t_gyeonggi_censor.rda")

### 2020-03-24 R_t_seoul_censor.R

@@ -1,50 +1,0 @@
-library(dplyr)
-load("reconstruct_time_series_seoul.rda")
-load("degree_censor.rda")
-
-reconstruct_censor <- R_t_seoul_censor <- vector('list', length(reconstruct_time_series_seoul))
-
-set.seed(101)
-for (i in 1:length(reconstruct_time_series_seoul)) {
-  rr <- reconstruct_time_series_seoul[[i]]
-  
-  # pgamma(7, 62, 62/5) - pgamma(4, 62, 62/5) approx .95
-  
-  meangen <- rgamma(1, 62, 62/5)
-  size <- rgamma(1, 20, 20/5)
-  
-  gen <- dnbinom(1:14, mu=meangen, size=size)
-  gen <- gen/sum(gen)
-  
-  recondata <- rr %>%
-    group_by(infection) %>%
-    summarize(
-      case=length(infection)
-    ) %>%
-    rename(
-      date=infection
-    ) %>%
-    filter(date >= as.Date("2020-01-19"), date <= as.Date("2020-03-10"))
-  
-  dd <- data.frame(
-    date=seq.Date(as.Date("2020-01-19"), as.Date("2020-03-10"), 1),
-    case=0
-  )
-  
-  dd$case[match(recondata$date, dd$date)] <- recondata$case
-  dd$case <- dd$case/(1-degree_censor$median)
-  
-  reconstruct_censor[[i]] <- dd
-  
-  IRt <- tail(dd$case, -14)/sapply(1:(length(dd$case)-14), function(x) sum(dd$case[x:(x+13)]*rev(gen)))
-  
-  Rt <- sapply(1:(length(IRt)-13), function(x) sum(IRt[x:(x+13)]*gen))
-  
-  R_t_seoul_censor[[i]] <- data.frame(
-    date=seq.Date(as.Date("2020-01-19")+14, as.Date("2020-03-10"), 1),
-    IRt=IRt,
-    Rt=c(Rt, rep(NA, 13))
-  )
-}
-
-save("R_t_seoul_censor", "reconstruct_censor", file="R_t_seoul_censor.rda")

### 2020-03-24 figure_R_t_seoul.R

@@ -3,24 +3,10 @@ library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
 source("wquant.R")
 
-load("R_t_seoul_censor.rda")
 load("R_t_seoul_censor_detect.rda")
 load("R_t_seoul_censor_raw.rda")
 
 R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
-
-rt_censor <- R_t_seoul_censor %>%
-  bind_rows(.id="sim") %>%
-  group_by(date) %>%
-  mutate(weight=R0prior(IRt)) %>%
-  summarize(
-    median=wquant(IRt, weights=weight,0.5),
-    lwr=wquant(IRt, weights=weight,0.025),
-    upr=wquant(IRt, weights=weight,0.975)
-  ) %>%
-  mutate(
-    group="adjusted for detection rate + number of tests"
-  )
 
 rt_censor_detect <- R_t_seoul_censor_detect %>%
   bind_rows(.id="sim") %>%
@@ -32,7 +18,7 @@ rt_censor_detect <- R_t_seoul_censor_detect %>%
     upr=wquant(IRt, weights=weight,0.975)
   ) %>%
   mutate(
-    group="adjusted for detection rate"
+    group="adjusted for testing criteria"
   )
 
 rt_censor_raw <- R_t_seoul_censor_raw %>%
@@ -48,28 +34,15 @@ rt_censor_raw <- R_t_seoul_censor_raw %>%
     group="raw"
   )
 
-rt_all <- bind_rows(rt_censor, rt_censor_detect, rt_censor_raw) %>%
+rt_all <- bind_rows(rt_censor_detect, rt_censor_raw) %>%
   mutate(
     group=factor(group, level=c("raw",
-                                "adjusted for detection rate",
-                                "adjusted for detection rate + number of tests"))
+                                "adjusted for testing criteria"))
   )
 
 geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3)
 
 geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
-
-reconstruct_case <- reconstruct_censor %>%
-  bind_rows(.id="sim") %>%
-  group_by(date) %>%
-  summarize(
-    median=median(case),
-    lwr=quantile(case, 0.025),
-    upr=quantile(case, 0.975)
-  ) %>%
[…]

### 2020-03-24 figure_adjusted_timeseries.R

@@ -1,98 +1,0 @@
-library(dplyr)
-library(tidyr)
-library(readxl)
-library(ggplot2); theme_set(theme_bw())
-library(gridExtra)
-
-load("test_weight.rda")
-
-geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
-  mutate(
-    date_report=as.Date(date_report)
-  ) 
-
-geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
-
-daegu <- geo %>% 
-  group_by(date_report) %>%
-  summarize(
-    cases=sum(Daegu, na.rm=TRUE)
-  ) %>%
-  merge(test_weight) %>%
-  mutate(
-    cases2=cases*weight,
-    cases3=round(cases2*weight2),
-    cases2=round(cases2)
-  ) %>%
-  gather(
-    key, value,-date_report, -meanp, -test, -weight, -weight2, -group
-  ) %>%
-  mutate(
-    date_report=as.Date(date_report),
-    key=factor(key, levels=c("cases", "cases2", "cases3"),
-               labels=c("raw", "adjusted for detection rate", "adjusted for detection rate + number of tests"))
-  )
-
-seoul <- geo %>% 
-  group_by(date_report) %>%
-  summarize(
-    cases=sum(Seoul, na.rm=TRUE)
-  ) %>%
-  merge(test_weight) %>%
-  mutate(
-    cases2=cases*weight,
-    cases3=round(cases2*weight2),
-    cases2=round(cases2)
-  ) %>%
-  gather(
-    key, value,-date_report, -meanp, -test, -weight, -weight2, -group
-  ) %>%
-  mutate(
-    date_report=as.Date(date_report),
-    key=factor(key, levels=c("cases", "cases2", "cases3"),
-               labels=c("raw", "adjusted for detection rate", "adjusted for detection rate + number of tests"))
-  )
-
-g1 <- ggplot(daegu) +
-  geom_point(aes(date_report, value, col=key, shape=key)) +
-  geom_line(aes(date_report, value, col=key, lty=key)) +
-  scale_x_date("Date", expand=c(0, 0)) +
-  scale_y_continuous("Scaled number of reported cases", expand=c(0, 0)) +
-  scale_color_manual(values=c(1, 2, 4)) +
-  ggtitle("A. Daegu") +
-  theme(
-    panel.grid = element_blank(),
[…]

### 2020-03-24 figure_cross.R

@@ -4,9 +4,12 @@ library(lubridate)
 library(dplyr)
 library(tidyr)
 library(readxl)
+source("wquant.R")
 
-load("R_t_daegu_censor.rda")
-load("R_t_seoul_censor.rda")
+R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
+
+load("R_t_daegu_censor_detect.rda")
+load("R_t_seoul_censor_detect.rda")
 load("traffic_daegu.rda")
 load("traffic_seoul.rda")
 
@@ -61,13 +64,14 @@ traffic_daegu4 <- dplyr::select(ungroup(filter(traffic_daegu,year==2020)), -월,
     date >= as.Date("2020-01-20"), date <= as.Date("2020-02-29")
   )
 
-rt_daegu <- R_t_daegu_censor %>%
+rt_daegu <- R_t_daegu_censor_detect %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE),
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
+    upr=wquant(IRt, weights=weight,0.975),
     location="Daegu"
   )
 
@@ -108,13 +112,14 @@ traffic_seoul4 <- ungroup(filter(traffic_seoul,year==2020)) %>%
     date >= as.Date("2020-01-20"), date <= as.Date("2020-03-10")
   )
 
-rt_seoul <- R_t_seoul_censor %>%
+rt_seoul <- R_t_seoul_censor_detect %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
   summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE),
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
+    upr=wquant(IRt, weights=weight,0.975),
     location="Seoul"
   )
 

### 2020-03-24 figure_report_delay.R

@@ -203,10 +203,10 @@ g3 <- ggplot(covid_delay3) +
   # annotate("text", x=as.Date("2020-01-20"), y=23.5, label="Imported cases introduced", angle=90) +
   # geom_vline(xintercept=as.Date("2020-01-19"), col="red", lwd=2) +
   scale_size_area("Number of samples", guide=FALSE, max_size=2) +
-  scale_y_continuous("Confirmation to symptom onset (days)", limit=c(0, 21), expand=c(0, 0)) +
+  scale_y_continuous("Symptom onset to confirmation (days)", limit=c(0, 21), expand=c(0, 0)) +
   scale_x_date("Date of symptom onset", expand=c(0, 0), limits=c(as.Date("2020-01-20"), as.Date("2020-03-16"))+c(0, 0.5)) +
   coord_cartesian(clip = 'off') +
-  ggtitle("C. Backward confirmation-to-onset delay") +
+  ggtitle("C. Backward onset-to-confirmation delay") +
   theme(
     panel.grid = element_blank(),
     panel.border = element_blank(),

### 2020-03-24 reconstruct_time_series_daegu.R

@@ -1,124 +1,0 @@
-library(dplyr)
-library(tidyr)
-library(readxl)
-library(lubridate)
-library(brms)
-library(rstan)
-
-load("test_weight.rda")
-load("report_delay_backward.rda")
-
-covid_line <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA") %>%
-  mutate(
-    age=as.numeric(age),
-    date_onset=as.Date(date_onset)
-  )
-
-casenum <- covid_line$case[!is.na(covid_line$case)]
-
-covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=1, na="NA") %>%
-  mutate(
-    date_onset=as.Date(date_onset)
-  ) %>%
-  filter(
-    !(case %in% casenum)
-  )
-
-covid_chungnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=2, na="NA") %>%
-  mutate(
-    date_onset=as.Date(date_onset)
-  ) %>%
-  filter(
-    !(case %in% casenum)
-  )
-
-covid_busan <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=3, na="NA") %>%
-  mutate(
-    date_onset=as.Date(date_onset)
-  ) %>%
-  filter(
-    !(case %in% casenum)
-  )
-
-covid_gyeongnam <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=4, na="NA") %>%
-  mutate(
-    date_onset=as.Date(date_onset)
-  ) %>%
-  filter(
-    !(case %in% casenum)
-  )
-
-covid_all <- rbind(covid_line, covid_seoul, covid_chungnam, covid_busan, covid_gyeongnam)
-
-covid_delay <- covid_all %>%
-  filter(!is.na(date_onset), !is.na(date_confirm)) %>%
-  mutate(
-    delay=yday(date_confirm)-yday(date_onset),
-    day=yday(date_confirm),
-    day=day-min(day)
-  ) %>%
-  filter(yday(date_confirm) <= yday(as.Date("2020-03-16")))
-
-covid_delay_fake <- data.frame(
-  day=covid_delay$day[covid_delay$date_confirm==as.Date("2020-01-20")]:56,
-  delay=1
-)
-
[…]

### 2020-03-24 wquant.R

@@ -1,0 +1,14 @@
+wquant <- function (x, weights, probs = c(0.025, 0.975)) {
+  which <- !is.na(weights)
+  x <- x[which]
+  weights <- weights[which]
+  
+  if (all(is.na(x)) || length(x) == 0) return(rep(NA, length(probs)))
+  
+  idx <- order(x)
+  x <- x[idx]
+  weights <- weights[idx]
+  w <- cumsum(weights)/sum(weights)
+  rval <- approx(w,x,probs,rule=1)
+  rval$y
+}

### 2020-03-24 korea.tex
korea

@@
 Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from \url{data.go.kr} and \url{data.seoul.go.kr}, respectively.
-We tabulated the total number of individuals who accessed the subway or monorail (\fref{data}) (using Seoul lines 1--9, and Daegu lines 1--3).
+We tabulated the total number of individuals who accessed the subway or monorail (using Seoul lines 1--9, and Daegu lines 1--3; \fref{data}).
 Soon after the first church-related case was confirmed in Daegu on Feb 18, 2020, the daily traffic volume decreased by about 80\% and 50\% in Daegu and Seoul, respectively.
@@
 
-In Seoul, incidence decreases slightly as traffic decreases but estimates of $\mathcal R_t$ remain around 1 throughout (\fref{eff}D).
+In Seoul, incidence decreases slightly as traffic decreased (\fref{eff}B) but estimates of $\mathcal R_t$ remain around 1 throughout (\fref{eff}D).
 Our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less effective.
@@
 We did not have sufficient data to account for these factors and therefore applied identical methods for adjusting the time series and estimating $\mathcal R_t$ across 4 regions.
-Nonetheless, the robustness of our findings is supported by the sensitivity analyses.
+Nonetheless, the robustness of our findings is supported by the sensitivity analyses (Supplementary Materials).
 We also did not distinguish local and imported cases, which may overestimate $\mathcal R_t$ \citep{thompson2019improved}.

### 2020-03-24 korea.tex
updates

@@
 
-While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.90$; 95\% CI: 0.79--0.95) and Seoul ($r=0.76$; 95\% CI: 0.57--0.87), these correlations are conflated by time trends, and are also likely conflated by other measures that could have affected $\mathcal R_t$ --
+While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.90$; 95\% CI: 0.79--0.95) and Seoul ($r=0.76$; 95\% CI: 0.59--0.87), these correlations are conflated by time trends, and are also likely conflated by other measures that could have affected $\mathcal R_t$ --
 in particular, the increase in the number of tests as well as the improvement in the accuracy of case identification (Supplementary Materials).

### 2020-03-24 supp.tex
updates

@@
 To do so, we calculated the relative proportion of positive cases during each period (divided by the the between-period mean) and multiplied the daily number of reported cases by the relative proportions of the corresponding criterion.
-A sensitivity analysis showed that results are robust to these adjustments.
+Sensitivity analyses showed that results are robust to these adjustments.
 
-We then estimated time-dependent \emph{backward} onset-to-confirmation delay distributions from the partial line list.
-We asked: Given a cohort of individuals who were reported on the same day, what is the probability distribution of the onset-to-confirmation delay?
+We then estimated time-dependent \emph{backward} onset-to-confirmation delay distributions from the partial line list: Given a cohort of individuals who were reported on the same day, what is the probability distribution of the onset-to-confirmation delay?
 The backward delay distribution depends on changes in the number of symptomatic cases ---
@@
 
-To account for right-censoring in the reported cases, we also estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using the same negative-binomial regression model.
-Here, we asked: Given a cohort of individuals who became symptomatic on the same day, what is the probability distribution of the onset-to-confirmation delay?
+To account for right-censoring in the reported cases, we also estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using the same negative-binomial regression model: Given a cohort of individuals who became symptomatic on the same day, what is the probability distribution of the onset-to-confirmation delay?
 The forward delay distribution reflects the changes in the accuracy of case identification --- e.g., a decrease in the delay reflects improvement in accuracy.
@@
 For each combination of date of infection and a posterior sample of the forward delay distribution, we drew 1000 samples of incubation periods and onset-to-confirmation delays and calculated the median probability that an individual infected on a given day will be confirmed before March 16, 2020.
-We divided the daily number of
-
-We combined the forward delay distribution with the , we calculated the median probability that a
-
-to calculate the median probability that a case infected on a given day will be reported before March 16, 2020.
-Changes in the mean backward delay (\fref{dist}C) reflect changes in incidence of symptomatic cases, whereas the decrease in the mean forward delay (\fref{dist}D) reflects the improvement in the accuracy of case identification \citep{lai2020effect}.
+Finally, we divided the daily number of infected cases by the median probability that an infected case on a given day will be confirmed before Marh 16, 2020.
 

### 2020-03-24 degree_censor.R

@@ -1,3 +1,4 @@
+library(dplyr)
 library(rstan)
 
 load("report_delay_censor.rda")
@@ -19,17 +20,19 @@ for (i in 1:length(shape)) {
   ## pgamma(4.2, 25, 25/2.6) - pgamma(1.8, 25, 25/2.6)
   sdinc <- rgamma(1, 25, 25/2.6)
   
+  shapeinc <- meaninc^2/sdinc^2
+  
   meandelay <- estmean[,i]
-  shapedelay <- ee$shape[i]
+  shapedelay <- shape[i]
   
   dd <- data.frame(
-    date=seq.Date(from=as.Date("2020-01-10"), to=as.Date("2020-03-10"), by=1),
+    date=seq.Date(from=as.Date("2020-01-10"), to=as.Date("2020-03-16"), by=1),
     delay=meandelay
   ) %>%
     filter(date >= as.Date("2020-01-19"))
   
   censor <- sapply(1:nrow(dd), function(x) {
-    inc <- floor(rgamma(nsim, shape=shape, rate=shape/meaninc))
+    inc <- floor(rgamma(nsim, shape=shapeinc, rate=shapeinc/meaninc))
     
     symp <- as.Date("2020-01-19")-1+inc+x
     
@@ -42,7 +45,7 @@ for (i in 1:length(shape)) {
   })
   
   degree_censor[[i]] <- data.frame(
-    date=seq.Date(from=as.Date("2020-01-19"), to=as.Date("2020-03-10"), by=1),
+    date=seq.Date(from=as.Date("2020-01-19"), to=as.Date("2020-03-16"), by=1),
     censor=censor
   )
 }
@@ -54,6 +57,7 @@ degree_censor <- degree_censor %>%
     median=median(censor),
     lwr=quantile(censor, 0.025),
     upr=quantile(censor, 0.975)
-  )
+  ) %>%
+  filter(date <= as.Date("2020-03-10"))
 
 save("degree_censor", file="degree_censor.rda")

### 2020-03-24 korea.tex
ok done

@@
 We also did not distinguish local and imported cases, which may overestimate $\mathcal R_t$ \citep{thompson2019improved}.
-\swp{The analysis of the Seoul outbreak which accounts for imported cases is presented in Supplementary Materials?}
+We performed a separate analysis of the epidemic in Seoul that accounts for imported cases using line list provided by the Seoul Metropolitan Government; our qualitative conclusions remained robust (Supplementary Materials).
 

### 2020-03-24 supp.tex
ok done

@@
 For each combination of date of infection and a posterior sample of the forward delay distribution, we drew 1000 samples of incubation periods and onset-to-confirmation delays and calculated the median probability that an individual infected on a given day will be confirmed before March 16, 2020.
-Finally, we divided the daily number of infected cases by the median probability that an infected case on a given day will be confirmed before Marh 16, 2020.
+Finally, we divided the daily number of infected cases by the median probability that an infected case on a given day will be confirmed before March 16, 2020.
 
@@
 Vertical lines indicate the date on which testing criteria expanded.
-Boxplots (C--D) represent the observed delays.
+Box plots (C--D) represent the observed delays.
 Black lines and gray ribbons represent the median estimates of the mean delays and their associated 95\% credible intervals.
@@
 \end{figure}
+
+\pagebreak
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figure_R_t_seoul_linelist.pdf}
+\caption{
+\textbf{Comparison of effective reproduction number in Seoul using the number of reported cases by the KCDC and the line list provided by the Seoul Metropolitan Government.}
+Using line list, we reconstructed incidence for local $I_t^{\textrm{local}}$ and imported $I_t^{\textrm{imported}}$ cases separately based on the method described in the main text.
+Then, we estimated the time-dependent reproduction number via $\mathcal R_t = I_t^{\textrm{local}}/\sum_{k=1}^{14} I_{t-k} w_k$, where $I_t=I_t^{\textrm{local}}+I_t^{\textrm{imported}}$.
+We did not account for changes in testing criteria in this analysis.
+}
+\end{figure}

### 2020-03-24 figure_R_t_seoul_linelist.R

@@ -1,0 +1,119 @@
+library(readxl)
+library(ggplot2); theme_set(theme_bw())
+library(gridExtra)
+source("wquant.R")
+
+load("R_t_seoul_censor_detect.rda")
+load("R_t_seoul_censor_linelist.rda")
+
+R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
+
+rt_censor_detect <- R_t_seoul_censor_detect %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
+  summarize(
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
+    upr=wquant(IRt, weights=weight,0.975)
+  ) %>%
+  mutate(
+    group="main analysis"
+  )
+
+rt_censor_linelist <- R_t_seoul_censor_linelist %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  mutate(weight=R0prior(IRt)) %>%
+  summarize(
+    median=wquant(IRt, weights=weight,0.5),
+    lwr=wquant(IRt, weights=weight,0.025),
+    upr=wquant(IRt, weights=weight,0.975)
+  ) %>%
+  mutate(
+    group="analysis based on the line list"
+  )
+
+rt_all <- bind_rows(rt_censor_detect, rt_censor_linelist) %>%
+  mutate(
+    group=factor(group, level=c("main analysis",
+                                "analysis based on the line list"))
+  )
+
+covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=1, na="NA") %>%
+  mutate(
+    date_onset=as.Date(date_onset)
+  ) %>%
+  mutate(
+    import=!is.na(import_source),
+    import=ifelse(import, "imported", "local")
+  )
+
+reconstruct_case_detect <- reconstruct_censor_detect %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(case),
+    lwr=quantile(case, 0.025),
+    upr=quantile(case, 0.975)
+  ) %>%
+  mutate(
+    group="main analysis"
+  )
+
+reconstruct_case_linelist <- reconstruct_censor_linelist %>%
+  bind_rows(.id="sim") %>%
+  group_by(date) %>%
+  summarize(
+    median=median(case),
[…]

### 2020-03-24 korea.tex
cut down around 80 words

--- 2020-03-24 korea.tex
@@
 
-Since its first appearance in Wuhan, China, in December 2019 \citep{pneumonia}, COVID-19 disease has spread internationally, including to South Korea.
+Since its first appearance in Wuhan, China, in December 2019 \citep{pneumonia}, coronavirus disease (COVID-19) has spread internationally, including to South Korea.
 The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
@@
 The chains of transmission that began from this cluster distinguish the epidemic in South Korea from that in any other countries:
-As of March 18, 2020, 8,413 cases were confirmed, of which 60\% were related to the church and 27\% were in their 20s \citep{kcdc}.
-Nonetheless, South Korea's intensive testing using novel contact tracing techniques allowed rapid identification and isolation cases and reduction of onward transmission \citep{fergusonimpact, tracingkorea, science}.
-Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 by using metro traffic data to compare epidemics in two geographically separated major cities in South Korea.
+As of March 22, 2020, 8,897 cases were confirmed, of which 57\% were related to the church and 27\% were in their 20s \citep{kcdc}.
+South Korea's intensive testing using novel contact tracing techniques allowed rapid identification and isolation cases and reduction of onward transmission \citep{fergusonimpact, tracingkorea, science}.
+Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 in South Korea by using metro traffic data to compare epidemics in two major cities.
 
@@
 Vertical lines indicate Feb 18, 2020, when the first case was confirmed in Daegu.
+See Supplementary Materials for details.
 }
@@
 Between January 20--March 16, 2020, 6,083 cases from Daegu and 248 from Seoul were reported by the KCDC.
-Unlike the epidemic in Daegu, which is characterized by a single, large peak followed by a gradual decrease, the epidemic in Seoul consists of several small outbreaks (\fref{data}; see Supplementary Materials for details).
+The epidemic in Daegu is characterized by a single, large peak followed by a gradual decrease, whereas the epidemic in Seoul consists of several small outbreaks (\fref{data}).
 
@@
 We adjusted the number of reported cases to account for changes in testing criteria, which occurred 4 times between January 20--March 16, 2020.
-Then, we inferred onset-to-confirmation delay distributions from the partial line list;
-we combined the inferred delay distribution with previously estimated incubation period distribution (Table 1) to obtain probability distributions for date of infection for each reported case.
+Then, we inferred onset-to-confirmation delay distributions from the partial line list and combined them with previously estimated incubation period distribution (Table 1) to obtain probability distributions for date of infection for each reported case.
 We accounted for right-censoring by dividing the daily incidence by the probability that a case infected on a given day would have been reported before March 16, 2020.
@@
 The initial decrease in $\mathcal R_t$ may reflect behavior change within the church; the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.
-Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
+% Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
 
@@
 
-While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.90$; 95\% CI: 0.79--0.95) and Seoul ($r=0.76$; 95\% CI: 0.59--0.87), these correlations are conflated by time trends, and are also likely conflated by other measures that could have affected $\mathcal R_t$ --
-in particular, the increase in the number of tests as well as the improvement in the accuracy of case identification (Supplementary Materials).
+While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.90$; 95\% CI: 0.79--0.95) and Seoul ($r=0.76$; 95\% CI: 0.59--0.87), these correlations are conflated by time trends, and are also likely conflated by other measures that could have affected $\mathcal R_t$ ---
+in particular, the contact tracing efforts \citep{fergusonimpact, tracingkorea, science}.
 We do not find clear signatures of lagged correlation between $\mathcal R_t$ and traffic volume (Supplementary Materials).
@@
 Our study is not without limitations.
-In particular, we did not account for differences in the delay distributions or changes in the number of tests among cities.
+We did not account for differences in the delay distributions or changes in the number of tests among cities.
 The intensity of intervention is likely to vary across regions given that majority of COVID-19 cases in South Korea were reported from Daegu.
-We did not have sufficient data to account for these factors and therefore applied identical methods for adjusting the time series and estimating $\mathcal R_t$ across 4 regions.
+We did not have sufficient data to account for these factors.
 Nonetheless, the robustness of our findings is supported by the sensitivity analyses (Supplementary Materials).
-We also did not distinguish local and imported cases, which may overestimate $\mathcal R_t$ \citep{thompson2019improved}.
-\swp{The analysis of the Seoul outbreak which accounts for imported cases is presented in Supplementary Materials?}
+We were also unable to distinguish local and imported cases, which may overestimate $\mathcal R_t$ \citep{thompson2019improved}.
+We were able to perform a separate analysis for Seoul that accounts for imported cases using line list provided by the Seoul Metropolitan Government; our qualitative conclusions remained robust (Supplementary Materials).
 
@@
 The 80\% decrease in traffic volume suggests that the strength of social distancing in Daegu may be comparable to that in Wuhan, China \citep{lai2020effect}.
-However, this analysis is indirect; we are not able to directly estimate the effect of social distancing on epidemiological dynamics.
-Moreover, other measures, such as intensive testing of core transmission groups and school closure, are also likely to have affected the changes in $\mathcal R_t$ \citep{kcdc}.
+However, we are not able to directly estimate the effect of social distancing on epidemiological dynamics.
+Other measures, such as intensive testing of core transmission groups and school closure, are also likely to have affected the changes in $\mathcal R_t$ \citep{kcdc}.
 Future studies should consider quantifying contributions of different measures in preventing the spread.

### 2020-03-25 korea.tex
k remove comment
tweak
Very light changes

@@
 
-To estimate the time-dependent reproduction number $R_t$ (the average number of secondary cases caused by an individual infected at time $t$ \citep{fraser2007estimating}), we first estimated daily incidence of infection from the daily number of reported cases by the KCDC \citep{kcdc}.
+To estimate the time-dependent reproduction number $R_t$ (the average number of secondary cases caused by an average individual, given conditions at time $t$ \citep{fraser2007estimating}), we first estimated daily incidence of infection from the daily number of reported cases by the KCDC \citep{kcdc}.
 We adjusted the number of reported cases to account for changes in testing criteria, which occurred 4 times between January 20--March 16, 2020.
@@
 
-\jd{Me to change backward/forward R description above, Daniel to approve.}
 We estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
@@
 \fref{eff} compares the reconstructed incidence (A,B) and estimates of $\mathcal R_t$ (C,D) in Daegu and Seoul.
-In Daegu, incidence peaked shortly after the first case was confirmed and decreased (\fref{eff}A).
+In Daegu, incidence peaked shortly after the first case was confirmed and then decreased (\fref{eff}A).
 Likewise, the estimates of $\mathcal R_t$ gradually decrease and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}C).
 The initial decrease in $\mathcal R_t$ may reflect behavior change within the church; the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.
-% Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
 
-In Seoul, incidence decreases slightly as traffic decreased (\fref{eff}B) but estimates of $\mathcal R_t$ remain around 1 throughout (\fref{eff}D).
-Our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less effective.
+In Seoul, estimates of $\mathcal R_t$ decreases slightly but remain around 1 (\fref{eff}D).
+Our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less sharp.
 Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.

### 2020-03-25 korea.tex
super minor tweaks

@@
 
-\section*{Abstract - 75 word limit; currently 67.}
+\section*{Abstract}
 
@@
 The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
-In February, the disease spread rapidly within a church community in the city of Daegu.
+In February, the disease spread rapidly within a church community in the city of Daegu \citep{kcdc}.
 The chains of transmission that began from this cluster distinguish the epidemic in South Korea from that in any other countries:
-As of March 22, 2020, 8,897 cases were confirmed, of which 57\% were related to the church and 27\% were in their 20s \citep{kcdc}.
+As of March 24, 2020, 9,037 cases were confirmed, of which 56\% were related to the church and 27\% were in their 20s \citep{kcdc}.
 South Korea's intensive testing using novel contact tracing techniques allowed rapid identification and isolation cases and reduction of onward transmission \citep{fergusonimpact, tracingkorea, science}.
@@
 The initial decrease in $\mathcal R_t$ may reflect behavior change within the church; the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.
+Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
 
@@
 Our analysis reveals that the epidemic may still persist in other regions, including Seoul and Gyeonggi-do;
-reports from Seoul and Gyeonggi-do (around 10 cases almost every day between March 11--22, 2020) provide further support for our conclusion \citep{kcdc}.
+reports from Seoul and Gyeonggi-do (around 10 cases almost every day between March 11--24, 2020) provide further support for our conclusion \citep{kcdc}.
 Unless the reproduction number can be reduced below 1 in all regions, small outbreaks may continue to occur in South Korea.

### 2020-03-25 supp.tex
supp
super minor tweaks

@@
 
-We then estimated time-dependent \emph{backward} onset-to-confirmation delay distributions from the partial line list: Given a cohort of individuals who were reported on the same day, what is the probability distribution of the onset-to-confirmation delay?
+We then estimated time-dependent \emph{backward} onset-to-confirmation delay distributions from the partial line list: Given a cohort of infected individuals who were confirmed on the same day, what is the probability distribution of the onset-to-confirmation delay?
 The backward delay distribution depends on changes in the number of symptomatic cases ---
@@
 We assumed weakly informative priors on the fixed effects: normal distributions with mean of 0 and standard deviation of 2;
-these distributions are priors on link scale.
+note that these distributions are priors on link scale.
 
@@
 This allowed us to obtain posterior samples of possible infection dates for each case,
-which was then converted into posterior samples of incidence time series.
+which were then converted into posterior samples of incidence time series.
 
-To account for right-censoring in the reported cases, we also estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using the same negative-binomial regression model: Given a cohort of individuals who became symptomatic on the same day, what is the probability distribution of the onset-to-confirmation delay?
+To account for right-censoring in the reported cases, we also estimated time-dependent \emph{forward} onset-to-confirmation delay distribution using the same negative-binomial regression model: Given a cohort of infected individuals who became symptomatic on the same day, what is the probability distribution of the onset-to-confirmation delay?
 The forward delay distribution reflects the changes in the accuracy of case identification --- e.g., a decrease in the delay reflects improvement in accuracy.
@@
 To estimate the forward delay distribution, we modified the stan code from the negative-binomial regression that we used to infer the backward delay distribution to account for right-censoring (in the observed delays) and ran the code using the \texttt{rstan} package \citep{rstan}.
-In particular, we modified the likelihood such that given a delay of $x_i$ days, symptom onset day $t_i$ and the day of measurement of $k$, the likelihood of observing the delay is given by:
+In particular, we modified the likelihood of the negative-binomial regression such that given a delay of $x_i$ days, symptom onset day $t_i$ and the day of measurement of $t_{\tiny\textrm{max}}$, the likelihood of observing the delay is given by:
 \begin{equation}
-\frac{f(x_i|\mu(t_i), \theta)}{F(k-t_i|\mu(t_i), \theta)},
+\frac{f(x_i|\mu(t_i), \theta)}{F(t_{\tiny\textrm{max}}-t_i|\mu(t_i), \theta)},
 \end{equation}
 where $f$ is the negative binomial distribution with time-dependent mean $\mu(t_i)$ and dispersion parameter $\theta$.
-This likelihood accounts for the fact that the delay between symptom onset and confirmation cannot be longer than $k-t_i$ (otherwise, the case will be reported after the date of measurement).
+This likelihood accounts for the fact that the delay between symptom onset and confirmation cannot be longer than $t_{\tiny\textrm{max}}-t_i$ (otherwise, the case will be reported after $t_{\tiny\textrm{max}}$).
 Convergence is assessed by the lack of warning messages from the \texttt{rstan} package \citep{rstan}.
@@
 For each combination of date of infection and a posterior sample of the forward delay distribution, we drew 1000 samples of incubation periods and onset-to-confirmation delays and calculated the median probability that an individual infected on a given day will be confirmed before March 16, 2020.
-Finally, we divided the daily number of infected cases by the median probability that an infected case on a given day will be confirmed before March 16, 2020.
+Finally, we divided the daily number of infected cases by the median probability this probability.
+We used the reconstructed incidence time series to estimate $\mathcal R_t$.
 
@@
 We did not account for changes in testing criteria in this analysis.
+The line lists were obtained from \url{http://news.seoul.go.kr/welfare/archives/513105}.
 }

### 2020-03-25 supp.tex
consistency

@@
 \caption{
-\textbf{Comparison of effective reproduction number and the daily number of reported cases in Daegu, Seoul, Gyeongsangbuk-do, and Gyeonggi-do.}
+\textbf{Comparison of time-dependent reproduction number and the daily number of reported cases in Daegu, Seoul, Gyeongsangbuk-do, and Gyeonggi-do.}
 }
@@
 \caption{
-\textbf{Comparison of effective reproduction number in Seoul using the number of reported cases by the KCDC and the line list provided by the Seoul Metropolitan Government.}
+\textbf{Comparison of time-dependent reproduction number in Seoul using the number of reported cases by the KCDC and the line list provided by the Seoul Metropolitan Government.}
 Using line list, we reconstructed incidence for local $I_t^{\textrm{local}}$ and imported $I_t^{\textrm{imported}}$ cases separately based on the method described in the main text.

### 2020-03-25 figure_R_t_all.R

@@ -72,7 +72,7 @@ g1 <- ggplot(filter(R_t_all, region=="Daegu")) +
   geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.3) +
   geom_line(aes(date, median)) +
   scale_x_date("Date") +
-  scale_y_continuous("Effective reproduction number", limits=c(0, NA), expand=c(0, 0),
+  scale_y_continuous("Time-dependent reproduction number", limits=c(0, NA), expand=c(0, 0),
                      sec.axis = sec_axis(~ .*150, name = "Number of reported cases")) +
   facet_wrap(~region, scale="free") +
   theme(
@@ -90,7 +90,7 @@ g2 <- ggplot(filter(R_t_all, region=="Gyeongsangbuk-do")) +
   geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.3) +
   geom_line(aes(date, median)) +
   scale_x_date("Date") +
-  scale_y_continuous("Effective reproduction number", limits=c(0, NA), expand=c(0, 0),
+  scale_y_continuous("Time-dependent reproduction number", limits=c(0, NA), expand=c(0, 0),
                      sec.axis = sec_axis(~ .*30, name = "Number of reported cases")) +
   facet_wrap(~region, scale="free") +
   theme(
@@ -108,7 +108,7 @@ g3 <- ggplot(filter(R_t_all, region=="Seoul")) +
   geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.3) +
   geom_line(aes(date, median)) +
   scale_x_date("Date") +
-  scale_y_continuous("Effective reproduction number", limits=c(0, NA), expand=c(0, 0),
+  scale_y_continuous("Time-dependent reproduction number", limits=c(0, NA), expand=c(0, 0),
                      sec.axis = sec_axis(~ .*8, name = "Number of reported cases")) +
   facet_wrap(~region, scale="free") +
   theme(
@@ -126,7 +126,7 @@ g4 <- ggplot(filter(R_t_all, region=="Gyeonggi-do")) +
   geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.3) +
   geom_line(aes(date, median)) +
   scale_x_date("Date") +
-  scale_y_continuous("Effective reproduction number", limits=c(0, NA), expand=c(0, 0),
+  scale_y_continuous("Time-dependent reproduction number", limits=c(0, NA), expand=c(0, 0),
[…]

### 2020-03-25 figure_R_t_daegu.R

@@ -102,7 +102,7 @@ g2 <- ggplot(rt_all) +
   scale_color_manual(values=c(2,1)) +
   scale_fill_manual(values=c(2,1)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-19", "2020-03-14"))) +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 7), expand=c(0, 0),
+  scale_y_continuous("Time-dependent reproduction number", limits=c(0, 7), expand=c(0, 0),
                      breaks=c(0, 3, 6)) +
   theme(
     panel.grid = element_blank(),

### 2020-03-25 figure_R_t_gyeongbuk.R

@@ -101,7 +101,7 @@ g2 <- ggplot(rt_all) +
   scale_color_manual(values=c(2, 1)) +
   scale_fill_manual(values=c(2, 1)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-19", "2020-03-14"))) +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 7), expand=c(0, 0),
+  scale_y_continuous("Time-dependent reproduction number", limits=c(0, 7), expand=c(0, 0),
                      breaks=c(0, 2, 4, 6)) +
   theme(
     panel.grid = element_blank(),

### 2020-03-25 figure_R_t_gyeonggi.R

@@ -101,7 +101,7 @@ g2 <- ggplot(rt_all) +
   scale_color_manual(values=c(2, 1)) +
   scale_fill_manual(values=c(2, 1)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-19", "2020-03-14"))) +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 8), expand=c(0, 0)) +
+  scale_y_continuous("Time-dependent reproduction number", limits=c(0, 8), expand=c(0, 0)) +
   theme(
     panel.grid = element_blank(),
     panel.border = element_blank(),

### 2020-03-25 figure_R_t_seoul_linelist.R

@@ -104,7 +104,7 @@ g2 <- ggplot(rt_all) +
   scale_color_manual(values=c(1, 2)) +
   scale_fill_manual(values=c(1, 2)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-19", "2020-03-14"))) +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 8), expand=c(0, 0)) +
+  scale_y_continuous("Time-dependent reproduction number", limits=c(0, 8), expand=c(0, 0)) +
   ggtitle("B") +
   theme(
     panel.grid = element_blank(),

### 2020-03-25 figure_compare_R_t.R

@@ -201,7 +201,7 @@ g3 <- ggplot(rt_daegu) +
   scale_color_manual(values=c(1, 2, 4)) +
   scale_fill_manual(values=c(1, 2, 4)) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-20", "2020-03-16"))+c(0,0.5)) +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 8), expand=c(0, 0),
+  scale_y_continuous("Time-dependent reproduction number", limits=c(0, 8), expand=c(0, 0),
                      sec.axis = sec_axis(~ .*1/6, name = "(Daily traffic, 2020)/(Mean daily traffic, 2017 - 2019)")) +
   ggtitle("C. Daegu") +
   theme(
@@ -226,7 +226,7 @@ g4 <- ggplot(rt_seoul) +
   geom_vline(xintercept=as.Date("2020-02-18"), lty=2) +
   scale_x_date("Date", expand=c(0, 0), limits=as.Date(c("2020-01-20", "2020-03-16"))+c(0,0.5)) +
   ggtitle("D. Seoul") +
-  scale_y_continuous("Effective reproduction number", limits=c(0, 8), expand=c(0, 0),
+  scale_y_continuous("Time-dependent reproduction number", limits=c(0, 8), expand=c(0, 0),
                      sec.axis = sec_axis(~ ./6, name = "(Daily traffic, 2020)/(Mean daily traffic, 2017 - 2019)")) +
   theme(
     panel.grid = element_blank(),
@@ -243,3 +243,17 @@ g4 <- ggplot(rt_seoul) +
 gtot <- arrangeGrob(g1, g2, g3, g4, nrow=2, widths=c(1.1, 1))
 
 ggsave("figure_compare_R_t.pdf", gtot, width=8, height=8)
+
+g1 <- ggplot(bind_rows(mutate(daegu_merge, region="Daegu"), mutate(seoul_merge, region="Seoul"))) +
+  geom_point(aes(traffic, median, shape=region, col=region), size=3) +
+  scale_x_continuous("Normalized traffic") +
+  scale_y_continuous("Time-dependent reproduction number") +
+  theme(
+    panel.grid = element_blank(),
+    panel.border = element_blank(),
+    axis.line = element_line(),
+    legend.position = c(0.1, 0.9),
+    legend.title = element_blank()
+  )
+
+ggsave("traffic.pdf", g1, width=6, height=6)

### 2020-03-25 korea.tex
done??

@@
 Daily number of reported cases in each geographic region was transcribed from press releases by Korea Centers for Disease Control and Prevention (KCDC) \cite{kcdc}.
-Partial line lists were transcribed from press releases from KCDC and various local and provincial governments.
+Partial line lists were transcribed from press releases from KCDC and various local and provincial governments \cite{seoul, busan, gwangju, ulsan, gyeonggi, daegu}.
 All data and original reports are stored in a publicly available GitHub repository.
@@
 
-While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.90$; 95\% CI: 0.79--0.95) and Seoul ($r=0.76$; 95\% CI: 0.59--0.87), these correlations are conflated by time trends, and are also likely conflated by other measures that could have affected $\mathcal R_t$ ---
-in particular, the contact tracing efforts \citep{fergusonimpact, tracingkorea, science}.
+While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.90$; 95\% CI: 0.79--0.95) and Seoul ($r=0.76$; 95\% CI: 0.59--0.87), these correlations are conflated by time trends, and are also likely conflated by other measures that could have affected $\mathcal R_t$.
 We do not find clear signatures of lagged correlation between $\mathcal R_t$ and traffic volume (Supplementary Materials).
@@
 The South Korean experience with COVID-19 provides evidence that epidemics can be suppressed with less extreme measures than those taken by China \citep{kickbusch2020response}.
+It demonstrates the necessity of prompt identification and isolation of cases in preventing further spread \citep{fergusonimpact, tracingkorea, science}.
 Our analysis reveals potential roles of social distancing in mitigating the COVID-19 epidemic in South Korea.

### 2020-03-25 korea.tex
separate title file

@@
 
-\begin{flushleft}{
-	\Large
-	\textbf\newline{
-		Potential roles of social distancing in mitigating the spread of coronavirus disease 2019 (COVID-19) in South Korea
-	}
-}
-\newline
-\\
-Sang Woo Park\textsuperscript{1,*}
-Kaiyuan Sun\textsuperscript{2}
-C\'ecile Viboud\textsuperscript{2}
-Bryan T.\ Grenfell\textsuperscript{1,2,3}
-Jonathan Dushoff\textsuperscript{4,5,6}
-\\
-\bigskip
-\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
-\\
-\textbf{2} Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
-\\
-\textbf{3} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
-\\
-\textbf{4} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
-\\
-\textbf{5} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
-\\
-\textbf{6} Department of Biology, McMaster University, Hamilton, ON, Canada
-\\
-\bigskip
-
-*Corresponding author: swp2@princeton.edu\\
-This article does not necessarily represent the views of the NIH or the US government.
-\end{flushleft}
-
-\section*{Abstract}
-
-On January 20, 2020, the first COVID-19 case was confirmed in South Korea.
-After a rapid outbreak, the number of incident cases has been consistently decreasing since early March;
-this decrease has been widely attributed to its intensive testing.
-We report here on the likely role of social distancing in reducing transmission in South Korea.
-Our analysis suggests that transmission may still be persisting in some regions.
-
-\section*{Contribution}
-
-Data collection: SWP; conceptualization: SWP, CV; analysis: SWP; first draft: SWP, JD. All authors contributed to the writing and approval of the final report.
-
-\section*{Data availability}
-
-All transcribed data and original reports used in this analysis are stored in a publicly available GitHub repository: \url{https://github.com/parksw3/COVID19-Korea}.
-All code and traffic data are stored in a separate GitHub repository: \url{https://github.com/parksw3/Korea-analysis}.
+\input{title.tex}
 

### 2020-03-25 title.tex
separate title file

\begin{flushleft}{
	\Large
	\textbf\newline{
		Potential roles of social distancing in mitigating the spread of coronavirus disease 2019 (COVID-19) in South Korea
	}
}
\newline
\\
Sang Woo Park\textsuperscript{1,*}
Kaiyuan Sun\textsuperscript{2}
C\'ecile Viboud\textsuperscript{2}
Bryan T.\ Grenfell\textsuperscript{1,2,3}
Jonathan Dushoff\textsuperscript{4,5,6}
\\
\bigskip
\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
\\
\textbf{2} Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
\\
\textbf{3} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
\\
\textbf{4} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
\\
\textbf{5} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
\\
\textbf{6} Department of Biology, McMaster University, Hamilton, ON, Canada
\\
\bigskip

*Corresponding author: swp2@princeton.edu\\
This article does not necessarily represent the views of the NIH or the US government.
\end{flushleft}

\section*{Abstract}

On January 20, 2020, the first COVID-19 case was confirmed in South Korea.
After a rapid outbreak, the number of incident cases has been consistently decreasing since early March;
this decrease has been widely attributed to its intensive testing.
We report here on the likely role of social distancing in reducing transmission in South Korea.
Our analysis suggests that transmission may still be persisting in some regions.

\section*{Contribution}

Data collection: SWP; conceptualization: SWP, CV; analysis: SWP; first draft: SWP, JD. All authors contributed to the writing and approval of the final report.

\section*{Data availability}

All transcribed data and original reports used in this analysis are stored in a publicly available GitHub repository: \url{https://github.com/parksw3/COVID19-Korea}.
All code and traffic data are stored in a separate GitHub repository: \url{https://github.com/parksw3/Korea-analysis}.

### 2020-03-25 korea.tex
one more citation

@@
 Daily number of reported cases in each geographic region was transcribed from press releases by Korea Centers for Disease Control and Prevention (KCDC) \cite{kcdc}.
-Partial line lists were transcribed from press releases from KCDC and various local and provincial governments \cite{seoul, busan, gwangju, ulsan, gyeonggi, daegu}.
+Partial line lists were transcribed from press releases from KCDC and various local and provincial governments \cite{seoul, busan, gwangju, ulsan, gyeonggi, daegu, gyeongnam}.
 All data and original reports are stored in a publicly available GitHub repository.

### 2020-03-26 korea.tex
prepraring

--- 2020-03-25 korea.tex
@@
 We analyzed epidemiological data describing the COVID-19 outbreak in South Korea between January 20--March 16, 2020.
-Daily number of reported cases in each geographic region was transcribed from press releases by Korea Centers for Disease Control and Prevention (KCDC) \cite{kcdc}.
-Partial line lists were transcribed from press releases from KCDC and various local and provincial governments \cite{seoul, busan, gwangju, ulsan, gyeonggi, daegu}.
+Daily number of reported cases in each geographic region was transcribed from press releases by the Korea Centers for Disease Control and Prevention (KCDC) \cite{kcdc}.
+Partial line lists were translated and transcribed from press releases by the KCDC and various local and provincial governments \cite{seoul, busan, gwangju, ulsan, gyeonggi, daegu, gyeongnam}.
 All data and original reports are stored in a publicly available GitHub repository.

### 2020-03-26 supp.tex
done??
separate title file
prepraring

@@
 Since then, the daily number of reported cases reflects the number of reported cases within the last 24 hours (12 AM to 12 AM).
-The daily number of reported cases by the KCDC may be different from the reports by each city's government as some cases may be transferred after they are confirmed.
+The number of negative cases was not reported on January 25 and 31, 2020; we took the average of cumulative negative cases from one day before and after these dates instead to impute missing values.
+The daily number of reported cases by the KCDC may be slightly different from the reports by each city's government as some cases may be transferred after they are confirmed.
+The sum of daily number of reported cases by the KCDC may be also slightly different from the cumulative number of cases reported the KCDC because it does not reflect possible location changes of the confirmed cases.
 
@@
 \begin{figure}[!ht]
+\includegraphics[width=\textwidth]{traffic.pdf}
+\caption{
+\textbf{Scatter plot of the normalized traffic volume and the median estimates of $\mathcal R_t$.}
+}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[!ht]
 \includegraphics[width=\textwidth]{figure_cross.pdf}

### 2020-03-26 supp_euro.tex
prepraring

--- 2020-03-25 korea.tex
@@
 
-\input{title.tex}
+\begin{flushleft}{
+	\Large
+	\textbf\newline{
+		Potential roles of social distancing in mitigating the spread of coronavirus disease 2019 (COVID-19) in South Korea -- Supplementary Materials
+	}
+}
+\newline
+\\
+Sang Woo Park\textsuperscript{1,*}
+Kaiyuan Sun\textsuperscript{2}
+C\'ecile Viboud\textsuperscript{2}
+Bryan T.\ Grenfell\textsuperscript{1,2,3}
+Jonathan Dushoff\textsuperscript{4,5,6}
+\\
+\bigskip
+\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
+\\
+\textbf{2} Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
+\\
+\textbf{3} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
+\\
+\textbf{4} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
+\\
+\textbf{5} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
+\\
+\textbf{6} Department of Biology, McMaster University, Hamilton, ON, Canada
+\\
+\bigskip
 
-\pagebreak
-
-\section{Background}
-
-Since its first appearance in Wuhan, China, in December 2019 \citep{pneumonia}, coronavirus disease (COVID-19) has spread internationally, including to South Korea.
-The first COVID-19 case in South Korea was confirmed on January 20, 2020, from a traveling resident of Wuhan, China \citep{kcdc}.
-In February, the disease spread rapidly within a church community in the city of Daegu \citep{kcdc}.
-The chains of transmission that began from this cluster distinguish the epidemic in South Korea from that in any other countries:
-As of March 24, 2020, 9,037 cases were confirmed, of which 56\% were related to the church and 27\% were in their 20s \citep{kcdc}.
-South Korea's intensive testing using novel contact tracing techniques allowed rapid identification and isolation cases and reduction of onward transmission \citep{fergusonimpact, tracingkorea, science}.
-Here, we describe potential roles of social distancing in mitigating the spread of COVID-19 in South Korea by using metro traffic data to compare epidemics in two major cities.
-
-\section{Data description}
-
-We analyzed epidemiological data describing the COVID-19 outbreak in South Korea between January 20--March 16, 2020.
-Daily number of reported cases in each geographic region was transcribed from press releases by Korea Centers for Disease Control and Prevention (KCDC) \cite{kcdc}.
-Partial line lists were transcribed from press releases from KCDC and various local and provincial governments \cite{seoul, busan, gwangju, ulsan, gyeonggi, daegu}.
-All data and original reports are stored in a publicly available GitHub repository.
-
-\begin{figure}[!h]
-\includegraphics[width=\textwidth]{figure_compare_report.pdf}
-\caption{
-\textbf{Comparison of epidemiological and traffic data from Daegu and Seoul.}
-Solid lines represent the daily metro traffic volume in 2020 (red) and mean daily traffic volume between 2017--2019 (black).
-Daily traffic from previous years have been shifted by 1--3 days to align day of the weeks.
-Vertical lines indicate Feb 18, 2020, when the first case was confirmed in Daegu.
-See Supplementary Materials for details.
-}
-\label{fig:data}
-\end{figure}
-
-We compared epidemiological dynamics of COVID-19 from two cities in which the largest number of COVID-19 cases have been reported: Daegu and Seoul.
-Between January 20--March 16, 2020, 6,083 cases from Daegu and 248 from Seoul were reported by the KCDC.
-The epidemic in Daegu is characterized by a single, large peak followed by a gradual decrease, whereas the epidemic in Seoul consists of several small outbreaks (\fref{data}).
-
-Daily metro traffic in Daegu and Seoul between 2017--2020 was obtained from \url{data.go.kr} and \url{data.seoul.go.kr}, respectively.
-We tabulated the total number of individuals who accessed the subway or monorail (using Seoul lines 1--9, and Daegu lines 1--3; \fref{data}).
-Soon after the first church-related case was confirmed in Daegu on Feb 18, 2020, the daily traffic volume decreased by about 80\% and 50\% in Daegu and Seoul, respectively.
-
-%% https://www.data.go.kr/dataset/15002503/fileData.do
-%% https://data.seoul.go.kr/dataList/OA-12914/S/1/datasetView.do
-
-\section{Trends in time-dependent reproduction number and traffic volume}
-
-To estimate the time-dependent reproduction number $R_t$ (the average number of secondary cases caused by an average individual, given conditions at time $t$ \citep{fraser2007estimating}), we first estimated daily incidence of infection from the daily number of reported cases by the KCDC \citep{kcdc}.
-We adjusted the number of reported cases to account for changes in testing criteria, which occurred 4 times between January 20--March 16, 2020.
-Then, we inferred onset-to-confirmation delay distributions from the partial line list and combined them with previously estimated incubation period distribution (Table 1) to obtain probability distributions for date of infection for each reported case.
-We accounted for right-censoring by dividing the daily incidence by the probability that a case infected on a given day would have been reported before March 16, 2020.
-Implementation details are provided in the Supplementary Materials.
-
-\begin{table}[t]
-\begin{center}
-\small
-\begin{tabular}{p{4cm}|p{4.5cm}|p{4.7cm}|l}
- & Parameterization & Priors & Source \\
-\hline
-Incubation period\newline distribution & $\mathrm{Gamma}(\mu_I, \mu_I^2/\sigma^2)$ & $\mu_I\sim \mathrm{Gamma}(6.5\,\textrm{days}, 145)$\newline$\sigma\sim \mathrm{Gamma}(2.6, 25)$ & \citep{backer2020incubation} \\
-\hline
-Generation-interval\newline distribution & $\mathrm{NegativeBinomial}(\mu_G, \theta)$ & $\mu_G\sim\mathrm{Gamma}(5\,\textrm{days},62)$\newline$\theta\sim\mathrm{Gamma}(5,20)$ & \citep{ferretti2020quantifying, ganyani2020estimating} \\
-\hline
-\end{tabular}
-\end{center}
-\caption{
-\textbf{Assumed incubation and generation-interval distributions.}
-Gamma distributions are parameterized using its mean and shape.
-Negative binomial distributions are parameterized using its mean and dispersion.
-Priors are chosen such that the 95\% quantiles of prior means and standard deviations are consistent with previous estimates.
-}
-\end{table}
-
-We estimated the time-dependent reproduction number using the renewal equation with a 14-day sliding window \citep{fraser2007estimating}:
-\begin{equation}
-\mathcal R_t = \frac{I_t}{\sum_{k=1}^{14} I_{t-k} w_k},
-\end{equation}
-where $I_t$ is the reconstructed incidence time series (i.e., the number of infected cases on day $t$) and $w_k$ is the generation-interval distribution randomly drawn from a prior distribution (Table 1).
-We weighted each sample of $\mathcal R_t$ by a gamma probability distribution with a mean of 2.6 and a standard deviation of 2 to reflect prior knowledge \citep{tempvar} and took weighted quantiles to calculate the medians and associated 95\% credible intervals.
-We estimated $\mathcal R_t$ between February 2, 2020 (14 days after the first confirmed case was imported) and March 10, 2020 (after this point the effects of censoring are too strong for reliable estimates).
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figure_compare_R_t.pdf}
-\caption{
-\textbf{Comparison of reconstructed incidence and time-dependent reproduction number in Daegu and Seoul.}
-Black lines and gray ribbons represent the median estimates of reconstructed incidence (A,B) and $R_t$ (C,D) and their corresponding 95\% credible intervals.
-Bar plots show the number of reported cases.
-Red lines represent the normalized traffic volume.
-Vertical lines indicate Feb 18, 2020, when the first case was confirmed in Daegu.
-}
-\label{fig:eff}
-\end{figure}
-
-\fref{eff} compares the reconstructed incidence (A,B) and estimates of $\mathcal R_t$ (C,D) in Daegu and Seoul.
-In Daegu, incidence peaked shortly after the first case was confirmed and then decreased (\fref{eff}A).
-Likewise, the estimates of $\mathcal R_t$ gradually decrease and eventually drop below 1 about a week after the reporting of its first case, coinciding with the decrease in the metro traffic volume (\fref{eff}C).
-The initial decrease in $\mathcal R_t$ may reflect behavior change within the church; the first confirmed case in Daegu became symptomatic on February 7, 2020, and visited the church on February 9 and 16, 2020 \citep{kcdc}.
-Our estimates of $\mathcal R_t$ for Daegu are consistent with the estimates of $\mathcal R_t$ for South Korea by Abbott \textit{et al.} \cite{tempvar} --- their estimates drop below 1 slightly later because they rely on number of symptomatic cases instead.
-
-In Seoul, estimates of $\mathcal R_t$ decreases slightly but remain around 1 (\fref{eff}D).
-Our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less sharp.
-Stronger distancing or further intervention will be necessary to reduce $\mathcal R_t$ below 1.
-
-While we find clear, positive correlations between the normalized traffic and the median estimates of $\mathcal R_t$ in both Daegu ($r=0.90$; 95\% CI: 0.79--0.95) and Seoul ($r=0.76$; 95\% CI: 0.59--0.87), these correlations are conflated by time trends, and are also likely conflated by other measures that could have affected $\mathcal R_t$.
-We do not find clear signatures of lagged correlation between $\mathcal R_t$ and traffic volume (Supplementary Materials).
-Similar patterns in the estimates of $\mathcal R_t$ are found in directly surrounding provinces (Gyeongsangbuk-do and Gyeonggi-do), providing support for the robustness of our analysis (Supplementary Materials).
-
-\section{Discussion}
-
-The South Korean experience with COVID-19 provides evidence that epidemics can be suppressed with less extreme measures than those taken by China \citep{kickbusch2020response}.
-It demonstrates the necessity of prompt identification and isolation of cases in preventing further spread \citep{fergusonimpact, tracingkorea, science}.
-Our analysis reveals potential roles of social distancing in mitigating the COVID-19 epidemic in South Korea.
-Even though social distancing alone may not be able to fully prevent the spread of the disease, its ability to flatten the epidemic curve (cf.
-\fref{eff}B,D) reduces burden for healthcare system and provides time to plan for the future \citep{anderson2020will}.
-
-Our study is not without limitations.
-We did not account for differences in the delay distributions or changes in the number of tests among cities.
-The intensity of intervention is likely to vary across regions given that majority of COVID-19 cases in South Korea were reported from Daegu.
-We did not have sufficient data to account for these factors.
-Nonetheless, the robustness of our findings is supported by the sensitivity analyses (Supplementary Materials).
-We were also unable to distinguish local and imported cases, which may overestimate $\mathcal R_t$ \citep{thompson2019improved}.
-We were able to perform a separate analysis for Seoul that accounts for imported cases using line list provided by the Seoul Metropolitan Government; our qualitative conclusions remained robust (Supplementary Materials).
-
-Our analysis focused on comparing metro traffic, which serves as a proxy for the degree of social distancing, with epidemiological dynamics in two cities.
-The 80\% decrease in traffic volume suggests that the strength of social distancing in Daegu may be comparable to that in Wuhan, China \citep{lai2020effect}.
-However, we are not able to directly estimate the effect of social distancing on epidemiological dynamics.
-Other measures, such as intensive testing of core transmission groups and school closure, are also likely to have affected the changes in $\mathcal R_t$ \citep{kcdc}.
-Future studies should consider quantifying contributions of different measures in preventing the spread.
-
-Finally, our study highlights the importance of considering geographical heterogeneity in estimating epidemic potential.
-The recent decrease in the number of reported cases in South Korea is driven by the sharp decrease in Daegu.
-Our analysis reveals that the epidemic may still persist in other regions, including Seoul and Gyeonggi-do;
-reports from Seoul and Gyeonggi-do (around 10 cases almost every day between March 11--24, 2020) provide further support for our conclusion \citep{kcdc}.
-Unless the reproduction number can be reduced below 1 in all regions, small outbreaks may continue to occur in South Korea.
-
-\pagebreak
-
-\bibliography{korea}
-
+*Corresponding author: swp2@princeton.edu\\
+This article does not necessarily represent the views of the NIH or the US government.
+\end{flushleft}
 \pagebreak
@@
 
+\bibliography{korea}
+
 \end{document}

### 2020-03-26 title_euro.tex
prepraring

--- 2020-03-26 supp_euro.tex
@@
 	\textbf\newline{
-		Potential roles of social distancing in mitigating the spread of coronavirus disease 2019 (COVID-19) in South Korea -- Supplementary Materials
+		Potential roles of social distancing in mitigating the spread of coronavirus disease 2019 (COVID-19) in South Korea
 	}
@@
 \end{flushleft}
-\pagebreak
 
-\input{supp.tex}
+\section*{Abstract}
 
-\bibliography{korea}
+On January 20, 2020, the first COVID-19 case was confirmed in South Korea.
+After a rapid outbreak, the number of incident cases has been consistently decreasing since early March;
+this decrease has been widely attributed to its intensive testing.
+We report here on the likely role of social distancing in reducing transmission in South Korea.
+Our analysis suggests that transmission may still be persisting in some regions.
+
+\section*{Keywords}
+
+coronavirus disease, COVID-19, SARS-CoV-2, South Korea, social distancing
+
+\section*{Contribution}
+
+Data collection: SWP; conceptualization: SWP, CV; analysis: SWP; first draft: SWP, JD. All authors contributed to the writing and approval of the final report.
+
+\section*{Data availability}
+
+All transcribed data and original reports used in this analysis are stored in a publicly available GitHub repository: \url{https://github.com/parksw3/COVID19-Korea}.
+All code and traffic data are stored in a separate GitHub repository: \url{https://github.com/parksw3/Korea-analysis}.
+
+\section*{Funding statement}
+
+JD was supported from the Canadian Institutes of Health Research. The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+
+\section*{Conflict of interest}
+
+None declared.
 

### 2020-03-26 korea.tex
update!!!!!

@@
 \textbf{Comparison of epidemiological and traffic data from Daegu and Seoul.}
-Solid lines represent the daily metro traffic volume in 2020 (red) and mean daily traffic volume between 2017--2019 (black).
+Solid lines represent the daily metro traffic volume in 2020 (red) and mean daily metro traffic volume between 2017--2019 (black).
 Daily traffic from previous years have been shifted by 1--3 days to align day of the weeks.
@@
 
-In Seoul, estimates of $\mathcal R_t$ decreases slightly but remain around 1 (\fref{eff}D).
+In Seoul, estimates of $\mathcal R_t$ decrease slightly but remain around 1 (\fref{eff}D).
 Our analysis suggests that social distancing in Seoul was less intense, and this could be why reduction in spread was less sharp.
@@
 Our analysis reveals that the epidemic may still persist in other regions, including Seoul and Gyeonggi-do;
-reports from Seoul and Gyeonggi-do (around 10 cases almost every day between March 11--24, 2020) provide further support for our conclusion \citep{kcdc}.
+recent reports from Seoul and Gyeonggi-do (around 10 new cases almost every day between March 11--24, 2020) provide further support for our conclusion \citep{kcdc}.
 Unless the reproduction number can be reduced below 1 in all regions, small outbreaks may continue to occur in South Korea.

### 2020-03-26 supp.tex
update!!!!!

@@
 The daily number of reported cases by the KCDC may be slightly different from the reports by each city's government as some cases may be transferred after they are confirmed.
-The sum of daily number of reported cases by the KCDC may be also slightly different from the cumulative number of cases reported the KCDC because it does not reflect possible location changes of the confirmed cases.
+The sum of daily number of reported cases by the KCDC may be also slightly different from the cumulative number of cases reported the KCDC because it does not reflect possible location changes of the confirmed cases after reporting.
 

### 2020-03-26 supp_euro.tex
update!!!!!

@@
 \newline
-\\
-Sang Woo Park\textsuperscript{1,*}
-Kaiyuan Sun\textsuperscript{2}
-C\'ecile Viboud\textsuperscript{2}
-Bryan T.\ Grenfell\textsuperscript{1,2,3}
-Jonathan Dushoff\textsuperscript{4,5,6}
-\\
-\bigskip
-\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
-\\
-\textbf{2} Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
-\\
-\textbf{3} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
-\\
-\textbf{4} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
-\\
-\textbf{5} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
-\\
-\textbf{6} Department of Biology, McMaster University, Hamilton, ON, Canada
-\\
-\bigskip
 
-*Corresponding author: swp2@princeton.edu\\
-This article does not necessarily represent the views of the NIH or the US government.
+This supplementary material is hosted by Eurosurveillance as supporting information alongside the article ``Potential roles of social distancing in mitigating the spread of coronavirus disease 2019 (COVID-19) in South Korea'', on behalf of the authors, who remain responsible for the accuracy and appropriateness of the content.
+The same standards for ethics, copyright, attributions and permissions as for the article apply.
+Supplements are not edited by Eurosurveillance and the journal is not responsible for the maintenance of any links or email addresses provided therein.
+
 \end{flushleft}

### 2020-03-27 title.tex
update!!!!!
tweak

@@
 
+\section*{Keywords}
+
+coronavirus disease, COVID-19, SARS-CoV-2, South Korea, social distancing
+
 \section*{Contribution}
@@
 
-\section*{Data availability}
+\section*{Funding statement}
 
-All transcribed data and original reports used in this analysis are stored in a publicly available GitHub repository: \url{https://github.com/parksw3/COVID19-Korea}.
-All code and traffic data are stored in a separate GitHub repository: \url{https://github.com/parksw3/Korea-analysis}.
+JD was supported from the Canadian Institutes of Health Research. The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+
+\section*{Conflict of interest}
+
+None declared.

### 2020-03-27 figure_compare_R_t.R

@@ -259,5 +259,5 @@ g1 <- ggplot(bind_rows(mutate(daegu_merge, region="Daegu"), mutate(seoul_merge, 
     legend.title = element_blank()
   )
 
-ggsave("traffic.pdf", g1, width=6, height=6)
+ggsave("traffic.pdf", g1, width=5, height=5)
 ggsave("traffic.png", g1, width=6, height=4)

### 2020-03-28 figure_R_t_daegu.R
misc edits

@@ -116,3 +116,6 @@ g2 <- ggplot(rt_all) +
 gtot <- arrangeGrob(g1, g2, nrow=2)
 
 ggsave("figure_R_t_daegu.pdf", gtot, width=6, height=5)
+
+ggsave("AppendixFigure5A.jpg", g1 + theme(plot.title = element_blank()), width=6,height=3, dpi=600)
+ggsave("AppendixFigure5B.jpg", g2 + theme(plot.title = element_blank()), width=6,height=3, dpi=600)

### 2020-03-28 figure_R_t_gyeongbuk.R

@@ -115,3 +115,6 @@ g2 <- ggplot(rt_all) +
 gtot <- arrangeGrob(g1, g2, nrow=2)
 
 ggsave("figure_R_t_gyeongbuk.pdf", gtot, width=6, height=5)
+
+ggsave("AppendixFigure7A.jpg", g1 + theme(plot.title = element_blank()), width=6,height=3, dpi=600)
+ggsave("AppendixFigure7B.jpg", g2 + theme(plot.title = element_blank()), width=6,height=3, dpi=600)

### 2020-03-28 figure_R_t_gyeonggi.R

@@ -114,3 +114,6 @@ g2 <- ggplot(rt_all) +
 gtot <- arrangeGrob(g1, g2, nrow=2)
 
 ggsave("figure_R_t_gyeonggi.pdf", gtot, width=6, height=5)
+
+ggsave("AppendixFigure8A.jpg", g1 + theme(plot.title = element_blank()), width=6,height=3, dpi=600)
+ggsave("AppendixFigure8B.jpg", g2 + theme(plot.title = element_blank()), width=6,height=3, dpi=600)

### 2020-03-28 figure_R_t_seoul.R

@@ -114,3 +114,6 @@ g2 <- ggplot(rt_all) +
 gseoul <- arrangeGrob(g1, g2, nrow=2)
 
 ggsave("figure_R_t_seoul.pdf", gseoul, width=6, height=5)
+
+ggsave("AppendixFigure6A.jpg", g1 + theme(plot.title = element_blank()), width=6,height=3, dpi=600)
+ggsave("AppendixFigure6B.jpg", g2 + theme(plot.title = element_blank()), width=6,height=3, dpi=600)

### 2020-03-28 figure_compare_R_t.R

@@ -245,6 +245,11 @@ gtot <- arrangeGrob(g1, g2, g3, g4, nrow=2, widths=c(1.1, 1))
 ggsave("figure_compare_R_t.pdf", gtot, width=8, height=8)
 ggsave("figure_compare_R_t.png", gtot, width=15, height=7)
 
+ggsave("figure2A.jpg", g1 + theme(plot.title = element_blank()), width=4, height=4, dpi=600)
+ggsave("figure2B.jpg", g2 + theme(plot.title = element_blank()), width=4, height=4, dpi=600)
+ggsave("figure2C.jpg", g3 + theme(plot.title = element_blank()), width=4, height=4, dpi=600)
+ggsave("figure2D.jpg", g4 + theme(plot.title = element_blank()), width=4, height=4, dpi=600)
+
 g1 <- ggplot(bind_rows(mutate(daegu_merge, region="Daegu"), mutate(seoul_merge, region="Seoul"))) +
   geom_point(aes(traffic, median, shape=region, col=region), size=3) +
   scale_x_continuous("(Daily traffic, 2020)/(Mean daily traffic, 2017 - 2019)") +
@@ -261,3 +266,4 @@ g1 <- ggplot(bind_rows(mutate(daegu_merge, region="Daegu"), mutate(seoul_merge, 
 
 ggsave("traffic.pdf", g1, width=5, height=5)
 ggsave("traffic.png", g1, width=6, height=4)
+ggsave("AppendixFigure2.jpg", g1, width=6, height=4, dpi=600)

### 2020-03-28 figure_compare_report.R

@@ -154,3 +154,6 @@ gtot <- arrangeGrob(g1, g2, nrow=1, widths=c(1.1, 1))
 
 ggsave("figure_compare_report.pdf", gtot, width=8, height=3.5)
 ggsave("figure_compare_report.png", gtot, width=8, height=3.5)
+
+ggsave("figure1A.jpg", g1+theme(plot.title = element_blank()), width=4, height=4, dpi=600)
+ggsave("figure1B.jpg", g2+theme(plot.title = element_blank()), width=4, height=4, dpi=600)

### 2020-03-28 figure_cross.R

@@ -175,3 +175,5 @@ g2 <- ggplot(c2data) +
 gtot <- arrangeGrob(g1, g2, nrow=1)
 
 ggsave("figure_cross.pdf", gtot, width=8, height=4)
+ggsave("AppendixFigure3A.jpg", g1 + theme(plot.title = element_blank()), width=4, height=4, dpi=600)
+ggsave("AppendixFigure3B.jpg", g2 + theme(plot.title = element_blank()), width=4, height=4, dpi=600)

### 2020-07-23 doc/response.tex
ok

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}

\usepackage{color}

\usepackage{xspace}

\usepackage{lmodern}
\usepackage{amssymb,amsmath}

\usepackage[pdfencoding=auto, psdextra]{hyperref}

\newcommand{\rR}{\mbox{$r$--$\cal R$}}
\newcommand{\RR}{\ensuremath{{\cal R}}}
\newcommand{\RRhat}{\ensuremath{{\hat \cal R}}}
\newcommand{\Rx}[1]{\ensuremath{{\cal R}_{#1}}} 
\newcommand{\Ro}{\ensuremath{{\mathcal R}_{0}}\xspace}
\newcommand{\Rpool}{\ensuremath{{\mathcal R}_{\textrm{\tiny{pool}}}}\xspace}
\newcommand{\Reff}{\Rx{\mathit{eff}}}
\newcommand{\Tc}{\ensuremath{C}}

\newcommand{\rev}{\subsection*}
\newcommand{\revtext}{\textsf}
\setlength{\parskip}{\baselineskip}
\setlength{\parindent}{0em}

\newcommand{\comment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
\newcommand{\jd}[1]{\comment{cyan}{JD}{#1}}
\newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}
\newcommand{\dc}[1]{\comment{blue}{DC}{#1}}
\newcommand{\jsw}[1]{\comment{green}{JSW}{#1}}
\newcommand{\hotcomment}[1]{\comment{red}{HOT}{#1}}

\begin{document}

\noindent Dear Editor:

Thank you for the chance to revise and resubmit our manuscript. 
We have made minor revisions to our manuscript to address reviewer's comments and to meet the word limit.
Below please find our responses to reviewers.

\rev{Reviewer \#1}

\revtext{The manuscript by Park et al. provided a measure of social distancing in South Korea and illustrated how it is connected with the reduced transmission. The analyses were robust and supported by relevant sensitivity analyses, considering change in testing criteria. The study illustrated clearly how traffic data can be used as a measure social distancing.}

\revtext{Comments to the authors}

\revtext{1.	Please clarify if Rt measured transmission before or after time t in the Figure caption. How would that affect interpretation of the results?}

We quantify the instantaneous reproduction number, which measures transmission at time $t$. This is the Rt that one should use to quantify real-time changes in transmission. Therefore, the use of this Rt does not affect interpretation of the results. We have added the following text in the Figure caption:

``The instantaneous reproduction number $\mathcal R_t$ reflects transmission dynamics at time $t$''

\revtext{2.	Figure 2C. Could the authors explain why Rt started to decrease even 1 week before confirmation of the first COVID-19 case in Daegu? }

This is likely due to our resampling method, which is known to over-smooth the incidence curve and estimates of $\mathcal R_t$. Nonetheless, it has been shown to capture the qualitative changes in $\mathcal R_t$. We have added the following text to acknowledge this limitation with appropriate citations:

`` The initial decrease in $\mathcal R_t$ is likely to be caused by our resampling method for infection times for each reported case, which over-smoothes the incidence curve and $\mathcal R_t$ estimates (7).''

``Finally, the method of resampling infection time over-smoothes $\mathcal R_t$ estimates, but still captures qualitatively changes in $\mathcal R_t$ (7).''

\revtext{3.	L79, it may be helpful to describe in the introduction if there were any official recommendations concerning social distancing or traffic}

We have added the following text to the introduction. We were not able to find any official recommendations concerning traffic.

``Several measures were implemented to prevent the spread of COVID-19. On February 20,
the Daegu Metropolitan City Government recommended refraining from going outside and
wearing masks in everyday life (2). On February 23, the national alert level was raised to the
highest level (1) and the beginning of school semesters were delayed (3). Intensive testing and
contact tracing efforts further allowed rapid identification and isolation cases and reduction of
onward transmission (4).''

\rev{EID Editorial Board Comments:}

\revtext{Even though the limits for dispatch articles in EID are 1200 words in length, 15 references and 2 tables and figures, note that EID does allow appendices to accompany dispatch articles, so as to accommodate supplemental figures, tables, and results.}

\revtext{Do not embed figures in MS Word document. Resubmit supplemental figures as separate JPEG, PDF, PPT or TIF files at 300 DPI or greater resolution. For supplemental figures 1, 3, 4, and 9, resubmit panels as separate files.}

We have revised the manuscript to meet the word limits and style guidelines.

\end{document}

### 2020-07-23 doc/response.tex
update

@@
 
+We have also added the following text in the study section as we think it better suits there:
+
+``To our knowledge, social distancing was first recommended by the KCDC on February 29 (1), and there were no official guidelines regarding public transportations, suggesting that distancing was, in part, voluntary.''
+
 \rev{EID Editorial Board Comments:}

### 2020-07-24 doc/response.tex
update

@@
 
-``Finally, the method of resampling infection time over-smoothes $\mathcal R_t$ estimates, but still captures qualitatively changes in $\mathcal R_t$ (7).''
+``Finally, while the method of resampling infection time can capture qualitative changes in $\mathcal R_t$, estimates of $\mathcal R_t$ can be over-smoothed and should be interpreted with care (7).
+Nonetheless, our estimates of $\mathcal R_t$ are broadly consistent with previous estimates (11).''
 
@@
 ``Several measures were implemented to prevent the spread of COVID-19. On February 20,
-the Daegu Metropolitan City Government recommended refraining from going outside and
+the Daegu Metropolitan City Government recommended refraining from going outside, and
 wearing masks in everyday life (2). On February 23, the national alert level was raised to the

### 2020-07-28 doc/response.tex
more updates

@@
 
-We quantify the instantaneous reproduction number, which measures transmission at time $t$.
-This is the Rt that one should use to quantify real-time changes in transmission.
-Therefore, the use of this Rt does not affect interpretation of the results.
+We quantify instantaneous reproduction number $\mathcal R_t$ using a proxy for incidence of infection, which measures transmission at time $t$.
+This $\mathcal R_t$ quantifies real-time changes in transmission.
+Therefore, the use of this $\mathcal R_t$ does not affect interpretation of the results. 
 We have added the following text in the Figure caption:
@@
 ``The instantaneous reproduction number $\mathcal R_t$ reflects transmission dynamics at time $t$''
+
+We also tried to explain this more carefully in the main text:
+
+``We reconstructed the time series of a proxy for incidence of infection $I_t$, representing the number of individuals who became infected at time $t$ and reported later, and estimated the instantaneous reproduction number, $\mathcal R_t$, defined as the average number of secondary infections caused by an infected individual, given conditions at time $t$ (5).''
 

### 2020-07-28 figure_supp/figure_R_t_all.R

@@ -1,12 +1,13 @@
 library(dplyr)
 library(tidyr)
-library(ggplot2); theme_set(theme_bw())
-source("wquant.R")
+library(ggplot2); theme_set(theme_bw(base_size = 12,
+                                     base_family = "Times"))
+source("../R/wquant.R")
 
-load("R_t_gyeongbuk_censor_detect.rda")
-load("R_t_daegu_censor_detect.rda")
-load("R_t_seoul_censor_detect.rda")
-load("R_t_gyeonggi_censor_detect.rda")
+load("../analysis_R_t/R_t_gyeongbuk_censor_detect.rda")
+load("../analysis_R_t/R_t_daegu_censor_detect.rda")
+load("../analysis_R_t/R_t_seoul_censor_detect.rda")
+load("../analysis_R_t/R_t_gyeonggi_censor_detect.rda")
 
 R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
 
@@ -44,7 +45,7 @@ R_t_all <- bind_rows(
     region=factor(region, levels=c("Daegu", "Seoul", "Gyeongsangbuk-do", "Gyeonggi-do"))
   )
 
-geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
+geo <- read_xlsx("../data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
   mutate(
     date_report=as.Date(date_report)
   ) 
@@ -67,9 +68,10 @@ geo_all <- geo %>% group_by(date_report) %>%
   )
  
 g1 <- ggplot(filter(R_t_all, region=="Daegu")) +
+  annotate("text", x=as.Date("2020-01-21"), y=Inf, label="Daegu", hjust=0, vjust=1) +
   geom_bar(data=filter(geo_all, region=="Daegu"), aes(date_report, value/150), stat="identity", alpha=0.5) +
   geom_hline(yintercept=1, lty=2) +
-  geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.3) +
+  geom_ribbon(aes(date, ymin=lwr, ymax=upr), alpha=0.15) +
   geom_line(aes(date, median)) +
   scale_x_date("Date") +
[…]

### 2020-07-28 figure_supp/figure_R_t_daegu.R

@@ -1,14 +1,14 @@
 library(readxl)
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
-source("wquant.R")
+source("../R/wquant.R")
 
-load("R_t_gyeonggi_censor_detect.rda")
-load("R_t_gyeonggi_censor_raw.rda")
+load("../analysis_R_t/R_t_daegu_censor_detect.rda")
+load("../analysis_R_t/R_t_daegu_censor_raw.rda")
 
 R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
 
-rt_censor_detect <- R_t_gyeonggi_censor_detect %>%
+rt_censor_detect <- R_t_daegu_censor_detect %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   mutate(weight=R0prior(IRt)) %>%
@@ -21,7 +21,7 @@ rt_censor_detect <- R_t_gyeonggi_censor_detect %>%
     group="adjusted for testing criteria"
   )
 
-rt_censor_raw <- R_t_gyeonggi_censor_raw %>%
+rt_censor_raw <- R_t_daegu_censor_raw %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   mutate(weight=R0prior(IRt)) %>%
@@ -40,7 +40,7 @@ rt_all <- bind_rows(rt_censor_detect, rt_censor_raw) %>%
                                 "adjusted for testing criteria"))
   )
 
-geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3)
+geo <- read_xlsx("../data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3)
 
 geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
 
@@ -75,14 +75,15 @@ reconstruct_case_all <- bind_rows(reconstruct_case_detect, reconstruct_case_raw)
   )
 
 g1 <- ggplot(geo) +
-  geom_bar(aes(as.Date(date_report)-0.5, `Gyeonggi-do`), stat="identity", alpha=0.5) +
+  geom_bar(aes(as.Date(date_report)-0.5, Daegu), stat="identity", alpha=0.5) +
   geom_vline(xintercept=as.Date("2020-03-10"), lty=2) +
[…]

### 2020-07-28 figure_supp/figure_R_t_gyeongbuk.R

@@ -1,14 +1,14 @@
 library(readxl)
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
-source("wquant.R")
+source("../R/wquant.R")
 
-load("R_t_seoul_censor_detect.rda")
-load("R_t_seoul_censor_raw.rda")
+load("../analysis_R_t/R_t_gyeongbuk_censor_detect.rda")
+load("../analysis_R_t/R_t_gyeongbuk_censor_raw.rda")
 
 R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
 
-rt_censor_detect <- R_t_seoul_censor_detect %>%
+rt_censor_detect <- R_t_gyeongbuk_censor_detect %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   mutate(weight=R0prior(IRt)) %>%
@@ -21,7 +21,7 @@ rt_censor_detect <- R_t_seoul_censor_detect %>%
     group="adjusted for testing criteria"
   )
 
-rt_censor_raw <- R_t_seoul_censor_raw %>%
+rt_censor_raw <- R_t_gyeongbuk_censor_raw %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   mutate(weight=R0prior(IRt)) %>%
@@ -40,7 +40,7 @@ rt_all <- bind_rows(rt_censor_detect, rt_censor_raw) %>%
                                 "adjusted for testing criteria"))
   )
 
-geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3)
+geo <- read_xlsx("../data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3)
 
 geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
 
@@ -75,15 +75,15 @@ reconstruct_case_all <- bind_rows(reconstruct_case_detect, reconstruct_case_raw)
   )
 
 g1 <- ggplot(geo) +
-  geom_bar(aes(as.Date(date_report)-0.5, Seoul), stat="identity", alpha=0.5) +
+  geom_bar(aes(as.Date(date_report)-0.5, `Gyeongsangbuk-do`), stat="identity", alpha=0.5) +
+  geom_vline(xintercept=as.Date("2020-03-10"), lty=2) +
[…]

### 2020-07-28 figure_supp/figure_R_t_seoul.R

@@ -1,14 +1,14 @@
 library(readxl)
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
-source("wquant.R")
+source("../R/wquant.R")
 
-load("R_t_daegu_censor_detect.rda")
-load("R_t_daegu_censor_raw.rda")
+load("../analysis_R_t/R_t_seoul_censor_detect.rda")
+load("../analysis_R_t/R_t_seoul_censor_raw.rda")
 
 R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
 
-rt_censor_detect <- R_t_daegu_censor_detect %>%
+rt_censor_detect <- R_t_seoul_censor_detect %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   mutate(weight=R0prior(IRt)) %>%
@@ -21,7 +21,7 @@ rt_censor_detect <- R_t_daegu_censor_detect %>%
     group="adjusted for testing criteria"
   )
 
-rt_censor_raw <- R_t_daegu_censor_raw %>%
+rt_censor_raw <- R_t_seoul_censor_raw %>%
   bind_rows(.id="sim") %>%
   group_by(date) %>%
   mutate(weight=R0prior(IRt)) %>%
@@ -40,7 +40,7 @@ rt_all <- bind_rows(rt_censor_detect, rt_censor_raw) %>%
                                 "adjusted for testing criteria"))
   )
 
-geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3)
+geo <- read_xlsx("../data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3)
 
 geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] <- geo$date_report[geo$time_report==16 & !is.na(geo$time_report)] + 1
 
@@ -75,16 +75,15 @@ reconstruct_case_all <- bind_rows(reconstruct_case_detect, reconstruct_case_raw)
   )
 
 g1 <- ggplot(geo) +
-  geom_bar(aes(as.Date(date_report)-0.5, Daegu), stat="identity", alpha=0.5) +
-  geom_vline(xintercept=as.Date("2020-03-10"), lty=2) +
+  geom_bar(aes(as.Date(date_report)-0.5, Seoul), stat="identity", alpha=0.5) +
[…]

### 2020-07-28 figure_supp/figure_R_t_seoul_linelist.R

@@ -1,10 +1,10 @@
 library(readxl)
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
-source("wquant.R")
+source("../R/wquant.R")
 
-load("R_t_seoul_censor_detect.rda")
-load("R_t_seoul_censor_linelist.rda")
+load("../analysis_R_t/R_t_seoul_censor_detect.rda")
+load("../analysis_R_t/R_t_seoul_censor_linelist.rda")
 
 R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
 
@@ -31,16 +31,16 @@ rt_censor_linelist <- R_t_seoul_censor_linelist %>%
     upr=wquant(IRt, weights=weight,0.975)
   ) %>%
   mutate(
-    group="analysis based on the line list"
+    group="analysis based on public line list"
   )
 
 rt_all <- bind_rows(rt_censor_detect, rt_censor_linelist) %>%
   mutate(
     group=factor(group, level=c("main analysis",
-                                "analysis based on the line list"))
+                                "analysis based on public line list"))
   )
 
-covid_seoul <- read_xlsx("data/COVID19-Korea-Regional-2020-03-16.xlsx", sheet=1, na="NA") %>%
+covid_seoul <- read_xlsx("../data/COVID19-Seoul.xlsx", sheet=1, na="NA") %>%
   mutate(
     date_onset=as.Date(date_onset)
   ) %>%
@@ -70,13 +70,13 @@ reconstruct_case_linelist <- reconstruct_censor_linelist %>%
     upr=quantile(case, 0.975)
   ) %>%
   mutate(
-    group="analysis based on the line list"
+    group="analysis based on public line list"
   )
 
 reconstruct_case_all <- bind_rows(reconstruct_case_detect, reconstruct_case_linelist) %>%
   mutate(
     group=factor(group, level=c("main analysis",
-                                "analysis based on the line list"))
+                                "analysis based on public line list"))
   )
 
 g1 <- ggplot(covid_seoul) +
@@ -84,16 +84,15 @@ g1 <- ggplot(covid_seoul) +
   geom_ribbon(data=reconstruct_case_linelist, aes(date, ymin=lwr, ymax=upr), alpha=0.3, fill=2, col=2) +
[…]

### 2020-07-28 figure_supp/figure_cross.R

@@ -1,19 +1,20 @@
-library(ggplot2); theme_set(theme_bw())
+library(ggplot2); theme_set(theme_bw(base_size = 12,
+                                     base_family = "Times"))
 library(gridExtra)
 library(lubridate)
 library(dplyr)
 library(tidyr)
 library(readxl)
-source("wquant.R")
+source("../R/wquant.R")
 
 R0prior <- function(x) dgamma(x, shape=(2.6/2)^2, rate=(2.6/2)^2/2.6)
 
-load("R_t_daegu_censor_detect.rda")
-load("R_t_seoul_censor_detect.rda")
-load("traffic_daegu.rda")
-load("traffic_seoul.rda")
+load("../analysis_R_t/R_t_daegu_censor_detect.rda")
+load("../analysis_R_t/R_t_seoul_censor_detect.rda")
+load("../data_processed/traffic_daegu.rda")
+load("../data_processed/traffic_seoul.rda")
 
-geo <- read_xlsx("data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
+geo <- read_xlsx("../data/COVID19-Korea-2020-03-16.xlsx", na="NA", sheet=3) %>%
   mutate(
     date_report=as.Date(date_report)
   ) 
@@ -37,7 +38,7 @@ traffic_daegu1 <- dplyr::select(ungroup(filter(traffic_daegu,year==2017)), -월,
     wday=wday(date)
   ) %>%
   filter(
-    date >= as.Date("2017-01-20")+3, date <= as.Date("2017-03-01")+3
+    date >= as.Date("2017-01-20")+3, date <= as.Date("2017-03-17")+3
   )
 
 traffic_daegu2 <- dplyr::select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) %>%
@@ -45,7 +46,7 @@ traffic_daegu2 <- dplyr::select(ungroup(filter(traffic_daegu,year==2018)), -월,
     wday=wday(date)
   ) %>%
   filter(
-    date >= as.Date("2018-01-20")+2, date <= as.Date("2018-03-01")+2
+    date >= as.Date("2018-01-20")+2, date <= as.Date("2018-03-17")+2
   )
 
 traffic_daegu3 <- dplyr::select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) %>%
[…]

### 2020-07-28 regression.R

@@ -1,103 +1,0 @@
-load("R_t_daegu_censor.rda")
-load("R_t_seoul_censor.rda")
-load("traffic_daegu.rda")
-load("traffic_seoul.rda")
-
-load("pred_report_delay.rda")
-
-traffic_daegu1 <- select(ungroup(filter(traffic_daegu,year==2017)), -월, -일) %>%
-  mutate(
-    wday=wday(date)
-  ) %>%
-  filter(
-    date >= as.Date("2017-02-02")+3, date <= as.Date("2017-03-01")+3
-  )
-
-traffic_daegu2 <- select(ungroup(filter(traffic_daegu,year==2018)), -월, -일) %>%
-  mutate(
-    wday=wday(date)
-  ) %>%
-  filter(
-    date >= as.Date("2018-02-02")+2, date <= as.Date("2018-03-01")+2
-  )
-
-traffic_daegu3 <- select(ungroup(filter(traffic_daegu,year==2019)), -월, -일) %>%
-  mutate(
-    wday=wday(date)
-  ) %>%
-  filter(
-    date >= as.Date("2019-02-02")+1, date <= as.Date("2019-03-01")+1
-  )
-
-traffic_daegu4 <- select(ungroup(filter(traffic_daegu,year==2020)), -월, -일) %>%
-  mutate(
-    wday=wday(date)
-  ) %>%
-  filter(
-    date >= as.Date("2020-02-02"), date <= as.Date("2020-02-29")
-  )
-
-rt_daegu <- R_t_daegu_censor %>%
-  bind_rows(.id="sim") %>%
-  group_by(date) %>%
-  summarize(
-    median=median(IRt, na.rm=TRUE),
-    lwr=quantile(IRt, 0.025, na.rm=TRUE),
-    upr=quantile(IRt, 0.975, na.rm=TRUE),
-    location="Daegu"
-  ) %>%
-  filter(date <= as.Date("2020-02-29")) %>%
-  mutate(
-    traffic=traffic_daegu4$total/((traffic_daegu1$total + traffic_daegu2$total + traffic_daegu3$total)/3)
-  )
-
-traffic_seoul1 <- ungroup(filter(traffic_seoul,year==2017)) %>%
-  mutate(
-    wday=wday(date)
-  ) %>%
-  filter(
-    date >= as.Date("2017-02-02")+3, date <= as.Date("2017-03-11")+3
-  )
-
[…]
